### 2023-04-20 README.Rmd
first commit

---
title: "robspack"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![GitHub commit
activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/robspack)
[![R build
status](https://github.com/ibarraespinosa/robspack/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
[![R build
status](https://github.com/ibarraespinosa/robspack/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)

[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of green house gases observations

`robspack` only depends on `data.table`, which is basically parallel C, 
so it can be installed in any machine.

## installation

```{r, eval = F}
remotes::install_github("ibarraespinosa/robspack")
```

```{r, include = T, message=F, warning=F}
library(robspack)
library(data.table)
```

## ObsPack summary

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

```{r}

categories <- c(
  "aircraft-pfp",
  "aircraft-insitu",
  "surface-insitu",
  "aircore",
  "surface-pfp",
  "tower-insitu",
  "shipboard-insitu",
  "flask"
)
obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs)

```

There are 362 files in the ObsPack directory.
The printed information also shows the total at the bottom,
as the sum of the individual file by sector. This
is to ensure that the sum of files is equal to the
total number of files found, shown at the top.
furthermore, the printed information
also shows that there are 136 files with the
`agl` explicitly mentioned in the name of the file.

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".
```{r}
dft <- obs_table(index = index,
                 categories = "tower-insitu",
                 verbose = FALSE)
dft[site_code == "SCT", ]$timeUTC |> 
  range()
```

### Read data

Once the summary is built, the function `obs_read`
will read the files available in the
index file previously generated.
Here we selected the category "tower-insitu".
The argument verbose prints which files are being read each time,
by default.
At the end, this function prints the total number
of observations by type of altitude (agl or asl).

```{r}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

```

We added a function to plot the data read from ObsPack.
The y-axis is the field `value` and the 
x-axis is by default `time`. The data illustrated sorted by
color is the field `site_code`, with the default
number of 3 sites. The argument `pal` is to define the
color palette, used by the internally imported function
`cptcity::cpt`.

```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
```

Here we can see `r round(nrow(df)/1000000, 2)` million observations 
for `tower-insitu`. These observations are made between 
`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
of the altitude and type is critical. The approach used here 
consists of:

1. Identify `agl` from the name of the tile.
2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
3. If `agl` is not present, `agl = altitude - elevation`.
4. If there are some NA in elevation, will result some NA in `agl`
5. A new column is added named `altitude_final` to store `agl` or `asl`
6. Another column named `type_altitude` is added to identify `agl` or `asl`.
7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"

### Filtering

ObsPack includes global observations and sometimes we need to extract data
for a specific region and periods of time.
In this part we include spatial and temporal parameters
to filter data.
The year of interest is 2020, but we also included
December of 2019 and January of 2021.
At this stage, we can apply the spatial filter
by using the coordinates.

```{r}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14:15

yy <- 2020
df <- rbind(df[year == yy - 1 & month == 12],
            df[year == yy],
            df[year == yy + 1 & month == 1])

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

After filtering by space and time, we have `r round(nrow(df), 2)`
million observations. Towers can have observations at different
heights. Here we need to select one site with the observations
registered at the highest height. The column
with the height is named `altitude_final` and the
max altitude was named `max_altitude`.

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
```

### Time

Here we need to start time columns.
The function `obs_addtime` adds time columns
`timeUTC`, `timeUTC_start` which shows the start time
of each observation and `timeUTC_end`
which shows the end time for each observation.
Then we need to identify the local time with the function
`add_ltime`. This is important because to identifying
observations in the evening in local time.
`add_ltime` uses two methods, first identify
the time difference with utc by identifying the
metadata column "site_utc2lst". If this information
is not available, with the aircrafts for instance,
the local time is calculated with an approximation
based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.
Then, the time is cut every two hours.
Now, we identify the local time to select evening hours.

```{r}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

Now there are `r nrow(df3)` observations.
At this point we can calculate the averages of several columns by
the cut time. The function `obs_agg` does this aggregation as shown
in the following lines of code. The argument `gby`
establish the function used to aggregate `cols`, in this case
the function mean by time and altitude. Finally, we
add local time again.

```{r}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value",
                        "latitude",
                        "longitude",
                        "type_altitude",
                        "dif_time",
                        "year_end",
                        "site_utc2lst"),
               verbose = FALSE,
               byalt = TRUE)
df5 <- obs_addltime(df4)
```

Now there are `r nrow(df5)` observations,
`r nrow(df3) - nrow(df5)` less observations.
Here we add the column `max_altitude` to identify the max altitude
by site.

```{r}
df5[,
    max_altitude := max(altitude_final),
    by = site_code]
df5[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()
```

### Saving master as text and csvy

Now that we have all the required information, 
we can save the files. Here, we name the data.frame 
as master, because it contains all the information.
This is important because some 
fields can be used in the future, and for traceability.
For convenience,
time variables are transformed into character
before writing into the disk.
The separation is space " ".

```{r, eval = T}
tmp <- tempfile()
master <- df5
master$timeUTC <- as.character(master$timeUTC)
master$timeUTC_end <- as.character(master$timeUTC_end)
master$local_time <- as.character(master$local_time)

fwrite(master, 
       file =  "paper/tower_insitu_2020.txt",
       sep = " ")
```

The format Comma Separated Value with YAML (CSVY)[^3]
consists in a typical CSV with a YAML header.
The function`obs_write_csvy` includes the argument
`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
the output of the R function `str`, which
provides a vertical summary of the data, known as
structure.

```{r, eval = T}
obs_write_csvy(dt = master,
          notes = "tower 2020",
          out = "paper/tower_insitu_2020.csvy")
```

To check the YAML header we read the first 38 lines of 
the files that were generated. Here we can see
the column names, type of data and first observations.
The YAML header is delimited by the characters "---".

```{r, eval = F}
readLines("paper/tower_insitu_2020.csvy")[1:38]
```

### Saving receptors

We need to filter some columns
from the master files in a new object called receptors.
This is needed because internally we run HYSPLIT [@hy]
using the information from the receptors.
In the case of a tower, we need to select 
observations with the highest 
altitude. The specific columns are selected
as shown on the following code. We are selecting the 
ending times, because later HYSPLIT is run backwards
based on the time of measurement, between ending and
starting times. The columns about time are formatted to
have two characters. For instance, the month 1,
is formatted as "01". We also need to filter for
`type_altitude` equal 0, representing `agl `observations
, or equal to 1, `asl`.

```{r, eval = F}

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour",
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "altitude_final",
                     "type_altitude",
                     "year_end",
                     "month_end",
                     "day_end",
                     "hour_end",
                     "minute_end",
                     "second_end")]
receptor$altitude_final <- round(receptor$altitude_final)
receptor <- obs_format(receptor)

if(nrow(receptor_agl) > 0) {
  fwrite(x = receptor_agl,
         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
  sep = " ")}

if(nrow(receptor_asl) > 0) {
  fwrite(x = receptor_asl,
         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
  sep = " ")}

```

## Recommendation for other applications
The approach to generate receptors depends on each type of 
observation and other considerations. For instance, aircraft with 
continuous observations at each second can be filtered and averaged 
every 20 seconds. In that way, the footprints are still 
representative and it would not be necessary to run HYSPLIT 
every second. Of course, it depends on the application 
and objective of the study. For this manuscript, we are 
presenting the generation of receptors based on tower observations.

## Application for other sectors
In this package we are sharing scripts to process other sectors
The scripts are available in the path 
`https://github.com/ibarraespinosa/robspack/tree/main/rscripts`

### 2023-04-20 paper/paper.Rmd
first commit

--- 2023-04-20 README.Rmd
@@
 ---
-title: "robspack"
-output: github_document
+# Example from https://joss.readthedocs.io/en/latest/submitting.html
+title: 'rObsPack: An R package for the NOAA global observations ObsPack methane GLOBALVIEW+'
+tags:
+  - R
+  - NOAA
+  - CH4
+  - methane
+  - ObsPack
+authors:
+  - name: Sergio Ibarra-Espinosa
+    orcid: 0000-0002-3162-1905
+    equal-contrib: true
+    affiliation: "1, 2"
+  - name: Lei Hu
+    orcid: 0000-0002-5624-3832
+    equal-contrib: true
+    affiliation: "2"
+affiliations:
+ - name: Cooperative Institute for Research in Environmental Sciences, University of Colorado Boulder, Boulder, CO, USA
+   index: 1
+ - name: NOAA Global Monitoring Laboratory, Boulder, CO, USA
+   index: 2
+citation_author: Ibarra-Espinosa and Hu
+date: 04 April 2023
+year: 2023
+bibliography: paper.bib
+output: rticles::joss_article
+csl: apa.csl
+journal: JOSS
 ---
+<!-- output: rticles::joss_article -->
+<!-- output: word_document -->
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Summary
 
-![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/robspack)
-[![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
-[![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
+Observations of greenhouse gas emissions are critical to monitor 
+the state of the atmosphere, quantify present and historical 
+emissions, and understand global climate change. Observation 
+Package (ObsPack) is a specific format to deliver atmospheric
+measurements data of trace gasses [@masarie2014obspack]. 
+The NOAA Global Monitoring Laboratory (GML) manages a 
+global network of atmospheric observations compiled and 
+delivered to the public as ObsPack CH4 GLOBALVIEW+. This 
+product provides access to global methane observations made 
+in surface, tower, aircrafts, aircores and ships. Processing 
+global data can be cumbersome since they are provided as 
+multiple text and NetCDF files. The text files consist of a 
+commented metadata followed with the tabulated data which can 
+have different headers across files. The metadata can have critical
+information which determines the nature of the file itself, such as
+altitude of the site, local hour, laboratory and more. Below the
+metadata, starts the section with tabulated data, with different
+header across files.
 
-[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of green house gases observations
+# Statement of need
 
-`robspack` only depends on `data.table`, which is basically parallel C, 
-so it can be installed in any machine.
+In order to analyze greenhouse gasses measurements we developed the 
+R package `robspack`, which reads and processes NOAA/GML ObsPack
+GLOBALVIEW+ CH4 data. This package imports functions from 
+`data.table` R package, which contains C bindings with 
+parallel implementation via Open-MP [@dt]. 
+`data.table` is faster than other Python, Julia and R 
+implementations for data-science, providing a strong basis
+for `robspack`. The main objective of `robspack` is to
+integrate different CH4 ObsPack text files in a long 
+format, adding fields for specific metadata information,
+with a tidy structure [@Silge2016]. Hence, the data 
+can easily be processed with `data-table`, allowing
+to produce plots with `ggplot2` and other R packages [@g2].
 
-## installation
+The steps followed by `robspack` are described in the 
+following lines. First, it is necessary to create a 
+summary of the data , read all the file names, identify 
+key-words, sectors and full paths. The data is read
+iterating on each file for each category such as aircraft, tower, 
+surface, flasks, ships and aircore. It is necessary to
+add UTC, local time and calculate altitude above ground 
+level (agl) from the data. For continuous ground sites, 
+afternoon observations are selected and averaged to reduce
+correlations and noise. Given the high frequency of 
+observations of data, usually every second, aircraft data is
+usually averaged every 20 seconds. Once the data is processed
+for each category, it is merged in a long format. The 
+installation and functions created to perform these tasks 
+are shown in the following sections.
+
+# Functions
+
+The `robspack` functions are shown in the table 1.
+
+Table 1. Functions and classes in `robspack`.
+
+| Function              | Description                                       |
+|-----------------------|---------------------------------------------------|
+| `invfile`             | Class with `print`, `summary` and `plot` methods  |
+| `obs_addltime()`      | Add local time based on metadata and longitude    |
+| `obs_addtime()`       | Add UTC time                                      |
+| `obs_addzero()`       | Format numbers to match specific requirements     |
+| `obs_agg()`           | Aggregates ObsPack by time                        |
+| `obs_find_receptors()`| Find expected receptors and NetCDF files          |
+| `obs_format()`        | Format for some columns of data.table             |
+| `obs_freq()`          | Return numeric vector in intervals                |
+| `obs_invfiles()`      | Construct `invfile` objects                       |
+| `obs_list.dt()`       | Rbind list of data.frames with different names    |
+| `obs_meta()`          | Reads ObsPack metadata                            |
+| `obs_out()`           | Outersect, opposed as intersect                   |
+| `obs_rbind()`         | Rbind data.frames with different names            |
+| `obs_read()`          | Read files, and add metadata as columns           |
+| `obs_read_csvy()`     | Read csvy file and prints yaml header             |
+| `obs_roundtime()`     | Round seconds from "POSIXct" "POSIXt" classes     |
+| `obs_summary()`       | Construct summary of ObsPack as a data.frame      |
+| `obs_table()`         | Return a data.frame with summary of data          |
+| `obs_trunc()`         | Trunc numbers with a desired number of decimals   |
+| `obs_write()`         | Write CSVY to disk, YAML followed by tabulated    |
+
+# Installation
+
+To install `robspack`, the user must have installed the
+R package `remotes` and run the following script. 
+This process will install all the required 
+dependencies, such as `data.table`, `cptcity`, an R package with
+more than 7000 color palettes, and `lubridate`, a package to manage 
+time and dates [@cpt;@lu].
+Then, we call the libraries to load the
+function into the environment.
 
@@
 remotes::install_github("ibarraespinosa/robspack")
-```
-
-```{r, include = T, message=F, warning=F}
 library(robspack)
@@
 ```
+
+```{r, include = F, message=F, warning=F}
+library(robspack)
+library(data.table)
+```
+
+For this manuscript, we are presenting
+the application of `robspack` on tower observations.
 
@@
 )
-obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
+obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
 index <- obs_summary(obs = obs)
@@
                gby = "mean",
-               cols = c("value",
-                        "latitude",
-                        "longitude",
-                        "type_altitude",
-                        "dif_time",
-                        "year_end",
-                        "site_utc2lst"),
+               cols = c("value", "latitude", "longitude", "type_altitude",
+                        "dif_time", "year_end", "site_utc2lst"),
                verbose = FALSE,
@@
 ```{r, eval = T}
-tmp <- tempfile()
 master <- df5
@@
 fwrite(master, 
-       file =  "paper/tower_insitu_2020.txt",
+       file = "tower_insitu_2020.txt",
        sep = " ")
@@
 consists in a typical CSV with a YAML header.
-The function`obs_write_csvy` includes the argument
-`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
+The function`obs_write` includes the argument
+`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write` adds 
 the output of the R function `str`, which
@@
               notes = "tower 2020",
-          out = "paper/tower_insitu_2020.csvy")
+              out = "tower_insitu_2020.csvy")
 ```
@@
 ```{r, eval = F}
-readLines("paper/tower_insitu_2020.csvy")[1:38]
+readLines("tower_insitu_2020.csvy")[1:38]
 ```
@@
                    c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour",
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "altitude_final",
-                     "type_altitude",
-                     "year_end",
-                     "month_end",
-                     "day_end",
-                     "hour_end",
-                     "minute_end",
-                     "second_end")]
+                     "year", "month", "day",
+                     "hour", "minute", "second",
+                     "latitude", "longitude",
+                     "altitude_final", "type_altitude",
+                     "year_end", "month_end", "day_end", "hour_end",
+                     "minute_end", "second_end")]
 receptor$altitude_final <- round(receptor$altitude_final)
@@
 ```
-
 ## Recommendation for other applications
@@
 presenting the generation of receptors based on tower observations. 
-
-## Application for other sectors
-In this package we are sharing scripts to process other sectors
+Furthermore, in this package we are sharing scripts to process other sectors
 The scripts are available in the path 
 `https://github.com/ibarraespinosa/robspack/tree/main/rscripts`
+
+## Conclusion
+
+In this manuscript we presented an `robskpack`,
+an R package to read and process
+$CH_4$ ObsPack GLOBALVIEW+ published by the Global Monitoring 
+Laboratory (GML) from the National Oceanographic and Atmospheric
+Administration (NOAA). `robspack` reads the text data
+which have different headers and organizes them in a common 
+format. Then, this software applies calculations to
+filter observations by time and space. Finally,
+this software generates receptors in a suitable format that 
+allows it to run HYSPLIT and generate footprints. 
+This software does 
+not provide methods to run HYSPLIT, but the user
+can follow the site https://www.ready.noaa.gov/HYSPLIT.php.
+
+[^1]: https://h2oai.github.io/db-benchmark/
+[^2]: https://gml.noaa.gov/ccgg/carbontracker-lagrange/
+[^3]: https://csvy.org/
+
+# Acknowledgements
+
+Funding: This project is funded by the NOAA Climate Program Office AC4
+and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was 
+supported by the NOAA cooperative agreement NA22OAR4320151.
+
+```{r, include=FALSE}
+file.copy(from = "paper.md", to = "../", overwrite = TRUE)
+file.copy(from = "paper.bib", to = "../", overwrite = TRUE)
+```
+
+# References

### 2023-04-20 R/01_index_obs.R

@@ -1,0 +1,100 @@
+#' Summary of the ObsPack files
+#'
+#' This function returns a data.frame index of all files
+#' available in obspack.
+#'
+#' @param obs Path to the Obspack GLOBALview txt data
+#' @param categories character; default are c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask"). The ideia is that,as the file names include these words,
+#' this function identifies which files has these words and add them as columns.
+#' @param lnchar Integer; last nchar, default = 11.
+#' @param out Path to the Obspack index output
+#' @param verbose Logical to show more information
+#' @return A data.frame with with an index of the obspack Globalview.
+#' @rdname obs_summary
+#' @importFrom data.table fwrite ".N" ":=" setDT data.table
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs)
+#' }
+#' }
+obs_index <- function(obs,
+                      categories = c("aircraft-pfp",
+                                     "aircraft-insitu",
+                                     "surface-insitu",
+                                     "tower-insitu",
+                                     "aircore",
+                                     "surface-pfp",
+                                     "shipboard-insitu",
+                                     "flask"),
+                      lnchar = 11,
+                      out = paste0(tempfile(), "_index.csv"),
+                      verbose = TRUE){
+  cat("Please, use `obs_summary` instead")
+
+
+  x <- list.files(obs,
+                  full.names = T)
+
+  na <- list.files(obs,
+                   full.names = F)
+
+  # create index ####
+  # the idea
+  index <- data.table::data.table(id = x, name = na)
+  index$n <- 1:nrow(index)
+  sector <- NULL
+  for(i in seq_along(categories)) {
+    index[grepl(pattern = categories[i], x = index$na),
+          sector := categories[i]]
+  }
+
+
+  cat(paste0("Number of files of index: ", nrow(index), "\n"))
+  xx <- index[, .N, by = sector]
+  dx <- data.table::data.table(sector = c("Total sectors"),
+                               N = sum(xx$N))
+  print(rbind(xx, dx))
+
+
[…]

### 2023-04-20 R/01_summary_obs.R

@@ -1,0 +1,98 @@
+#' Summary of the ObsPack files
+#'
+#' This function returns a data.frame index of all files
+#' available in ObsPack.
+#'
+#' @param obs Path to the Obspack GLOBALview txt data
+#' @param categories character; default are c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask"). The idea is that,as the file names include these words,
+#' this function identifies which files has these words and add them as columns.
+#' @param lnchar Integer; last nchar, default = 11.
+#' @param out Path to the ObsPack index output
+#' @param verbose Logical to show more information
+#' @return A data.frame with with an index of the obspack Globalview.
+#' @importFrom data.table fwrite ".N" ":=" setDT data.table
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs)
+#' }
+#' }
+obs_summary <- function(obs,
+                        categories = c("aircraft-pfp",
+                                       "aircraft-insitu",
+                                       "surface-insitu",
+                                       "tower-insitu",
+                                       "aircore",
+                                       "surface-pfp",
+                                       "shipboard-insitu",
+                                       "flask"),
+                        lnchar = 11,
+                        out = paste0(tempfile(), "_index.csv"),
+                        verbose = TRUE){
+
+
+  x <- list.files(obs,
+                  full.names = T)
+
+  na <- list.files(obs,
+                   full.names = F)
+
+  # create index ####
+  # the idea
+  index <- data.table::data.table(id = x, name = na)
+  index$n <- 1:nrow(index)
+  sector <- NULL
+  for(i in seq_along(categories)) {
+    index[grepl(pattern = categories[i], x = index$na),
+          sector := categories[i]]
+  }
+
+
+  cat(paste0("Number of files of index: ", nrow(index), "\n"))
+  xx <- index[, .N, by = sector]
+  dx <- data.table::data.table(sector = c("Total sectors"),
+                               N = sum(xx$N))
+  print(rbind(xx, dx))
+
+
+  # function to extract last n characters
+  sr <- function(x, n){
[…]

### 2023-04-20 R/02_read_obs.R

@@ -1,0 +1,303 @@
+#' Read obspack
+#'
+#' Each obspack file has a header with metadata and this
+#' function reads selected fields from the metadata and add
+#' them as columns. This new columns are used later to
+#' be filtered
+#'
+#' @param index data.table
+#' @param categories character; ONE category
+#' : of c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask").
+#' @param verbose Logical to show more information
+#' @param n_site_code number of characters extraced from metadata after search
+#' @param n_site_name number of characters extraced from metadata after search
+#' @param n_site_country number of characters extraced from metadata after search
+#' @param n_dataset_project number of characters extraced from metadata after search
+#' @param n_lab number of characters extraced from metadata after search
+#' @param n_scales number of characters extraced from metadata after search
+#' @param n_site_elevation number of characters extraced from metadata after search
+#' @param n_altitude_comment number of characters extraced from metadata after search
+#' @param n_utc number of characters extraced from metadata after search
+#' @param fill_value fill value. Appeared in aoa_aircraft-flask_19_allvalid.txt
+#' @param as_list Logical to return as list
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table fwrite ".N" ":=" rbindlist "%chin%"
+#' @export
+#' @note The identification of the altitude and type is critical.
+#' The approach used here consists of:
+#' 1. Identify agl from the name of the tile.
+#' 2. If magl not present, search dill_values used in elevation and
+#' transform them into NA (not available)
+#' 3. If magl is not present, agl = altitude - elevation
+#' 4. If there are some NA in elevation, will result some NA in agl
+#' 5. A new column is added named `altitude_final` to store agl or asl
+#' 6. Another column named `type_altitude` is added to identify "magl" or "masl"
+#' 7. If there is any case NA in `altitude_final`,
+#' `type_altitude` is "not available"
+#'
+#' Then, the relationship with hysplit is:
+#' \tabular{ccc}{
+#'   type_altitude \tab hysplit  \cr
+#'   magl          \tab agl      \cr
+#'   masl          \tab asl      \cr
+#'   not available \tab f-PBL    \cr
+#'}
+#'
+#' @examples {
+#' # Do not run
[…]

### 2023-04-20 R/03_addtime_obs.R

@@ -1,0 +1,86 @@
+#' Add times into obspack
+#'
+#' This function add timeUTC as POSIX class,
+#' local time and ending sampling time
+#'
+#'
+#'
+#' @param dt obspack data.table
+#' @param verbose obspack data.table
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table year month hour minute second setDT fifelse uniqueN
+#' @note timeUTC is calculated based on the field column start_time,
+#' the timeUTC_end is calculated using this approach:
+#' 1. If the column time_interval is not found, proceed
+#' with the calculation using midpoint_time
+#' 2. Else, use column time_interval
+#'
+#' @export
+#' @examples {
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <-  obs_summary(obs)
+#' dt <- obs_read(index)
+#' dt <- obs_addtime(dt)
+#' }
+obs_addtime <- function(dt,
+                        verbose = TRUE){
+
+  if(!inherits(dt, "data.table")) data.table::setDT(dt)
+
+  if(verbose) cat("Adding timeUTC\n")
+  # Time is already UTC
+  dt$timeUTC <- as.POSIXct(dt$time,
+                           origin = "1970-01-01",
+                           tz = "UTC")
+
+  if(verbose) cat("Adding timeUTC_start\n")
+  # Time is already UTC
+  dt$timeUTC_start <- as.POSIXct(dt$start_time,
+                                 origin = "1970-01-01",
+                                 tz = "UTC")
+
+  if(verbose) cat("Adding timeUTC_end\n")
+  # adding timeUTC_end
+  midpoint_time <- NULL
+  start_time <- NULL
+  timeUTC <- NULL
+  time_interval <- NULL
+  timeUTC_start <- NULL
+  if(any("time_interval" == names(dt))) {
+    dt[, timeUTC_end := fifelse(
+      is.na(time_interval),
+      timeUTC_start + (midpoint_time - start_time)*2,
+      timeUTC_start + time_interval
+    )]
+  } else {
+    #if there is no column named time_interval
+    time_inter <- (dt$midpoint_time - dt$start_time)*2
+    dt$timeUTC_end <- dt$timeUTC_start + time_inter
+  }
+
+  dt$dif_time <- dt$timeUTC_end - dt$timeUTC_start
+  time_warning <- NULL
[…]

### 2023-04-20 R/04_aggregate_obs.R

@@ -1,0 +1,129 @@
+#' Aggregates ObsPack by time
+#'
+#' This function add timeUTC as POSIX class,
+#' local time and ending sampling time
+#'
+#'
+#'
+#' @param dt obspack data.table
+#' @param gby (groupby) Character indicating the function to group the
+#' data. Default is "mean" meaning that all numeric variables
+#' will be averaged by the new_second column established in the `idcol`
+#' @param cols Character which defines columns to be aggregated
+#' @param byalt Logical, to aggregate by altitude or not (used in tower)
+#' @param verbose logical to show more information
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table  second setDT setorderv mday
+#' @importFrom lubridate decimal_date
+
+#' @export
+#' @examples \dontrun{
+#' # Do not run
+#' dt <- obs_addtime(dt)
+#' }
+obs_agg <- function(dt,
+                     gby = "mean",
+                     cols = c("year",
+                               "month",
+                               "day",
+                               "hour",
+                               "minute",
+                               "second",
+                               "time",
+                               "time_decimal",
+                               "value",
+                               "latitude",
+                               "longitude",
+                               "altitude",
+                               "pressure",
+                               "u",
+                               "v",
+                               "temperature",
+                               "type_altitude",
+                               "altitude_final"),
+                     byalt = FALSE,
+                     verbose = TRUE){
+
+  if(!inherits(dt, "data.table")){
+    if(verbose) cat("Transforming your data into data.table\n")
+    data.table::setDT(dt)
+  }
+
+
+  if(!any(names(dt) %in% "timeUTC")) stop("missing `timeUTC``")
+  if(!any(names(dt) %in% "site_code")) stop("missing `site_code``")
+  if(!any(names(dt) %in% "altitude_final")) stop("missing `altitude_final``")
+
+  # Remember aircraft
+  # Only aircraft need to be aggregated every 20 seconds
+  # For hysplit, the ending times are not needed
+  .SD <- NULL
+  timecol <- NULL
+  site_code <- NULL
+  altitude_final <- NULL
+  timeUTC <- NULL
+  lab_1_abbr <- NULL
[…]

### 2023-04-20 R/05_find_missing_receptors.R

@@ -1,0 +1,81 @@
+#' Compares expected receptors
+#'
+#' This function creates a data.frame with the expected
+#' footprints in .nc and compare with the actual footprints.
+#' The idea is verify is there are missing footprints.
+#'
+#' @param path String with path hwere are stored the footprints
+#' @param year numeric number
+#' @param month numeric number
+#' @param day numeric number
+#' @param hour numeric number
+#' @param minute numeric number
+#' @param lat numeric number
+#' @param lon numeric number
+#' @param alt numeric number
+#' @param out outfile path.
+#' @param verbose logical to show more information
+#' @return A data.frame with with expected footprints
+#' @export
+#' @examples \dontrun{
+#' # do not run
+#' p <- "/path/to/continuous/"
+#' # here we have year/month/hysplit*.nc
+#' x <- dt
+#' dt <- obs_find_receptors(p, year, month....)
+#' }
+obs_find_receptors <- function(path,
+                               year,
+                               month,
+                               day,
+                               hour,
+                               minute,
+                               lat,
+                               lon,
+                               alt,
+                               out = paste0(tempfile(), ".csvy"),
+                               verbose = FALSE){
+
+
+  dt <- data.table::data.table(
+    expect = obs_footname(year = year,
+                          month  = month,
+                          day = day,
+                          hour = hour,
+                          minute = minute,
+                          lat = lat,
+                          lon = lon,
+                          alt = alt
+    )
+  )
+
+
+  ld <- list.files(path = path,
+                   pattern = ".nc",
+                   recursive = T,
+                   full.names = T)
+
+  ln <- list.files(path = path,
+                   pattern = ".nc",
+                   recursive = T,
+                   full.names = F)
+
+  if(verbose) print(ld)
+
+  df <- data.table::data.table(file = ln,
+                               expect = ln)
+
+  dx <- merge(dt, df, by = "expect", all.x = T)
+
+  # missing <- NULL
+  dx[, missing := ifelse(is.na(file), T, F)]
+
+
+  if(!missing(out)) {
+    obs_write_csvy(dt = dx,
+              notes = "expected files",
+              out = out)
+  }
+  return(dx)
+
+}

### 2023-04-20 R/06_inv_files.R

@@ -1,0 +1,193 @@
+#' Generate files to perform inverse modeling
+#'
+#' This function return a list of three elements.
+#'
+#' 1. Footprints_hera_hysplit with the  full path to the NetCDF footprints\
+#'
+#' 2. Obs_hysplit concentration associazted iwth each receptor
+#'
+#' 3. Receptor_info_hysplit with receptor info
+#'
+#' @param master data.table master file after filtering obspack
+#' @param nc column name of master which indicate NetCDF footprints
+#' @param path first part of the full path of the .nc footprint files
+#' @param value column name with the pollutant
+#' @param value_factor numeric factor
+#' @param Type "continuous", "flask" or "hatsflask"
+#' @param SubType "aircraft-insitu", "tower-insitu", "surface-insitu" and so on
+#' @param Surface_Elev Site elevation, default 99999
+#' @param Model_agl Model for agl, if missing is column `altitude_final`
+#' @param Data_agl Model for agl, if missing is column `altitude_final`
+#' @param Event Number of event
+#' @param Institution Institution
+#' @param Scale Scale
+#' @param outdir String for the output dir
+#' @return A list of data.frame
+#' @export
+#' @examples \dontrun{
+#' # do not run
+#' }
+obs_invfiles <- function(master,
+                         nc = "nc",
+                         path,
+                         value = "value",
+                         value_factor = 1e+9,
+                         Type = "continuous",
+                         SubType = "tower-insitu",
+                         Surface_Elev = 99999,
+                         Model_agl,
+                         Data_agl,
+                         Event = 99999,
+                         Institution,
+                         Scale,
+                         outdir) {
+
+  # check existing files and then merge
+  x <- list.files(path = path,
+                  pattern = ".nc",
+                  full.names = TRUE,
+                  recursive = TRUE)
+  hs <- pol <- NULL
+  if(!nc %in% names(master)) stop(paste0("Need column ", nc))
+  master$hs <- paste0(path, master$nc)
+
+  if(!value %in% names(master)) stop(paste0("Need column ", value))
+  master$pol <- master[[value]]*value_factor
+
+  # receptor_info_hysplit
+  master$File <- substr(x = master[[nc]],
+                        start = 9,
+                        stop = 100)
+
+  master$Type <- Type
+
+  master$SubType <- SubType
+
+  Model <- "hysplit"
+  master$Model <- Model
+
+  if(!"timeUTC" %in% names(master)) stop("Need column 'timeUTC'")
+
[…]

### 2023-04-20 R/07_meta_obs.R

@@ -1,0 +1,212 @@
+#' Read obspack metadata
+#'
+#' @param index data.table
+#' @param verbose Logical to show more information
+#' @param n_site_code number of characters extraced from metadata after search
+#' @param n_site_name number of characters extraced from metadata after search
+#' @param n_site_latitude number of characters extraced from metadata after search
+#' @param n_site_longitude number of characters extraced from metadata after search
+#' @param n_site_country number of characters extraced from metadata after search
+#' @param n_dataset_project number of characters extraced from metadata after search
+#' @param n_lab number of characters extraced from metadata after search
+#' @param n_scales number of characters extraced from metadata after search
+#' @param n_site_elevation number of characters extraced from metadata after search
+#' @param n_altitude_comment number of characters extraced from metadata after search
+#' @param n_utc number of characters extraced from metadata after search
+#' @param fill_value fill value. Appeared in aoa_aircraft-flask_19_allvalid.txt
+#' @param as_list Logical to return as list
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table fwrite ".N" ":=" rbindlist "%chin%"
+#' @export
+#' @examples {
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs)
+#' dt <- obs_meta(index)
+#' }
+obs_meta <- function(index,
+                     verbose = TRUE,
+                     n_site_code = 15,
+                     n_site_name = 15,
+                     n_site_latitude = 18,
+                     n_site_longitude = 19,
+                     n_site_country = 18,
+                     n_dataset_project = 21,
+                     n_lab = 16,
+                     n_scales = 31,
+                     n_site_elevation = 20,
+                     n_altitude_comment = 22,
+                     n_utc = 18,
+                     fill_value = -1e+34,
+                     as_list = FALSE){
+
+  df <- index
+  if(nrow(df) == 0) stop("empty index")
+
+  x1 <- df$id
+
+  lapply(seq_along(x1), function(i) {
+
+    agl <- df$agl[i]
+    n <- df$n
+    categories <- df$sector[i]
+
+    if(verbose) cat(paste0(i, ": ", df$name[i], "\n"))
+
+    nr <- data.table::fread(x1[i], nrows = 1)$V4
+
+    att <- readLines(x1[i], n = nr)
+
+    # site code ####
+    pattern <- grep(pattern = "site_code",
+                    x = att,
+                    value = T)
+
+    # everything after char 15
[…]

### 2023-04-20 R/08_plot.R

@@ -1,0 +1,139 @@
+#' Read obspack metadata
+#'
+#' @param dt data.table
+#' @param time x axis column (time)
+#' @param y y axis column, default "value"
+#' @param yfactor factor of y
+#' @param colu column to plot by color, default site_code
+#' @param type type of plot, default "p"
+#' @param n Character indicating `colu` to subset, for instance,
+#' if you want to plot "site_code", here include all the
+#' site_code that you want to plot
+#' @param pal Color palette name see \link[cptcity]{cpt},
+#' default "cb_qual_Accent_08"
+#' @param xlab Character, xlab
+#' @param ylab Character, ylab
+#' @param verbose Logical to show more information
+#' @param ... plot arguments
+#' @param xlim x limits
+#' @param ylim y limits
+#' @return Plot
+#' @importFrom cptcity cpt find_cpt
+#' @importFrom graphics par lines
+#' @importFrom data.table as.data.table
+#' @export
+#' @examples \dontrun{
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs)
+#' dt <- obs_read(index)
+#' obs_plot(dt, time = "time")
+#' }
+obs_plot <- function(dt,
+                     time,
+                     y = "value",
+                     yfactor = 1,
+                     colu = "site_code",
+                     type = "p",
+                     n = if(length(unique(dt[[colu]])) == 1) unique(dt[[colu]]) else unique(dt[[colu]])[1:2] ,
+                     pal = cptcity::find_cpt("qual")[6],
+                     verbose = TRUE,
+                     xlab = "time",
+                     ylab = "value",
+                     xlim = range(dt[[time]], na.rm = TRUE),
+                     ylim = range(dt[[y]], na.rm = TRUE),
+                     ...){
+
+  oldpar <- graphics::par(no.readonly = TRUE)       # code line i
+  on.exit(graphics::par(oldpar))                    # code line i + 1
+
+  dt <- data.table::as.data.table(dt)
+  dt[[y]] <- dt[[y]]*yfactor
+  value <- NULL
+  nmiss <- nrow(dt[is.na(value)])
+
+  if(nmiss > 0) if(verbose) cat("Found ", nmiss, "missing `value` on dt\n")
+
+  dt <- dt[!is.na(value)]
+
[…]

### 2023-04-20 R/09_obs_table.R

@@ -1,0 +1,361 @@
+#' Obspack Table
+#'
+#' This function reads the obsPack directory
+#' providing a summary for the columns: "value", "time",
+#' "time_decimal", "latitude" and "longitude". These summary
+#' are made by the columns "name", "sector", "site_name",
+#' "site_country", "type_altitude", "lab_1_abbr" and
+#' "site_utc2lst"
+#'
+#' @param index data.table
+#' @param categories character; ONE category
+#' : of c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask").
+#' @param verbose Logical to show more information
+#' @param n_site_code number of characters extraced from metadata after search
+#' @param n_site_name number of characters extraced from metadata after search
+#' @param n_site_country number of characters extraced from metadata after search
+#' @param n_dataset_project number of characters extraced from metadata after search
+#' @param n_lab number of characters extraced from metadata after search
+#' @param n_scales number of characters extraced from metadata after search
+#' @param n_site_elevation number of characters extraced from metadata after search
+#' @param n_altitude_comment number of characters extraced from metadata after search
+#' @param n_utc number of characters extraced from metadata after search
+#' @param fill_value fill value. Appeared in aoa_aircraft-flask_19_allvalid.txt
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table fwrite ".N" ":=" rbindlist "%chin%"
+#' @export
+#' @note Categories are defined in `obs_summary`. If aircraft* is used,
+#' then `obs_table` will calculate stats for altitude_final.
+#' @examples {
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs, "aircraft-flask")
+#' dx <- obs_table(index, categories = "aircraft-flask")
+#' }
+obs_table <- function(index,
+                     categories = "tower-insitu",
+                     verbose = TRUE,
+                     n_site_code = 15,
+                     n_site_name = 15,
+                     n_site_country = 18,
+                     n_dataset_project = 21,
+                     n_lab = 16,
+                     n_scales = 31,
+                     n_site_elevation = 20,
+                     n_altitude_comment = 22,
+                     n_utc = 18,
+                     fill_value = -1e+34){
+
+
+  # categories = "tower-insitu"
[…]

### 2023-04-20 R/class.R

@@ -1,0 +1,147 @@
+#' Methods for objects with class `invfile`
+#'
+#'
+#' @return Objects of class "invfile"
+#'
+#' @param x Object with class "invfile"
+#' @param ... plot arguments
+#' @aliases invfile
+#' @examples \dontrun{
+#' #do not run
+#' }
+
+
+#' @rdname invfile
+#' @method print invfile
+#' @export
+print.invfile <- function(x, ...) {
+  cat("footprints_hera_hysplit: ",
+      length(x$footprints_hera_hysplit),"Observations\n:")
+  print(x$footprints_hera_hysplit[1:5])
+  print("...")
+
+
+  cat("obs_hysplit:",
+      length(x$obs_hysplit),"Observations\n:")
+  print(x$obs_hysplit[1:5])
+  print("...")
+
+  cat("receptor_info_hysplit:",
+      nrow(x$receptor_info_hysplit),"Observations\n:")
+  print(str(x$receptor_info_hysplit))
+
+}
+
+
+
+#' @rdname invfile
+#' @param object invfile
+#' @method summary invfile
+#' @export
+summary.invfile <- function(object, ...) {
+
+  cat("footprints_hera_hysplit\n:")
+  cat(length(object[[1]]), "files\n:")
+  cat(object[[1]][1], "\n...\n", object[[1]][length(object[[1]])], "\n")
+
+
+  cat("obs_hysplit\n:")
+  print(summary(object[[2]]))
+
+  cat("receptor_info_hysplit\n:")
+  print(lapply(object[[3]],  summary))
+
+}
+
+#' @rdname invfile
+#' @param time x axis column (time)
+#' @method plot invfile
+#' @importFrom cptcity cpt find_cpt
+#' @importFrom graphics par lines
+#' @export
+plot.invfile <- function(x,
+                         time = "timeUTC",
+                         ...) {
+
+  # oldpar <- graphics::par(no.readonly = TRUE)       # code line i
+  # on.exit(graphics::par(oldpar))                    # code line i + 1
+
+  master <- x$master
+
+  if(is.numeric(x$obs_hysplit)) {
+    value = x$obs_hysplit
+  } else {
+    value = x$obs_hysplit[[1]]
+  }
+
+  obs_plot(dt = master,
+           time = time,
+           y = value,
+           ...)
+
+  # value <- NULL
+  # nmiss <- nrow(master[is.na(value)])
+  #
[…]

### 2023-04-20 R/helpers.R

@@ -1,0 +1,525 @@
+#' Helpers
+#'
+#'
+#' @title outersect
+#' @family helpers
+#' @name obs_out
+#' @description Just the opposite of intersect
+#' @param x vector
+#' @param y vector
+#' @return vector opposite of intersect
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' }
+#' }
+obs_out <- function(x, y) {
+  sort(c(setdiff(x, y),
+         setdiff(y, x)))
+}
+
+
+#' @title list.dt
+#' @family helpers
+#' @name obs_list.dt
+#' @description Some treatments for list of data.frames
+#' @param ldf list of data.frames
+#' @param na common names in the final data.frame
+#' @param verbose Logical to show more information
+#' @return long data.table
+#' @note 1. Filter out empty data.frames
+#' 2. identify common names
+#' 3. rbindlist and return data.table
+#' @importFrom data.table rbindlist
+#' @importFrom utils object.size
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' }
+#' }
+obs_list.dt <- function(ldf, na, verbose = TRUE) {
+
+  ldf <- Filter(function(x) dim(x)[1] > 0, ldf)
+
+  lm <- lapply(ldf, names)
+
+  na <- Reduce(intersect, lapply(ldf, names))
+
+
+  lxx <- function(x, names) {
+    lapply(seq_along(x), function(i) {
+      x[[i]][, names, with = FALSE]
+    }) -> lx
+    data.table::rbindlist(lx)
+  }
+  dt <- lxx(ldf, na)
+  if(verbose) print(format(utils::object.size(dt), units = "Mb"))
+  return(dt)
+}
+
+
+#' @title Number to character
+#' @family helpers
+#' @description transform integer to character and add zero at left
+#' @param x integer
+#' @param char Character to be added ad the beginning of x
+#' @return transform integer in character
+#' @name robspack-deprecated
+#' @seealso \code{\link{robspack-deprecated}}
+#' @keywords internal
+NULL
+# obs_addzero <- function(x, char) {
+#
+#   # if(!inherits(x, "integer")) stop("x must be integer")
+#
+#   dd <- as.character(x)
+#   nch <- nchar(dd)
+#   nchmax <- max(nchar(dd))
+#
+#   difch <- nchmax - nch
+#
+#   if(length(unique(nch)) == 1 ) {
+#     if (unique(nch) == 1) {
[…]

### 2023-04-20 R/robspack-deprecated.R

@@ -1,0 +1,9 @@
+## robspack-deprecated.r
+#' @title Deprecated functions in package \pkg{robspack}.
+#' @description The functions listed below are deprecated and will be defunct in
+#'   the near future. When possible, alternative functions with similar
+#'   functionality are also mentioned. Help pages for deprecated functions are
+#'   available at \code{help("-deprecated")}.
+#' @name robspack-deprecated
+#' @keywords internal
+NULL

### 2023-04-20 rscripts/00_README.R

@@ -1,0 +1,38 @@
+
+# README ####
+# The directory `rscripts` contains the required files
+# to generate the list of receptors using `robspack`
+
+# The scripts will run from a defined working directory
+
+
+# 1. Run index (summary)
+
+# The script r/index.R will create the tree directory
+# to store the processed ObsPack files
+# Also, it will create summaries and store them
+# for each category in the directory `obs`
+
+# 2. Categories in the directory r
+# These scripts need to be run using Rscript command
+#examples:
+# Rscript r/aircore_year.R 2020 obs/obs_aircore.csv.gz
+# Rscript r/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv.gz
+# Rscript r/flask_year.R 2020 obs/obs_flask.csv.gz
+# Rscript r/surface_year.R 2020 obs/obs_surface-insitu.csv.gz
+# Rscript r/tower_year.R 2020 obs/obs_tower-insitu.csv.gz
+
+# Windows users can adapt the following lines
+# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/tower_insitu_year.R 2020 obs/obs_tower-insitu.csv.gz
+# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/aircore_year.R 2020 obs/obs_aircore.csv.gz
+
+
+# 3. Run Hysplit
+# Once if we have the receptor, we can
+# run Hysplit to generate footprints in NetCDF format
+
+# 4. Generate receptor info
+# Once we have the receptor list and the footprints were generated,
+# we can generate the receptor info, which
+# consists in text files.
+# The script is r/inputs_inv.R

### 2023-04-20 rscripts/r/aircore_year.R

@@ -1,0 +1,173 @@
+#example Rscript rscripts/aircore_year.R 2020 obs/obs_aircore.csv
+
+#!/usr/bin/env Rscript
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) != 2) {
+  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
+} else {
+  yy <-  as.numeric(args[1] )
+  print(yy)
+
+  ind  <- as.character(args[2])
+  print(ind)
+
+}
+
+out <- paste0("master/master_aircore", yy)
+
+message("All missing values are filled with 999999")
+
+cat("                 _     robspackfilter              \n")
+cat("               (`  ).                  _           \n")
+cat("              (     ).             .:(`  )`.       \n")
+cat(")           _(       '`.          :(   .    )      \n")
+cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
+cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
+cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
+cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
+cat(")  )  ( )       --'       `- __.'         :(      )) \n")
+cat(".-'  (_.'          .')                    `(    )  ))\n")
+cat("                  (_  )                     ` __.:'          \n")
+cat("\n")
+cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
+
+library(data.table)
+library(robspack)
+
+# 0 filters ####
+north <- 80
+south <- 10
+west <- -170
+east <- -50
+max_altitude <- 8000
+
+# 1 Index ####
+#cat("Running obs_index(obs)\n")
+#index <- obs_index(obs = obs)
+
+# 2 Read ####
+#cat("Running obs_read(index = index,
+#               categories = 'aircore',
+#               verbose = T)\n")
+#df <- obs_read(index = index,
+#               categories = "aircore",
+#               verbose = T)
+
+df <- fread(ind)
+
+# 3 Filtering space and time ###
+cat("Filtering\n")
+df <- rbind(df[year == yy - 1 & month == 12],
+            df[year == yy],
+            df[year == yy + 1 & month == 1])
+
+df <- df[altitude_final < max_altitude &
+           latitude < north &
+           latitude > south &
+           longitude < east &
[…]

### 2023-04-20 rscripts/r/aircraft_year.R

@@ -1,0 +1,207 @@
+#example Rscript rscripts/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv
+
+#!/usr/bin/env Rscript
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) != 2) {
+  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
+} else {
+  yy <-  as.numeric(args[1] )
+  print(yy)
+
+  ind  <- as.character(args[2])
+  print(ind)
+
+}
+
+out <- paste0("master/master_aircraft_insitu_", yy)
+
+message("All missing values are filled with 999999")
+
+cat("                 _     robspackfilter              \n")
+cat("               (`  ).                  _           \n")
+cat("              (     ).             .:(`  )`.       \n")
+cat(")           _(       '`.          :(   .    )      \n")
+cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
+cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
+cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
+cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
+cat(")  )  ( )       --'       `- __.'         :(      )) \n")
+cat(".-'  (_.'          .')                    `(    )  ))\n")
+cat("                  (_  )                     ` __.:'          \n")
+cat("\n")
+cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
+
+library(data.table)
+library(robspack)
+
+# 0 filters ####
+north <- 80
+south <- 10
+west <- -170
+east <- -50
+max_altitude <- 8000
+
+# evening <- 14 # for aircraft keep all the local hours
+
+# 1 Index ####
+#cat("Running obs_index(obs)\n")
+#index <- obs_index(obs = obs)
+
+# 2 Read ####
+#cat("Running obs_read(index = index,
+#               categories = 'aircraft-insitu',
+#               verbose = T)\n")
+#df <- obs_read(index = index,
+#               categories = "aircraft-insitu",
+#               verbose = T)
+df  <- fread(ind)
+
+# 3 ECO 2020 ####
+(cat("Reading ECO 2020\n"))
+
+if(Sys.info()[["sysname"]] == "Windows") {
[…]

### 2023-04-20 rscripts/r/flask_non_noaa_year.R

@@ -1,0 +1,205 @@
+#example Rscript rscripts/flask_year.R 2020 obs/obs_flask.csv
+
+#!/usr/bin/env Rscript
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) != 2) {
+  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
+} else {
+  yy <-  as.numeric(args[1] )
+  print(yy)
+
+  ind  <- as.character(args[2])
+  print(ind)
+
+}
+out <- paste0("master/master_flask_no_noaa_", yy)
+
+
+
+message("All missing values are filled with 999999")
+
+cat("                 _     robspackfilter              \n")
+cat("               (`  ).                  _           \n")
+cat("              (     ).             .:(`  )`.       \n")
+cat(")           _(       '`.          :(   .    )      \n")
+cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
+cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
+cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
+cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
+cat(")  )  ( )       --'       `- __.'         :(      )) \n")
+cat(".-'  (_.'          .')                    `(    )  ))\n")
+cat("                  (_  )                     ` __.:'          \n")
+cat("\n")
+cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
+
+library(data.table)
+library(robspack)
+
+# 0 filters ####
+north <- 80
+south <- 10
+west <- -170
+east <- -50
+max_altitude <- 8000
+
+# 1 Index ####
+#cat("Running obs_index(obs)\n")
+#index <- obs_index(obs = obs)
+
+# 2 Read ####
+#cat("Running obs_read(index = index,
+#               categories = 'flask',
+#               verbose = T)\n")
+#df <- obs_read(index = index,
+#               categories = "flask",
+#               verbose = T)
+
+df <- fread(ind)
+
+# 3 Filtering space and time ###
+cat("Filtering\n")
+df <- rbind(df[year == yy - 1 & month == 12],
+            df[year == yy],
+            df[year == yy + 1 & month == 1])
+
+df <- df[altitude_final < max_altitude &
+           latitude < north &
+           latitude > south &
+           longitude < east &
+           longitude > west]
+
[…]

### 2023-04-20 rscripts/r/index.R

@@ -1,0 +1,54 @@
+wd <- paste0(getwd(), "/rscripts") #  you can change
+# if you change, define a new path
+# wd <- "path"
+# then set working directory
+setwd(wd)
+
+cat("This is your working directory", getwd(), "\n")
+
+library(data.table)
+library(robspack)
+
+categories <- c(
+  #"aircraft-pfp",
+  "aircraft-insitu",
+  "surface-insitu",
+  "aircore",
+  #"surface-pfp",
+  "tower-insitu",
+  "shipboard-insitu",
+  "flask"
+)
+
+# ObsPack ####
+obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt/"
+
+# Create Directories ####
+# master have files with all columns
+dir.create("master", showWarnings = FALSE)
+
+# receptors have files with required columns
+dir.create("receptors", showWarnings = FALSE)
+
+# receptors have files with required columns
+dir.create("obs", showWarnings = FALSE)
+
+# 1 summary (previously named index) ####
+cat("Running obs_summary(obs)\n")
+index <- obs_summary(obs = obs)
+
+
+# 2 read ####
+
+for(i in seq_along(categories)) {
+
+cat(paste0("Running obs_read(index = index,
+               categories = ", categories[i], "\n"))
+
+df <- obs_read(index = index,
+               categories = categories[i],
+               verbose = T)
+
+fwrite(df, paste0("obs/obs_",categories[i], ".csv.gz"))
+
+}

### 2023-04-20 rscripts/r/inputs_inv.R

@@ -1,0 +1,310 @@
+
+library(colorout)
+library(data.table)
+library(robspack)
+
+dir.create("invfiles")
+
+# aircraft insitu ####
+ma1 <- paste0("/path/master/master_aircraft_insitu_", 2017:2021, ".csv")
+na <- gsub(pattern = ".csv", replacement = "", x = ma1)
+na <- gsub(pattern = "master/master_", replacement = "", x = na)
+
+lx <- list()
+
+for(i in seq_along(ma1)) {
+  if(!file.exists(ma1[i])) {
+    cat("skipping ", ma1[i], "\n")
+    next
+  }
+  dx <- fread(ma1[i])
+  dx$file <- na[i]
+  print(dim(dx))
+  lx[[i]] <- dx
+}
+master <- obs_list.dt(lx)
+# filtering between dec 2017 and jan 2021
+
+master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
+                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
+
+master$nc <- obs_footname(year = master$year,
+                          month = master$month,
+                          day = master$day,
+                          hour = master$hour,
+                          minute = master$minute,
+                          lat = master$latitude,
+                          lon = master$longitude,
+                          alt = master$altitude_final)
+
+path <-  "/path/aircraft/flask/" # param
+
+master$fnc <- paste0(path, master$nc)
+
+master$is_fnc <- file.exists(master$fnc)
+
+table(master$is_fnc)
+
+master <- master[is_fnc == TRUE]
+
+master <- master[!is.na(value)]
+
+master <- master[altitude_final > 10]
+
+dfaircraft <- obs_invfiles(master = master,
+                           path = path,
+                           Type = "continuous",
+                           SubType = "aircraft")
+obs_plot(dfaircraft$master,
+         time = "timeUTC",
+         y = "value",
+         colu = "site_code",
+         yfactor = 1e9,
+         cex = 0.5,
+         type = "l"
+)
+
+
+# tower insitu ####
+ma1 <- paste0("master/master_tower_insitu_", 2017:2021, ".csv")
+na <- gsub(pattern = ".csv", replacement = "", x = ma1)
+na <- gsub(pattern = "master/master_", replacement = "", x = na)
+
+lx <- list()
+
+for(i in seq_along(ma1)) {
+  if(!file.exists(ma1[i])) {
+    cat("skipping ", ma1[i], "\n")
+    next
+  }
[…]

### 2023-04-20 rscripts/r/inputs_invs.R

@@ -1,0 +1,314 @@
+
+library(colorout)
+library(data.table)
+library(robspackfilter)
+
+# dir.create("invfiles")
+
+# aircraft insitu ####
+ma1 <- paste0("master/master_aircraft_insitu_", 2017:2021, ".csv")
+na <- gsub(pattern = ".csv", replacement = "", x = ma1)
+na <- gsub(pattern = "master/master_", replacement = "", x = na)
+
+lx <- list()
+
+for(i in seq_along(ma1)) {
+  if(!file.exists(ma1[i])) {
+    cat("skiiping ", ma1[i], "\n")
+    next
+  }
+  dx <- fread(ma1[i])
+  dx$file <- na[i]
+  print(dim(dx))
+  lx[[i]] <- dx
+}
+master <- obs_list.dt(lx)
+# filtering between dec 2017 and jan 2021
+
+master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
+                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
+#  footprints_hera_hysplit.txt
+master$nc <- obs_footname(year = master$year,
+                          month = master$month,
+                          day = master$day,
+                          hour = master$hour,
+                          minute = master$minute,
+                          lat = master$latitude,
+                          lon = master$longitude,
+                          alt = master$altitude_final)
+
+path <-  "/work/noaa/co2/footprint/hysplit/CPO2021/hysplit4/aircraft/flask/" # param
+
+master$fnc <- paste0(path, master$nc)
+
+master$is_fnc <- file.exists(master$fnc)
+
+table(master$is_fnc)
+
+master <- master[is_fnc == TRUE]
+
+master <- master[!is.na(value)]
+
+master <- master[altitude_final > 10]
+
+dfaircraft <- obs_invfiles(master = master,
+                           path = path,
+                           Type = "continuous",
+                           SubType = "aircraft")
+plot(dfaircraft)
+
+obs_plot(dfsurface$master,
+         time = "timeUTC",
+         y = "value",
+         colu = "site_code",
+         n = c("BRA", "EGB", "EST", "ETL", "LLB"),
+         yfactor = 1e9,
+         cex = 0.5,
+         type = "l",
+         ylim = c(1800, 3000)
+)
+
+
+# Found  760 missing `value` on master
+# It is possible that this script must be run again to explude missing value
+
+
[…]

### 2023-04-20 rscripts/r/tower_insitu_year.R

@@ -1,0 +1,238 @@
+#example Rscript rscripts/tower_year.R 2020 obs/obs_tower-insitu.csv
+
+#!/usr/bin/env Rscript
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) != 2) {
+  stop("At least one argument must be supplied (year).n and obs_index",
+       call.=FALSE)
+} else {
+  yy <-  as.numeric(args[1] )
+  print(yy)
+
+  ind  <- as.character(args[2])
+  print(ind)
+
+}
+out <- paste0("master/master_tower_insitu_", yy)
+
+message("All missing values are filled with 999999")
+
+n01 <- ("                  _     robspackfilter              \n")
+n02 <- ("               (`  ).                  _           \n")
+n03 <- ("              (     ).             .:(`  )`.       \n")
+n04 <- (")           _(       '`.          :(   .    )      \n")
+n05 <- ("        .=(`(      .   )     .--  `.  (    ) )      \n")
+n06 <- ("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
+n07 <- ("`.     `(       ) )       (   .  )     (   )  ._   \n")
+n08 <- ("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
+n09 <- (")  )  ( )       --'       `- __.'         :(      )) \n")
+n10 <- (".-'  (_.'          .')                    `(    )  ))\n")
+n11 <- ("                  (_  )                     ` __.:'          \n")
+n12 <- ("\n")
+n13 <- ("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
+cat(c(n01, n02,n03, n04, n05,
+      n06, n07, n08, n09, n10,
+      n11, n12, n13))
+
+library(data.table)
+library(robspack)
+
+# 0 filters ####
+north <- 80
+south <- 10
+west <- -170
+east <- -50
+max_altitude <- 8000
+evening <- 14:15
+
+# 1 Index ####
+#cat("Running obs_index(obs)\n")
+#index <- obs_index(obs = obs)
+
+# 2 Read ####
+#cat("Running obs_read(index = index,
+#               categories = 'tower-insitu',
+#               verbose = T)\n")
+# df <- obs_read(index = index,
+#               categories = "tower-insitu",
+#               verbose = T)
+df <- fread(ind)
+
+# 3 Filtering space and time ####
+cat("Filterinig\n")
[…]

### 2023-05-08 paper/paper.Rmd
paper
readme

@@
 | `obs_addtime()`       | Add UTC time                                      |
-| `obs_addzero()`       | Format numbers to match specific requirements     |
 | `obs_agg()`           | Aggregates ObsPack by time                        |

### 2023-05-08 R/09_obs_table.R

@@ -35,18 +35,18 @@
 #' dx <- obs_table(index, categories = "aircraft-flask")
 #' }
 obs_table <- function(index,
-                     categories = "tower-insitu",
-                     verbose = TRUE,
-                     n_site_code = 15,
-                     n_site_name = 15,
-                     n_site_country = 18,
-                     n_dataset_project = 21,
-                     n_lab = 16,
-                     n_scales = 31,
-                     n_site_elevation = 20,
-                     n_altitude_comment = 22,
-                     n_utc = 18,
-                     fill_value = -1e+34){
+                      categories = "tower-insitu",
+                      verbose = TRUE,
+                      n_site_code = 15,
+                      n_site_name = 15,
+                      n_site_country = 18,
+                      n_dataset_project = 21,
+                      n_lab = 16,
+                      n_scales = 31,
+                      n_site_elevation = 20,
+                      n_altitude_comment = 22,
+                      n_utc = 18,
+                      fill_value = -1e+34){
 
 
   # categories = "tower-insitu"
@@ -262,9 +262,11 @@ obs_table <- function(index,
                                "not available",
                                dt$type_altitude)
     dt$id <- i
-.SD <- . <- name <- altitude_final <- NULL
+    .SD <- . <- name <- altitude_final <- NULL
+
     if(length(grep("aircraft", categories)) > 0) {
-      dt[, lapply(.SD, summary, na.rm = TRUE),
+      dt[,
+         lapply(.SD, summary, na.rm = TRUE),
          by = .(name,
                 sector,
                 site_name,
@@ -351,11 +353,11 @@ obs_table <- function(index,
 
   type_altitude <- NULL
   .N <- NULL
-    dt <- data.table::rbindlist(lx, use.names = T)
-    # if(verbose) cat("Adding id\n")
-    # dt$id <- 1:nrow(dt)
-    # print(dt[, .N, by = type_altitude])
-    return(dt)
+  dt <- data.table::rbindlist(lx, use.names = T)
+  # if(verbose) cat("Adding id\n")
+  # dt$id <- 1:nrow(dt)
+  # print(dt[, .N, by = type_altitude])
+  return(dt)
 
 
 }

### 2023-06-12 rscripts/00_README.Rmd
paper

---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### files:
```{r, echo=F}
list.files(recursive = T)
```

The directory `rscripts` contains the required files
to generate the list of receptors using `robspack`
The scripts will run from a defined working directory

### 1. Run index (summary)

The script `r/index.R` will create the tree directory
to store the processed ObsPack files
Also, it will create summaries and store them
for each category in the directory `obs`

### 2. Categories in the directory r
These scripts need to be run using Rscript command
examples:

- `Rscript r/aircore_year.R 2020 obs/obs_aircore.csv.gz`
- `Rscript r/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv.gz`
- `Rscript r/flask_year.R 2020 obs/obs_flask.csv.gz`
- `Rscript r/surface_year.R 2020 obs/obs_surface-insitu.csv.gz`
- `Rscript r/tower_year.R 2020 obs/obs_tower-insitu.csv.gz`

Windows users can adapt the following lines

- "C:/Users/*User*/My Documents/R/R-4.1.3/bin/Rscript.exe" r/tower_insitu_year.R 2020 obs/obs_tower-insitu.csv.gz
- "C:/Users/*User*/My Documents/R/R-4.1.3/bin/Rscript.exe" r/aircore_year.R 2020 obs/obs_aircore.csv.gz

### 3. Run Hysplit
Once if we have the receptor, we can
run Hysplit to generate footprints in NetCDF format

### 4. Generate receptor info
Once we have the receptor list and the footprints were generated,
we can generate the receptor info, which
consists in text files.
The script is `r/inputs_inv.R`

### 2023-07-12 paper_elsevier/paper_elsevier.Rmd
add paper

---
title: robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
author:
  - name: Sergio Ibarra-Espinosa
    email: sergio.ibarra-espinosa@noaa.gov
    affiliation: a,b
    correspondingauthor: true
    footnote: 1
  - name: Lei Hu
    email: lhu@noaa.gov
    affiliation: b
address:
  - code: a
    organization: Cooperative Institute for Research in Environmental Sciences, CU Boulder
    addressline: 216 UCB, University of Colorado Boulder campus
    city: Boulder
    state: Colorado
    postcode: 80309
    country: United States
  - code: b
    organization: NOAA Global Monitoring Laboratory
    addressline: 325 Broadway
    state: Colorado
    postcode: 80309
    country: United States
footnote:
  - code: 1
    text: "This is the first author footnote."
abstract: |
  In this study, we present a new open-source R package `robspack`, 
  to read, process, select, and plot NOAA Observation Package (ObsPack)
  data products. We use a methane ObsPack data product as an example
  in this code base, but it can be easily modified to analyze ObsPack
  products for other greenhouse gasses. The R package starts with 
  creating a catalog of  all ObsPack files in each product. It then 
  reads all files and creates one database. While reading each ObsPack 
  file, it extracts site elevation and time zone information from the 
  file header and calculates sampling altitude in meters above ground 
  level and local time for individual samping events. Finally, it 
  processes and selects observations for inverse modeling purposes. 
  This package imports functions from data.table R package, which 
  contains C bindings with parallel implementation via Open-MP 
  [@dt]. data.table is faster than other 
  Python, Julia and R implementations for data-science, 
  providing a strong basis for robspack. robspack provides functions 
  to perform these tasks in a transparent and efficient way, 
  supporting open-source communities in environmental sciences.
keywords: 
  - ObsPack
  - NOAA
  - Greenhouse gases
journal: "Environmental Modelling & Software"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

<!-- output:  -->
<!--   rticles::elsevier_article: -->
<!--     keep_tex: true -->
<!--     citation_package: natbib -->

<!-- output: word_document -->
# Introduction

<!-- New Introduction -->
<!-- 1. Global Warming - Paris Accord - Methane Pledge -> obs -->
The world is experiencing an accelerated global warming
due to the accumulation of greenhouse gases (GHG)
since the industrial revolution [@us2018].
Greenhouse gas observations are critical to monitor the state of the
atmosphere, quantify present and historical emissions, and understand 
global climate change. 
During the 21th Conference of Parties (COP21), it was established the Paris 
Accord, a multilateral effort reduce greenhouse emissions
in order to limit the temperature increment of 1.5 degrees
[@rhodes20162015]. 
Methane is a greenhouse gas responsible for half of the 
temperature increase since preindustrial levels.
Furthermore, methane has a 9 years lifetime and a
global warming potential of 30 over 100 years [@epagwp], with
a current global radiative forcing of 0.650 $Wm^{-2}$ [@aggi].
Hence, in the 26 version of COP conference 
[@hunter2021glasgow], it was 
signed the Global Methane Pledge aiming reduce at least 
methane emissions 30\% from 2020 levels by 2030, with
U.S. as one of the parties [@wh]. Therefore, monitoring $CH_4$ observations,
emissions and sinks has become critical.

<!-- 2. Importance of greenhouse gases observations -->
<!-- 3. NOAA - mandate by congress - Obspack -->

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire,
evaluate and make available long-term records of atmospheric gases[^2].
To achieve that goal, GML gather own and other laboratories
data, releasing observation in a compendium named ObsPack
[@masarie2014obspack]. Specifically, the $CH_4$ ObsPack GLOBALVIEW+
is a comprehensive product consisting in observations from aircrafts, 
ships, surface stations, towers and aircores.
However, each 
ObsPack product generally contains hundreds of files, each of which 
has different sampling frequencies, hours, and hundreds of lines of 
headers. It takes time and effort to develop tools to read and process 
each ObsPack product and select observations of interest for specific 
modeling and data analysis purposes.

NOAA ObsPack data has been used to support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases Observing 
SATellite (GOSAT) data. 
@saunois2016global 
At regional scale, @lu2022methane performed another studied focused
on north america using as priors local emissions inventories. 
Furthermore, @hu2023trend presented trends and cycles of methane over the 
US.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrfats is different than surface stations, but all the files
include concentrations and other critical fields. 
An important information is the scale of measurement, which isXX.
Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `robspack` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `robspack`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focued on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output and plots.

<!-- 4. Objective -->

# Installation

To install `robspack`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment.

```{r install, eval = F}
remotes::install_github("ibarraespinosa/robspack")
library(robspack)
```

# Overview

`robspack` is a collection of function organized together to read and 
process ObsPack files [@masarie2014obspack]. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. In other to facilitate this task,
we are letting to the public the directory shown below. This directory 
includes a README file with all the required information and scripts 
generate process each category. The structure of the directory is:

https://github.com/ibarraespinosa/robspack/tree/main/rscripts

```{bash, eval = F}
|--"README.md"
|-- r
    |-- index.R
    |-- aircore_year.R 
    |-- aircraft_year.R
    |-- flask_non_noaa_year.R
    |-- surface_insitu_year.R
    |-- tower_insitu_year.R
    |-- inputs_inv.R
```

```{r flow, echo = FALSE, message=FALSE, fig.cap="Methane observations (ppb) by towers-insitu", out.height="50%", out.width="100%"}
knitr::include_graphics("flow/Slide1.PNG")
```

The file `index.R` creates a summary of ObsPack, generates the directories
master, receptor and obs, and store the summaries in obs directory for each
category. Then, the scripts for each category are run to generate the master
and receptor files. The next step consists in running HYSPLIT and obtaining
footprints (not in the scope of this manuscript). At last, the script
`inputs_inv.R` checks each footprint file for each category and generates the
final receptor list files. The process is described in detail in the following
example for tower insitu.

The first step consists in constructing a summary for the ObsPack. 
This is required to read the data, but also, identify agl, which is
present in some of the file names. This function returns a data.frame.
Optionally, the user can indicate a path to store the data.frame. 
obs_summary also prints a summary of the data. The second argument 
is the categories, and by default includes the categories shown below, 
to account for all the files. Then the summary data.frame contains the 
columns id as the full path to each file, name which is the name or 
relative path of the file, n which is just an id, sector such as tower,
and the column agl which indicates the agl indicated in the name of the
file if available. To read the documentation of this function, the 
user must run `?obs_summary`.

```{r libs, include = F, message=F, warning=F}
library(robspack)
library(data.table)
```

```{r}

categories <- c("aircraft-pfp","aircraft-insitu","surface-insitu",
  "aircore","surface-pfp","tower-insitu","shipboard-insitu","flask")
obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs, verbose = F)
```

```{r kable, echo = FALSE}
ta <- obs_summary(obs = obs, verbose = F, aslist = TRUE)$summary
knitr::kable(ta, format = "latex", caption = "Summary of ObsPack")
```

There are 362 files in the ObsPack directory. The printed information also 
shows the total at the bottom, as the sum of the individual file by sector. 
This is to ensure that the sum of files is equal to the total number of files 
found, shown at the top. furthermore, the printed information also shows that 
there are 136 files with the agl explicitly mentioned in the name of the file.
Sometimes we need more information about the site. For instance, what do the 
observations start and end. Then, we added the function `obs_table`, which 
calculates statistics summary of “time” and other numeric variables by file 
name, sector, site, altitude and mode. For instance, the observations in the 
site "SCT" in South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and 
"2020-12-31 23:30:00 UTC". In figure \@ref(fig:tiobs) we see the average of 
methane concentrations over each tower-insitu site. Higher concentration
are found over Russia. Over the United States (U.S.), most of
towers are found in the south to east coast.

```{r sf, echo = FALSE, message=FALSE}
library(sf)
library(ggplot2)
dft <- obs_table(index =  index, categories = "tower-insitu")
dft <- dft[stat == "mean"]
sdft <- st_as_sf(dft, coords = c("longitude", "latitude"), crs = 4326)
sdft$value <- sdft$value*1e09

```

```{r tiobs, echo = FALSE, message=FALSE, fig.cap="Methane observations (ppb) by towers-insitu"}
plot(sdft["value"], reset = FALSE, axes = TRUE, graticule = TRUE,
     pch = 16, cex = 1.5, pal = cptcity::cpt(rev = T, colorRampPalette = T, 
                                             alpha = 0.5), main = NULL)
maps::map(add = T)
```

## Functions

The `robspack` functions are shown in table 1.

Table 1. Functions and classes in `robspack`.

| Function              | Description                                       |
|-----------------------|---------------------------------------------------|
| `invfile`             | Class with `print`, `summary` and `plot` methods  |
| `obs_addltime()`      | Add local time based on metadata and longitude    |
| `obs_addtime()`       | Add UTC time                                      |
| `obs_agg()`           | Aggregates ObsPack by time                        |
| `obs_find_receptors()`| Find expected receptors and NetCDF files          |
| `obs_format()`        | Format for some columns of data.table             |
| `obs_freq()`          | Return numeric vector in intervals                |
| `obs_invfiles()`      | Construct `invfile` objects                       |
| `obs_list.dt()`       | Rbind list of data.frames with different names    |
| `obs_meta()`          | Reads ObsPack metadata                            |
| `obs_out()`           | Outersect, opposed as intersect                   |
| `obs_rbind()`         | Rbind data.frames with different names            |
| `obs_read()`          | Read files, and add metadata as columns           |
| `obs_read_csvy()`     | Read csvy file and prints yaml header             |
| `obs_roundtime()`     | Round seconds from "POSIXct" "POSIXt" classes     |
| `obs_summary()`       | Construct summary of ObsPack as a data.frame      |
| `obs_table()`         | Return a data.frame with summary of data          |
| `obs_trunc()`         | Trunc numbers with a desired number of decimals   |
| `obs_write()`         | Write CSVY to disk, YAML followed by tabulated    |

# Application for towers in situObsPack summary

## Read data

Once the summary is built, the function obs_read will read the files available in the index file previously generated. Here we selected the category "tower-insitu". The argument verbose prints which files are being read each time, by default. At the end, this function prints the total number of observations by type of altitude (agl or asl).

```{r read}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

```

We added a function to plot the data read from ObsPack. The y-axis is
the field value and the x-axis is by default time. The data illustrated
sorted by color is the field site_code, with the default number of 3 
sites. The argument pal is to define the color palette, used by the 
internally imported function cptcity::cpt.

```{r obsplot, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
```

Before sub setting the data,  tower-insitu has about 
`r round(nrow(df)/1000000, 2)` million 
observations. These observations are made between 2004 and 2020. 
The identification of the altitude and type is critical, then we 
developed an approach based on the availability of data:

1. Identify agl from the name of the tile.
2. If agl not present, search fill_values used in elevation and 
transform them into NA (not available)
3. If agl is not present, agl = altitude - elevation.
4. If there are some NA in elevation, will result some NA in agl
5. A new column is added named altitude_final to store agl or asl
6. Another column named type_altitude is added to identify agl or asl.
7. If there is any case NA in altitude_final, type_altitude is 
“not available”

## Filtering

ObsPack includes global observations and sometimes we need to 
extract data for a specific region and periods of time. In this 
part we include spatial and temporal parameters to filter data. 
The year of interest is 2020, but we also included December of 
2019 and January of 2021. The spatial filtering is done by using 
the coordinates, in this case covering North America. After filtering
by space and time, we have `r round(nrow(df), 2)` million observations.

```{r filtering}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14:15

yy <- 2020
df <- rbind(df[year == yy - 1 & month == 12],
            df[year == yy],
            df[year == yy + 1 & month == 1])

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

## Time

The function obs_addtime adds time columns timeUTC, timeUTC_start 
which shows the start time of each observation and timeUTC_end which 
shows the end time for each observation. Then we need to identify the 
local time with the function add_ltime. This is important because to 
identify observations in the evening in local time for modeling 
purposes. add_ltime uses two methods, first identifying the time 
difference with utc by identifying the metadata column "site_utc2lst". 
If this information is not available, with the aircrafts for instance, 
the local time is calculated with an approximation based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where lt is the local time, UTC the time, and longitude is the 
coordinate. Then, the time is cut every two hours. We also identify 
the local time to select evening hours.

```{r time}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

Now there are `r nrow(df3)` observations. At this point we can calculate the 
averages of several columns by the cut time. The function obs_agg 
does this aggregation as shown in the following lines of code. The 
argument gby establish the function used to aggregate cols, in this 
case the function "mean" by time and altitude. Finally, we add 
local time again.

```{r agg}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value", "latitude", "longitude", "type_altitude",
                        "dif_time", "year_end", "site_utc2lst"),
               verbose = FALSE,
               byalt = TRUE)
df5 <- obs_addltime(df4)
```

Now there are `r nrow(df5)` observations. Towers can have observations at 
different heights. Here we need to select one site with the observations 
registered at the highest height. The column with the height is named 
altitude_final and the max altitude was named max_altitude. Then, 
we print the altitudes of each site.

```{r maxalt}
df5[,
    max_altitude := max(altitude_final),
    by = site_code]
df5[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()
```

## Saving master as text and csvy

Now that we have all the required information, we can save the files. 
Here, we name the data.frame as master, because it contains all the 
information. This is important because some fields can be used in the future, 
and for traceability. For convenience, time variables are transformed into 
character before writing into the disk. The separation is space ” “.

```{r fwrite, eval = T}
master <- df5
master$timeUTC <- as.character(master$timeUTC)
master$timeUTC_end <- as.character(master$timeUTC_end)
master$local_time <- as.character(master$local_time)

fwrite(master, 
       file = "tower_insitu_2020.txt",
       sep = " ")
```

The format Comma Separated Value with YAML (CSVY)[^1] consists in a 
typical CSV with a YAML header. The function obs_write includes the 
argument notes which allows adding custom notes at the header of the 
file. Below the notes, obs_write adds the output of the R function 
`str`, which provides a vertical summary of the data, known as structure.

```{r obs_write_csvy, eval = T}
obs_write_csvy(dt = master,
              notes = "tower 2020",
              out = "tower_insitu_2020.csvy")
```

To check the YAML header we read the first 38 lines of the files that 
were generated. Here we can see the column names, type of data and 
first observations. The YAML header is delimited by the characters 
"- - -" (not shown here).

```{r csvy, eval = F}
readLines("tower_insitu_2020.csvy")[1:38]
```

## Saving receptors

We need to filter some columns from the master files in a new object 
called receptors. This is needed because internally we run HYSPLIT 
[@hy] using the information from the receptors. In the case of a 
tower, we need to select observations with the highest altitude. The 
specific columns are selected as shown on the following code. We are 
selecting the ending times, because later HYSPLIT is run backwards 
based on the time of measurement, between ending and starting times. 
The columns about time are formatted to have two characters. For 
instance, the month 1, is formatted as “01”. We also need to filter 
for type_altitude equal 0, representing agl observations , or equal 
to 1, asl.

```{r receptors, eval = F}

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year", "month", "day",
                     "hour", "minute", "second",
                     "latitude", "longitude",
                     "altitude_final", "type_altitude",
                     "year_end", "month_end", "day_end", "hour_end",
                     "minute_end", "second_end")]
receptor$altitude_final <- round(receptor$altitude_final)
receptor <- obs_format(receptor)

if(nrow(receptor_agl) > 0) {
  fwrite(x = receptor_agl,
         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
  sep = " ")}

if(nrow(receptor_asl) > 0) {
  fwrite(x = receptor_asl,
         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
  sep = " ")}

```

## Recommendation for other applications

The approach to generate receptors depends on each type of observation
and other considerations. For instance, aircraft with continuous 
observations at each second can be filtered and averaged every 20 
seconds. In that way, the footprints are still representative and it 
would not be necessary to run HYSPLIT every second. Of course, it 
depends on the application and objective of the study. For this 
manuscript, we are presenting the generation of receptors based 
on tower observations.

# Conclusion
In this manuscript we presented an robskpack, an R package to read 
and process CH4 ObsPack GLOBALVIEW+ published by the Global 
Monitoring Laboratory (GML) from the National Oceanographic and 
Atmospheric Administration (NOAA). robspack reads the text data which 
have different headers and organizes them in a common format. Then, 
this software applies calculations to filter observations by time and 
space. Finally, this software generates receptors in a suitable format 
that allows it to run HYSPLIT and generate footprints. This software 
does not provide methods to run HYSPLIT, but the user can follow the 
site https://www.ready.noaa.gov/HYSPLIT.php.

[^1]: https://csvy.org/
[^2]: https://gml.noaa.gov/about/aboutgml.html

# Acknowledgements

Funding: This project is funded by the NOAA Climate Program Office 
AC4 and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was 
supported by the NOAA cooperative agreement NA22OAR4320151.

# References {-}

### 2023-07-12 paper_elsevier/paper_elsevier.tex
add paper

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Environmental Modelling \& Software} % Sets Journal name

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\begin{document}

\begin{frontmatter}

  \title{robspack An R package for processing atmospheric greenhouse gas
observations in NOAA's ObsPack}
    \author[a,b]{Sergio Ibarra-Espinosa%
  \corref{cor1}%
  \fnref{1}}
   \ead{sergio.ibarra-espinosa@noaa.gov} 
    \author[b]{Lei Hu%
  %
  }
   \ead{lhu@noaa.gov} 
      \affiliation[a]{
    organization={Cooperative Institute for Research in Environmental
Sciences, CU Boulder},addressline={216 UCB, University of Colorado
Boulder
campus},city={Boulder},postcode={80309},state={Colorado},country={United
States},}
    \affiliation[b]{
    organization={NOAA Global Monitoring Laboratory},addressline={325
Broadway},postcode={80309},state={Colorado},country={United States},}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{This is the first author footnote.}

  \begin{abstract}
  In this study, we present a new open-source R package
  \texttt{robspack}, to read, process, select, and plot NOAA Observation
  Package (ObsPack) data products. We use a methane ObsPack data product
  as an example in this code base, but it can be easily modified to
  analyze ObsPack products for other greenhouse gasses. The R package
  starts with creating a catalog of all ObsPack files in each product.
  It then reads all files and creates one database. While reading each
  ObsPack file, it extracts site elevation and time zone information
  from the file header and calculates sampling altitude in meters above
  ground level and local time for individual samping events. Finally, it
  processes and selects observations for inverse modeling purposes. This
  package imports functions from data.table R package, which contains C
  bindings with parallel implementation via Open-MP \citep{dt}.
  data.table is faster than other Python, Julia and R implementations
  for data-science, providing a strong basis for robspack. robspack
  provides functions to perform these tasks in a transparent and
  efficient way, supporting open-source communities in environmental
  sciences.
  \end{abstract}
    \begin{keyword}
    ObsPack \sep NOAA \sep 
    Greenhouse gases
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The world is experiencing an accelerated global warming due to the
accumulation of greenhouse gases (GHG) since the industrial revolution
\citep{us2018}. Greenhouse gas observations are critical to monitor the
state of the atmosphere, quantify present and historical emissions, and
understand global climate change. During the 21th Conference of Parties
(COP21), it was established the Paris Accord, a multilateral effort
reduce greenhouse emissions in order to limit the temperature increment
of 1.5 degrees \citep{rhodes20162015}. Methane is a greenhouse gas
responsible for half of the temperature increase since preindustrial
levels. Furthermore, methane has a 9 years lifetime and a global warming
potential of 30 over 100 years \citep{epagwp}, with a current global
radiative forcing of 0.650 \(Wm^{-2}\) \citep{aggi}. Hence, in the 26
version of COP conference \citep{hunter2021glasgow}, it was signed the
Global Methane Pledge aiming reduce at least methane emissions 30\% from
2020 levels by 2030, with U.S. as one of the parties \citep{wh}.
Therefore, monitoring \(CH_4\) observations, emissions and sinks has
become critical.

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire, evaluate
and make available long-term records of atmospheric gases\footnote{https://gml.noaa.gov/about/aboutgml.html}.
To achieve that goal, GML gather own and other laboratories data,
releasing observation in a compendium named ObsPack
\citep{masarie2014obspack}. Specifically, the \(CH_4\) ObsPack
GLOBALVIEW+ is a comprehensive product consisting in observations from
aircrafts, ships, surface stations, towers and aircores. However, each
ObsPack product generally contains hundreds of files, each of which has
different sampling frequencies, hours, and hundreds of lines of headers.
It takes time and effort to develop tools to read and process each
ObsPack product and select observations of interest for specific
modeling and data analysis purposes.

NOAA ObsPack data has been used to support many studies. For instance,
the global methane budget for the year 2017 was 596 \(Tgy^{-1}\), in
agreement with other studies
\citep{saunois2020global, saunois2016global} \citet{lu2021global},
characterized global methane emissions in between 2014 and 2017,
including a comparison with Greenhouse gases Observing SATellite (GOSAT)
data. \citet{saunois2016global} At regional scale, \citet{lu2022methane}
performed another studied focused on north america using as priors local
emissions inventories. Furthermore, \citet{hu2023trend} presented trends
and cycles of methane over the US.

The NOAA ObsPack data is delivered to the public as NetCDF and text
files. The structure of the files including descriptor fields depend on
the type of file. For instance, the metadata from aircrfats is different
than surface stations, but all the files include concentrations and
other critical fields. An important information is the scale of
measurement, which isXX. Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome. The \texttt{robspack} package provides the GHG
science and research community a transparent and efficient tool to
process ObsPack products for GHG modeling and analyses.

In this manuscript we present \texttt{robspack}, an R package to read,
process and plot NOAA ObsPack data. For this release, we are focued on
the \(CH_4\) ObsPack GLOBALVIEW+ product. The general process consists
in create a summary of the ObsPack files, reading them in an iteration
process, filter and generating another output and plots.

\hypertarget{installation}{%
\section{Installation}\label{installation}}

To install \texttt{robspack}, the user must have installed the R package
remotes and run the following script. This process will install all the
required dependencies, such as data.table, \texttt{cptcity}, an R
package with more than 7000 color palettes, and \texttt{lubridate}, a
package to manage time and dates \citep{lu, cpt}. Then, we call the
libraries to load the function into the environment.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"ibarraespinosa/robspack"}\NormalTok{)}
\FunctionTok{library}\NormalTok{(robspack)}
\end{Highlighting}
\end{Shaded}

\hypertarget{overview}{%
\section{Overview}\label{overview}}

\texttt{robspack} is a collection of function organized together to read
and process ObsPack files \citep{masarie2014obspack}. The general
process consists in create a summary of the ObsPack files, reading them
in an iteration process, filter and generating another output. In other
to facilitate this task, we are letting to the public the directory
shown below. This directory includes a README file with all the required
information and scripts generate process each category. The structure of
the directory is:

https://github.com/ibarraespinosa/robspack/tree/main/rscripts

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{|}\ExtensionTok{{-}{-}}\StringTok{"README.md"}
\KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ r}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ index.R}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ aircore\_year.R }
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ aircraft\_year.R}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ flask\_non\_noaa\_year.R}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ surface\_insitu\_year.R}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ tower\_insitu\_year.R}
    \KeywordTok{|}\ExtensionTok{{-}{-}}\NormalTok{ inputs\_inv.R}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=1\linewidth,height=0.5\textheight]{flow/Slide1} \caption{Methane observations (ppb) by towers-insitu}\label{fig:flow}
\end{figure}

The file \texttt{index.R} creates a summary of ObsPack, generates the
directories master, receptor and obs, and store the summaries in obs
directory for each category. Then, the scripts for each category are run
to generate the master and receptor files. The next step consists in
running HYSPLIT and obtaining footprints (not in the scope of this
manuscript). At last, the script \texttt{inputs\_inv.R} checks each
footprint file for each category and generates the final receptor list
files. The process is described in detail in the following example for
tower insitu.

The first step consists in constructing a summary for the ObsPack. This
is required to read the data, but also, identify agl, which is present
in some of the file names. This function returns a data.frame.
Optionally, the user can indicate a path to store the data.frame.
obs\_summary also prints a summary of the data. The second argument is
the categories, and by default includes the categories shown below, to
account for all the files. Then the summary data.frame contains the
columns id as the full path to each file, name which is the name or
relative path of the file, n which is just an id, sector such as tower,
and the column agl which indicates the agl indicated in the name of the
file if available. To read the documentation of this function, the user
must run \texttt{?obs\_summary}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{categories }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"aircraft{-}pfp"}\NormalTok{,}\StringTok{"aircraft{-}insitu"}\NormalTok{,}\StringTok{"surface{-}insitu"}\NormalTok{,}
  \StringTok{"aircore"}\NormalTok{,}\StringTok{"surface{-}pfp"}\NormalTok{,}\StringTok{"tower{-}insitu"}\NormalTok{,}\StringTok{"shipboard{-}insitu"}\NormalTok{,}\StringTok{"flask"}\NormalTok{)}
\NormalTok{obs }\OtherTok{\textless{}{-}} \StringTok{"../../../obspack\_ch4\_1\_GLOBALVIEWplus\_v4.0\_2021{-}10{-}14/data/txt"}
\NormalTok{index }\OtherTok{\textless{}{-}} \FunctionTok{obs\_summary}\NormalTok{(}\AttributeTok{obs =}\NormalTok{ obs, }\AttributeTok{verbose =}\NormalTok{ F)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:kable}Summary of ObsPack}
\centering
\begin{tabular}[t]{l|r}
\hline
sector & N\\
\hline
aircraft-pfp & 40\\
\hline
aircraft-insitu & 11\\
\hline
flask & 101\\
\hline
surface-insitu & 124\\
\hline
aircore & 1\\
\hline
surface-pfp & 33\\
\hline
tower-insitu & 51\\
\hline
shipboard-insitu & 1\\
\hline
Total sectors & 362\\
\hline
\end{tabular}
\end{table}

There are 362 files in the ObsPack directory. The printed information
also shows the total at the bottom, as the sum of the individual file by
sector. This is to ensure that the sum of files is equal to the total
number of files found, shown at the top. furthermore, the printed
information also shows that there are 136 files with the agl explicitly
mentioned in the name of the file. Sometimes we need more information
about the site. For instance, what do the observations start and end.
Then, we added the function \texttt{obs\_table}, which calculates
statistics summary of ``time'' and other numeric variables by file name,
sector, site, altitude and mode. For instance, the observations in the
site ``SCT'' in South Carolina, USA, were between ``2015-08-19 21:30:00
UTC'' and ``2020-12-31 23:30:00 UTC''. In figure @ref(fig:tiobs) we see
the average of methane concentrations over each tower-insitu site.
Higher concentration are found over Russia. Over the United States
(U.S.), most of towers are found in the south to east coast.

\begin{figure}
\centering
\includegraphics{paper_elsevier_files/figure-latex/tiobs-1.pdf}
\caption{Methane observations (ppb) by towers-insitu}
\end{figure}

\hypertarget{functions}{%
\subsection{Functions}\label{functions}}

The \texttt{robspack} functions are shown in table 1.

Table 1. Functions and classes in \texttt{robspack}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3108}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6892}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Function
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{invfile} & Class with \texttt{print}, \texttt{summary} and
\texttt{plot} methods \\
\texttt{obs\_addltime()} & Add local time based on metadata and
longitude \\
\texttt{obs\_addtime()} & Add UTC time \\
\texttt{obs\_agg()} & Aggregates ObsPack by time \\
\texttt{obs\_find\_receptors()} & Find expected receptors and NetCDF
files \\
\texttt{obs\_format()} & Format for some columns of data.table \\
\texttt{obs\_freq()} & Return numeric vector in intervals \\
\texttt{obs\_invfiles()} & Construct \texttt{invfile} objects \\
\texttt{obs\_list.dt()} & Rbind list of data.frames with different
names \\
\texttt{obs\_meta()} & Reads ObsPack metadata \\
\texttt{obs\_out()} & Outersect, opposed as intersect \\
\texttt{obs\_rbind()} & Rbind data.frames with different names \\
\texttt{obs\_read()} & Read files, and add metadata as columns \\
\texttt{obs\_read\_csvy()} & Read csvy file and prints yaml header \\
\texttt{obs\_roundtime()} & Round seconds from ``POSIXct'' ``POSIXt''
classes \\
\texttt{obs\_summary()} & Construct summary of ObsPack as a
data.frame \\
\texttt{obs\_table()} & Return a data.frame with summary of data \\
\texttt{obs\_trunc()} & Trunc numbers with a desired number of
decimals \\
\texttt{obs\_write()} & Write CSVY to disk, YAML followed by
tabulated \\
\end{longtable}

\hypertarget{application-for-towers-in-situobspack-summary}{%
\section{Application for towers in situObsPack
summary}\label{application-for-towers-in-situobspack-summary}}

\hypertarget{read-data}{%
\subsection{Read data}\label{read-data}}

Once the summary is built, the function obs\_read will read the files
available in the index file previously generated. Here we selected the
category ``tower-insitu''. The argument verbose prints which files are
being read each time, by default. At the end, this function prints the
total number of observations by type of altitude (agl or asl).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{obs\_read}\NormalTok{(}\AttributeTok{index =}\NormalTok{ index,}
               \AttributeTok{categories =} \StringTok{"tower{-}insitu"}\NormalTok{,}
               \AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We added a function to plot the data read from ObsPack. The y-axis is
the field value and the x-axis is by default time. The data illustrated
sorted by color is the field site\_code, with the default number of 3
sites. The argument pal is to define the color palette, used by the
internally imported function cptcity::cpt.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{obs\_plot}\NormalTok{(}\AttributeTok{dt =}\NormalTok{ df, }\AttributeTok{time =} \StringTok{"time"}\NormalTok{, }\AttributeTok{yfactor =} \FloatTok{1e+09}\NormalTok{, }\AttributeTok{cex =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Found the following sites: 
##  [1] AZV   BRZ   BSD   CRV   DEM   DVV   GCI01 GCI02 GCI03 GCI04 HUN   IGR  
## [13] KRS   LEF   MRC   NOY   RGL   SCT   SVV   TAC   VGN   WGC   WSD   YAK  
## Plotting the following sites: 
## [1] AZV BRZ
\end{verbatim}

\begin{figure}
\centering
\includegraphics{paper_elsevier_files/figure-latex/obsplot-1.pdf}
\caption{First two sites in ObsPack}
\end{figure}

Before sub setting the data, tower-insitu has about 2.32 million
observations. These observations are made between 2004 and 2020. The
identification of the altitude and type is critical, then we developed
an approach based on the availability of data:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Identify agl from the name of the tile.
\item
  If agl not present, search fill\_values used in elevation and
  transform them into NA (not available)
\item
  If agl is not present, agl = altitude - elevation.
\item
  If there are some NA in elevation, will result some NA in agl
\item
  A new column is added named altitude\_final to store agl or asl
\item
  Another column named type\_altitude is added to identify agl or asl.
\item
  If there is any case NA in altitude\_final, type\_altitude is ``not
  available''
\end{enumerate}

\hypertarget{filtering}{%
\subsection{Filtering}\label{filtering}}

ObsPack includes global observations and sometimes we need to extract
data for a specific region and periods of time. In this part we include
spatial and temporal parameters to filter data. The year of interest is
2020, but we also included December of 2019 and January of 2021. The
spatial filtering is done by using the coordinates, in this case
covering North America. After filtering by space and time, we have
\ensuremath{2.322369\times 10^{6}} million observations.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{north }\OtherTok{\textless{}{-}} \DecValTok{80}
\NormalTok{south }\OtherTok{\textless{}{-}} \DecValTok{10}
\NormalTok{west }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{170}
\NormalTok{east }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{50}
\NormalTok{max\_altitude }\OtherTok{\textless{}{-}} \DecValTok{8000}
\NormalTok{evening }\OtherTok{\textless{}{-}} \DecValTok{14}\SpecialCharTok{:}\DecValTok{15}

\NormalTok{yy }\OtherTok{\textless{}{-}} \DecValTok{2020}
\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(df[year }\SpecialCharTok{==}\NormalTok{ yy }\SpecialCharTok{{-}} \DecValTok{1} \SpecialCharTok{\&}\NormalTok{ month }\SpecialCharTok{==} \DecValTok{12}\NormalTok{],}
\NormalTok{            df[year }\SpecialCharTok{==}\NormalTok{ yy],}
\NormalTok{            df[year }\SpecialCharTok{==}\NormalTok{ yy }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{\&}\NormalTok{ month }\SpecialCharTok{==} \DecValTok{1}\NormalTok{])}

\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ df[altitude\_final }\SpecialCharTok{\textless{}}\NormalTok{ max\_altitude }\SpecialCharTok{\&}
\NormalTok{           latitude }\SpecialCharTok{\textless{}}\NormalTok{ north }\SpecialCharTok{\&}
\NormalTok{           latitude }\SpecialCharTok{\textgreater{}}\NormalTok{ south }\SpecialCharTok{\&}
\NormalTok{           longitude }\SpecialCharTok{\textless{}}\NormalTok{ east }\SpecialCharTok{\&}
\NormalTok{           longitude }\SpecialCharTok{\textgreater{}}\NormalTok{ west]}
\end{Highlighting}
\end{Shaded}

\hypertarget{time}{%
\subsection{Time}\label{time}}

The function obs\_addtime adds time columns timeUTC, timeUTC\_start
which shows the start time of each observation and timeUTC\_end which
shows the end time for each observation. Then we need to identify the
local time with the function add\_ltime. This is important because to
identify observations in the evening in local time for modeling
purposes. add\_ltime uses two methods, first identifying the time
difference with utc by identifying the metadata column
``site\_utc2lst''. If this information is not available, with the
aircrafts for instance, the local time is calculated with an
approximation based on longitude:

\[
lt = UTC + longitude/15 * 60 * 60
\] Where lt is the local time, UTC the time, and longitude is the
coordinate. Then, the time is cut every two hours. We also identify the
local time to select evening hours.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df2 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_addtime}\NormalTok{(df)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Adding timeUTC
## Adding timeUTC_start
## Adding timeUTC_end
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df2}\SpecialCharTok{$}\NormalTok{timeUTC }\OtherTok{\textless{}{-}} \FunctionTok{cut}\NormalTok{(}\AttributeTok{x =}\NormalTok{ df2}\SpecialCharTok{$}\NormalTok{timeUTC}\SpecialCharTok{+}\DecValTok{3600}\NormalTok{,}
                   \AttributeTok{breaks =} \StringTok{"2 hour"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{as.character}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{as.POSIXct}\NormalTok{(}\AttributeTok{tz =} \StringTok{"UTC"}\NormalTok{)}
\NormalTok{df3 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_addltime}\NormalTok{(df2)}
\NormalTok{df3 }\OtherTok{\textless{}{-}}\NormalTok{ df3[lh }\SpecialCharTok{\%in\%}\NormalTok{ evening]}
\end{Highlighting}
\end{Shaded}

Now there are 8391 observations. At this point we can calculate the
averages of several columns by the cut time. The function obs\_agg does
this aggregation as shown in the following lines of code. The argument
gby establish the function used to aggregate cols, in this case the
function ``mean'' by time and altitude. Finally, we add local time
again.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df4 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_agg}\NormalTok{(}\AttributeTok{dt =}\NormalTok{ df3,}
               \AttributeTok{gby =} \StringTok{"mean"}\NormalTok{,}
               \AttributeTok{cols =} \FunctionTok{c}\NormalTok{(}\StringTok{"value"}\NormalTok{, }\StringTok{"latitude"}\NormalTok{, }\StringTok{"longitude"}\NormalTok{, }\StringTok{"type\_altitude"}\NormalTok{,}
                        \StringTok{"dif\_time"}\NormalTok{, }\StringTok{"year\_end"}\NormalTok{, }\StringTok{"site\_utc2lst"}\NormalTok{),}
               \AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{,}
               \AttributeTok{byalt =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Detecting dif_time. Adding ending times
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df5 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_addltime}\NormalTok{(df4)}
\end{Highlighting}
\end{Shaded}

Now there are 4394 observations. Towers can have observations at
different heights. Here we need to select one site with the observations
registered at the highest height. The column with the height is named
altitude\_final and the max altitude was named max\_altitude. Then, we
print the altitudes of each site.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df5[,}
\NormalTok{    max\_altitude }\SpecialCharTok{:=} \FunctionTok{max}\NormalTok{(altitude\_final),}
\NormalTok{    by }\OtherTok{=}\NormalTok{ site\_code]}
\NormalTok{df5[,}
    \FunctionTok{c}\NormalTok{(}\StringTok{"site\_code"}\NormalTok{,}
      \StringTok{"altitude\_final"}\NormalTok{,}
      \StringTok{"max\_altitude"}\NormalTok{)] }\SpecialCharTok{|\textgreater{}} \FunctionTok{unique}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     site_code altitude_final max_altitude
##  1:       CRV           17.0           32
##  2:       CRV           32.0           32
##  3:       CRV            4.9           32
##  4:       LEF          122.0          396
##  5:       LEF           30.0          396
##  6:       LEF          396.0          396
##  7:       SCT          305.0          305
##  8:       SCT           31.0          305
##  9:       SCT           61.0          305
## 10:       WGC           30.0          483
## 11:       WGC          483.0          483
## 12:       WGC           91.0          483
\end{verbatim}

\hypertarget{saving-master-as-text-and-csvy}{%
\subsection{Saving master as text and
csvy}\label{saving-master-as-text-and-csvy}}

Now that we have all the required information, we can save the files.
Here, we name the data.frame as master, because it contains all the
information. This is important because some fields can be used in the
future, and for traceability. For convenience, time variables are
transformed into character before writing into the disk. The separation
is space '' ``.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{master }\OtherTok{\textless{}{-}}\NormalTok{ df5}
\NormalTok{master}\SpecialCharTok{$}\NormalTok{timeUTC }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(master}\SpecialCharTok{$}\NormalTok{timeUTC)}
\NormalTok{master}\SpecialCharTok{$}\NormalTok{timeUTC\_end }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(master}\SpecialCharTok{$}\NormalTok{timeUTC\_end)}
\NormalTok{master}\SpecialCharTok{$}\NormalTok{local\_time }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(master}\SpecialCharTok{$}\NormalTok{local\_time)}

\FunctionTok{fwrite}\NormalTok{(master, }
       \AttributeTok{file =} \StringTok{"tower\_insitu\_2020.txt"}\NormalTok{,}
       \AttributeTok{sep =} \StringTok{" "}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The format Comma Separated Value with YAML (CSVY)\footnote{https://csvy.org/}
consists in a typical CSV with a YAML header. The function obs\_write
includes the argument notes which allows adding custom notes at the
header of the file. Below the notes, obs\_write adds the output of the R
function \texttt{str}, which provides a vertical summary of the data,
known as structure.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{obs\_write\_csvy}\NormalTok{(}\AttributeTok{dt =}\NormalTok{ master,}
              \AttributeTok{notes =} \StringTok{"tower 2020"}\NormalTok{,}
              \AttributeTok{out =} \StringTok{"tower\_insitu\_2020.csvy"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To check the YAML header we read the first 38 lines of the files that
were generated. Here we can see the column names, type of data and first
observations. The YAML header is delimited by the characters ``- - -''
(not shown here).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{readLines}\NormalTok{(}\StringTok{"tower\_insitu\_2020.csvy"}\NormalTok{)[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{38}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\hypertarget{saving-receptors}{%
\subsection{Saving receptors}\label{saving-receptors}}

We need to filter some columns from the master files in a new object
called receptors. This is needed because internally we run HYSPLIT
\citep{hy} using the information from the receptors. In the case of a
tower, we need to select observations with the highest altitude. The
specific columns are selected as shown on the following code. We are
selecting the ending times, because later HYSPLIT is run backwards based
on the time of measurement, between ending and starting times. The
columns about time are formatted to have two characters. For instance,
the month 1, is formatted as ``01''. We also need to filter for
type\_altitude equal 0, representing agl observations , or equal to 1,
asl.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{receptor }\OtherTok{\textless{}{-}}\NormalTok{ master[altitude\_final }\SpecialCharTok{==}\NormalTok{ max\_altitude,}
                   \FunctionTok{c}\NormalTok{(}\StringTok{"site\_code"}\NormalTok{,}
                     \StringTok{"year"}\NormalTok{, }\StringTok{"month"}\NormalTok{, }\StringTok{"day"}\NormalTok{,}
                     \StringTok{"hour"}\NormalTok{, }\StringTok{"minute"}\NormalTok{, }\StringTok{"second"}\NormalTok{,}
                     \StringTok{"latitude"}\NormalTok{, }\StringTok{"longitude"}\NormalTok{,}
                     \StringTok{"altitude\_final"}\NormalTok{, }\StringTok{"type\_altitude"}\NormalTok{,}
                     \StringTok{"year\_end"}\NormalTok{, }\StringTok{"month\_end"}\NormalTok{, }\StringTok{"day\_end"}\NormalTok{, }\StringTok{"hour\_end"}\NormalTok{,}
                     \StringTok{"minute\_end"}\NormalTok{, }\StringTok{"second\_end"}\NormalTok{)]}
\NormalTok{receptor}\SpecialCharTok{$}\NormalTok{altitude\_final }\OtherTok{\textless{}{-}} \FunctionTok{round}\NormalTok{(receptor}\SpecialCharTok{$}\NormalTok{altitude\_final)}
\NormalTok{receptor }\OtherTok{\textless{}{-}} \FunctionTok{obs\_format}\NormalTok{(receptor)}

\ControlFlowTok{if}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(receptor\_agl) }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{) \{}
  \FunctionTok{fwrite}\NormalTok{(}\AttributeTok{x =}\NormalTok{ receptor\_agl,}
         \AttributeTok{file =} \StringTok{"paper/receptor\_tower\_insitu\_2020\_AGL.txt"}\NormalTok{),}
\NormalTok{  sep }\OtherTok{=} \StringTok{" "}\ErrorTok{)}\NormalTok{\}}

\ControlFlowTok{if}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(receptor\_asl) }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{) \{}
  \FunctionTok{fwrite}\NormalTok{(}\AttributeTok{x =}\NormalTok{ receptor\_asl,}
         \AttributeTok{file =} \StringTok{"paper/receptor\_tower\_insitu\_2020\_ASL.txt"}\NormalTok{),}
\NormalTok{  sep }\OtherTok{=} \StringTok{" "}\ErrorTok{)}\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{recommendation-for-other-applications}{%
\subsection{Recommendation for other
applications}\label{recommendation-for-other-applications}}

The approach to generate receptors depends on each type of observation
and other considerations. For instance, aircraft with continuous
observations at each second can be filtered and averaged every 20
seconds. In that way, the footprints are still representative and it
would not be necessary to run HYSPLIT every second. Of course, it
depends on the application and objective of the study. For this
manuscript, we are presenting the generation of receptors based on tower
observations.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In this manuscript we presented an robskpack, an R package to read and
process CH4 ObsPack GLOBALVIEW+ published by the Global Monitoring
Laboratory (GML) from the National Oceanographic and Atmospheric
Administration (NOAA). robspack reads the text data which have different
headers and organizes them in a common format. Then, this software
applies calculations to filter observations by time and space. Finally,
this software generates receptors in a suitable format that allows it to
run HYSPLIT and generate footprints. This software does not provide
methods to run HYSPLIT, but the user can follow the site
https://www.ready.noaa.gov/HYSPLIT.php.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Funding: This project is funded by the NOAA Climate Program Office AC4
and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was
supported by the NOAA cooperative agreement NA22OAR4320151.

\renewcommand\refname{References}
\bibliography{mybibfile.bib}

\end{document}

### 2023-07-12 R/01_index_obs.R

@@ -58,7 +58,7 @@ obs_index <- function(obs,
   xx <- index[, .N, by = sector]
   dx <- data.table::data.table(sector = c("Total sectors"),
                                N = sum(xx$N))
-  print(rbind(xx, dx))
+  if(verbose) print(rbind(xx, dx))
 
 
   # function to extract last n characters

### 2023-07-12 R/01_summary_obs.R

@@ -11,6 +11,7 @@
 #' @param lnchar Integer; last nchar, default = 11.
 #' @param out Path to the ObsPack index output
 #' @param verbose Logical to show more information
+#' @param aslist Logical to return list of index and summary
 #' @return A data.frame with with an index of the obspack Globalview.
 #' @importFrom data.table fwrite ".N" ":=" setDT data.table
 #' @export
@@ -32,7 +33,8 @@ obs_summary <- function(obs,
                                        "flask"),
                         lnchar = 11,
                         out = paste0(tempfile(), "_index.csv"),
-                        verbose = TRUE){
+                        verbose = TRUE,
+                        aslist = FALSE){
 
 
   x <- list.files(obs,
@@ -52,11 +54,13 @@ obs_summary <- function(obs,
   }
 
 
-  cat(paste0("Number of files of index: ", nrow(index), "\n"))
+  if(verbose) cat(paste0("Number of files of index: ", nrow(index), "\n"))
   xx <- index[, .N, by = sector]
   dx <- data.table::data.table(sector = c("Total sectors"),
                                N = sum(xx$N))
-  print(rbind(xx, dx))
+
+  lx1 <- rbind(xx, dx)
+  if(verbose) print(lx1)
 
 
   # function to extract last n characters
@@ -94,5 +98,11 @@ obs_summary <- function(obs,
     cat(paste0("Detected ", sum(nai, na.rm = T), " files without agl\n"))
   }
 
-  return(index)
+  if(aslist) {
+    return(list(summary = lx1,
+                files = index))
+
+  } else {
+    return(index)
+  }
 }

### 2023-07-12 R/09_obs_table.R

@@ -36,7 +36,7 @@
 #' }
 obs_table <- function(index,
                       categories = "tower-insitu",
-                      verbose = TRUE,
+                      verbose = FALSE,
                       n_site_code = 15,
                       n_site_name = 15,
                       n_site_country = 18,

### 2023-07-12 rscripts/r/inputs_inv.R

@@ -260,7 +260,7 @@ hela <- c(dfaircraft$footprints_hera_hysplit,
 
 
 fwrite(x = as.data.table(hela),
-       file = "invfiles/footprints_hera_hysplit.txt",
+       file = "receptors/footprints_hera_hysplit.txt",
        quote = FALSE,
        row.names = FALSE,
        col.names = FALSE,
@@ -274,7 +274,7 @@ obsh <- c(dfaircraft$obs_hysplit,
 
 
 fwrite(x = as.data.table(obsh),
-       file = "invfiles/obs_hysplit.txt",
+       file = "receptors/obs_hysplit.txt",
        quote = FALSE,
        row.names = FALSE,
        col.names = FALSE,
@@ -289,7 +289,7 @@ reih <- rbind(dfaircraft$receptor_info_hysplit,
 
 
 fwrite(x = reih,
-       file = "invfiles/receptor_info_hysplit.txt",
+       file = "receptors/receptor_info_hysplit.txt",
        quote = FALSE,
        sep = "\t",
        row.names = FALSE,
@@ -301,7 +301,7 @@ reihnos <- reih
 reihnos$Scale <- NULL
 
 fwrite(x = reihnos,
-       file = "invfiles/receptor_info_hysplit_noscale.txt",
+       file = "receptors/receptor_info_hysplit_noscale.txt",
        quote = FALSE,
        sep = "\t",
        row.names = FALSE,

### 2023-07-14 paper_elsevier/paper_elsevier.Rmd
update robs

@@
 @lu2021global, characterized global methane emissions in 
-between 2014 and 2017, including a comparison with Greenhouse gases Observing 
-SATellite (GOSAT) data. 
-@saunois2016global 
+between 2014 and 2017, including a comparison with Greenhouse gases 
+Observing SATellite (GOSAT) data. 
+@saunois2016global.
 At regional scale, @lu2022methane performed another studied focused
@@
 depend on the type of file. For instance, the metadata
-from aircrfats is different than surface stations, but all the files
+from aircrafts is different than surface stations, but all the files
 include concentrations and other critical fields. 
-An important information is the scale of measurement, which isXX.
-Cite Importance of scale and what is WMOX2014.
 Given the complexity of ObsPack format, reading and analyzing the data
@@
 process and plot NOAA ObsPack data. For this release, we are 
-focued on the $CH_4$ ObsPack GLOBALVIEW+ product.
-The general process 
-consists in create a 
+focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
+The general process consists in creating a 
 summary of the ObsPack files, reading them in an iteration process, 
-filter and generating another output and plots.
+filtering, and generating another output and plots.
 

### 2023-10-03 paper_elsevier/paper_elsevier.Rmd
adding NetCDF support

@@
 obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
-index <- obs_summary(obs = obs, verbose = F)
+index <- obs_summary(obs = obs, 
+                     categories = c("aircraft",
+                                    "surface",
+                                    "aircore",
+                                    "shipboard",
+                                    "tower"),
+                     verbose = T)
 ```
@@
 ```{r kable, echo = FALSE}
-ta <- obs_summary(obs = obs, verbose = F, aslist = TRUE)$summary
+ta <- obs_summary(obs = obs, 
+                  categories =  c("aircraft",
+                                    "surface",
+                                    "aircore",
+                                    "shipboard",
+                                    "tower"),
+                  verbose = F, aslist = TRUE)$summary
 knitr::kable(ta, format = "latex", caption = "Summary of ObsPack")
@@
 library(ggplot2)
-dft <- obs_table(index =  index, categories = "tower-insitu")
+dft <- obs_table(index =  index, categories = "tower")
 dft <- dft[stat == "mean"]

### 2023-10-03 R/01_index_obs.R

@@ -1,100 +1,0 @@
-#' Summary of the ObsPack files
-#'
-#' This function returns a data.frame index of all files
-#' available in obspack.
-#'
-#' @param obs Path to the Obspack GLOBALview txt data
-#' @param categories character; default are c("aircraft-pfp", "aircraft-insitu",
-#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
-#' "flask"). The ideia is that,as the file names include these words,
-#' this function identifies which files has these words and add them as columns.
-#' @param lnchar Integer; last nchar, default = 11.
-#' @param out Path to the Obspack index output
-#' @param verbose Logical to show more information
-#' @return A data.frame with with an index of the obspack Globalview.
-#' @rdname obs_summary
-#' @importFrom data.table fwrite ".N" ":=" setDT data.table
-#' @export
-#' @examples {
-#' \dontrun{
-#' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
-#' index <- obs_summary(obs)
-#' }
-#' }
-obs_index <- function(obs,
-                      categories = c("aircraft-pfp",
-                                     "aircraft-insitu",
-                                     "surface-insitu",
-                                     "tower-insitu",
-                                     "aircore",
-                                     "surface-pfp",
-                                     "shipboard-insitu",
-                                     "flask"),
-                      lnchar = 11,
-                      out = paste0(tempfile(), "_index.csv"),
-                      verbose = TRUE){
-  cat("Please, use `obs_summary` instead")
-
-
-  x <- list.files(obs,
-                  full.names = T)
-
-  na <- list.files(obs,
-                   full.names = F)
-
-  # create index ####
-  # the idea
-  index <- data.table::data.table(id = x, name = na)
-  index$n <- 1:nrow(index)
-  sector <- NULL
-  for(i in seq_along(categories)) {
-    index[grepl(pattern = categories[i], x = index$na),
-          sector := categories[i]]
-  }
-
-
-  cat(paste0("Number of files of index: ", nrow(index), "\n"))
-  xx <- index[, .N, by = sector]
-  dx <- data.table::data.table(sector = c("Total sectors"),
-                               N = sum(xx$N))
-  if(verbose) print(rbind(xx, dx))
-
-
[…]

### 2023-10-03 R/01_summary_obs.R

@@ -1,8 +1,10 @@
-#' Summary of the ObsPack files
+#' @title Summary of the ObsPack files (.txt)
 #'
+#' @description
 #' This function returns a data.frame index of all files
 #' available in ObsPack.
 #'
+#' @family obs_summary
 #' @param obs Path to the Obspack GLOBALview txt data
 #' @param categories character; default are c("aircraft-pfp", "aircraft-insitu",
 #' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
@@ -36,7 +38,6 @@ obs_summary <- function(obs,
                         verbose = TRUE,
                         aslist = FALSE){
 
-
   x <- list.files(obs,
                   full.names = T)
 
@@ -54,35 +55,35 @@ obs_summary <- function(obs,
   }
 
 
-  if(verbose) cat(paste0("Number of files of index: ", nrow(index), "\n"))
+  cat(paste0("Number of files of index: ", nrow(index), "\n"))
   xx <- index[, .N, by = sector]
   dx <- data.table::data.table(sector = c("Total sectors"),
                                N = sum(xx$N))
-
-  lx1 <- rbind(xx, dx)
-  if(verbose) print(lx1)
+  if(verbose) print(rbind(xx, dx))
 
 
-  # function to extract last n characters
-  sr <- function(x, n){
-    substr(x, nchar(x)-n+1, nchar(x))
-  }
+  # detecting file extension
+  # using tools::file_ext function instead of import
+  fiex <- fex(index$id[1])
 
-  # add the last 11 characters of each name file'
-  agl <- NULL
-  id <- NULL
-  index[grepl(pattern = "magl", x = x),
-        agl := sr(id, 11)]
+  idfiex <- ifelse(fiex == "nc", 10, 11)
+    # add the last 11 characters of each name file'
+    agl <- NULL
+    id <- NULL
+    index[grepl(pattern = "magl", x = x),
+          agl := sr(id, idfiex)]
 
-  # This line replaces removes the characters magl.txt
-  # for instance, remove "magl.txt" from -11magl.txt
-  index$agl <- gsub("magl.txt", "", index$agl)
+    # This line replaces removes the characters magl.txt
+    # for instance, remove "magl.txt" from -11magl.txt
+    index$agl <- gsub(paste0("magl.", fiex), "", index$agl)
+    # assuming d-{number}magl.txt
+    index$agl <- gsub("d", "", index$agl)
 
-  # check
-  index
[…]

### 2023-10-03 R/02_read_obs.R

@@ -1,5 +1,6 @@
-#' Read obspack
+#' @title Read obspack (.txt)
 #'
+#' @description
 #' Each obspack file has a header with metadata and this
 #' function reads selected fields from the metadata and add
 #' them as columns. This new columns are used later to
@@ -300,4 +301,307 @@ obs_read <- function(index,
 
   }
 
+}
+
+#' @title Read obspack (.nc)
+#'
+#' @description
+#' Each obspack file has a header with metadata and this
+#' function reads selected fields from the metadata and add
+#' them as columns. This new columns are used later to
+#' be filtered
+#'
+#' @param index data.table
+#' @param categories character; ONE category
+#' : of c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask").
+#' @param verbose Logical to show more information
+#' @param n_site_code number of characters extraced from metadata after search
+#' @param n_site_name number of characters extraced from metadata after search
+#' @param n_site_country number of characters extraced from metadata after search
+#' @param n_dataset_project number of characters extraced from metadata after search
+#' @param n_lab number of characters extraced from metadata after search
+#' @param n_scales number of characters extraced from metadata after search
+#' @param n_site_elevation number of characters extraced from metadata after search
+#' @param n_altitude_comment number of characters extraced from metadata after search
+#' @param n_utc number of characters extraced from metadata after search
+#' @param fill_value fill value. Appeared in aoa_aircraft-flask_19_allvalid.txt
+#' @param as_list Logical to return as list
+#' @return A data.frame with with an index obspack.
+#' @importFrom data.table fwrite ".N" ":=" rbindlist "%chin%"
+#' @export
+#' @note The identification of the altitude and type is critical.
+#' The approach used here consists of:
+#' 1. Identify agl from the name of the tile.
+#' 2. If magl not present, search dill_values used in elevation and
+#' transform them into NA (not available)
+#' 3. If magl is not present, agl = altitude - elevation
+#' 4. If there are some NA in elevation, will result some NA in agl
+#' 5. A new column is added named `altitude_final` to store agl or asl
+#' 6. Another column named `type_altitude` is added to identify "magl" or "masl"
+#' 7. If there is any case NA in `altitude_final`,
[…]

### 2023-10-03 R/03_addtime_obs.R

@@ -1,10 +1,10 @@
-#' Add times into obspack
+#' @title Add times into obspack
 #'
+#' @description
 #' This function add timeUTC as POSIX class,
 #' local time and ending sampling time
 #'
-#'
-#'
+#' @family time
 #' @param dt obspack data.table
 #' @param verbose obspack data.table
 #' @return A data.frame with with an index obspack.
@@ -48,7 +48,7 @@ obs_addtime <- function(dt,
   time_interval <- NULL
   timeUTC_start <- NULL
   if(any("time_interval" == names(dt))) {
-    dt[, timeUTC_end := fifelse(
+    dt[, timeUTC_end := data.table::fifelse(
       is.na(time_interval),
       timeUTC_start + (midpoint_time - start_time)*2,
       timeUTC_start + time_interval
@@ -82,5 +82,49 @@ obs_addtime <- function(dt,
   # 0,
   # dt$new_second)
 
+  return(dt)
+}
+
+
+
+#' @title local hour (bsed on longitude and time)
+#' @family time
+#' @name obs_addltime
+#' @description Calculate an approximation of local hour
+#' @param dt data.table
+#' @param timeUTC Character indicating the Time column as POSIXct
+#' @param utc2lt Character indicating the integer column to convert to local time
+#' if available
+#' @param longitude Character indicating the column with the lingitude
+#' @importFrom data.table hour
+#' @return data.table with local time columns
+#' @note time depending n longitude is by John Miller (GML/NOAA)
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' }
+#' }
+obs_addltime <- function(dt,
+                         timeUTC = "timeUTC",
+                         utc2lt = "site_utc2lst",
+                         longitude = "longitude") {
+
+  # Priority site_utc2lst and if it is not available,
+  dt$local_time <- dt[[timeUTC]] + as.numeric(dt[[utc2lt]])*60*60
+
+
+  dt$local_time <- ifelse(
+    is.na(dt$local_time),
+    dt[[timeUTC]] + dt[[longitude]]/15*60*60, #john miller approach
+    dt$local_time)
+
+  if(inherits(dt$local_time, "numeric")) {
+    dt$local_time <- as.POSIXct(dt$local_time,
+                                origin = "1970-01-01",
+                                tz = "UTC")
+  }
+
+  dt$lh <- data.table::hour(dt$local_time)
   return(dt)
 }

### 2023-10-03 R/helpers.R

@@ -60,51 +60,6 @@ obs_list.dt <- function(ldf, na, verbose = TRUE) {
 }
 
 
-#' @title Number to character
-#' @family helpers
-#' @description transform integer to character and add zero at left
-#' @param x integer
-#' @param char Character to be added ad the beginning of x
-#' @return transform integer in character
-#' @name robspack-deprecated
-#' @seealso \code{\link{robspack-deprecated}}
-#' @keywords internal
-NULL
-# obs_addzero <- function(x, char) {
-#
-#   # if(!inherits(x, "integer")) stop("x must be integer")
-#
-#   dd <- as.character(x)
-#   nch <- nchar(dd)
-#   nchmax <- max(nchar(dd))
-#
-#   difch <- nchmax - nch
-#
-#   if(length(unique(nch)) == 1 ) {
-#     if (unique(nch) == 1) {
-#       message("Character has same length. Adding one zero.")
-#       y <- paste0("0", x)
-#     } else {
-#       y <- x
-#     }
-#
-#   } else {
-#     unlist(lapply(seq_along(difch), function(i) {
-#       paste(rep("0", difch[i]), collapse = "")
-#     })) -> zero
-#
-#     y <- paste0(zero, x)
-#
-#   }
-#
-#
-#   if(!missing(char)) {
-#     y <- paste0(char, y)
-#   }
-#
-#   return(y)
-# }
-
 
 #' @title return numeric vector in intervals
 #' @family helpers
@@ -204,50 +159,6 @@ obs_rbind <- function(dt1,
   dx <- rbind(dt1, dt2, use.names = TRUE)
 
   return(dx)
-}
-
-
-#' @title local hour (bsed on longitude and time)
-#' @family helpers
-#' @name obs_addltime
-#' @description Calculate an approximation of local hour
-#' @param dt data.table
-#' @param timeUTC Character indicating the Time column as POSIXct
-#' @param utc2lt Character indicating the integer column to convert to local time
-#' if available
-#' @param longitude Character indicating the column with the lingitude
-#' @importFrom data.table hour
-#' @return data.table with local time columns
-#' @note time depending n longitude is by John Miller (GML/NOAA)
-#' @export
-#' @examples {
-#' \dontrun{
-#' # Do not run
-#' }
-#' }
-obs_addltime <- function(dt,
-                         timeUTC = "timeUTC",
[…]

### 2023-10-03 R/robspack-deprecated.R

@@ -7,3 +7,163 @@
 #' @name robspack-deprecated
 #' @keywords internal
 NULL
+
+
+
+
+#' @title Index of the ObsPack files (Deprecated)
+#'
+#' @description
+#' This function returns a data.frame index of all files
+#' available in obspack.
+#'
+#' @param obs Path to the Obspack GLOBALview txt data
+#' @param categories character; default are c("aircraft-pfp", "aircraft-insitu",
+#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
+#' "flask"). The ideia is that,as the file names include these words,
+#' this function identifies which files has these words and add them as columns.
+#' @param lnchar Integer; last nchar, default = 11.
+#' @param out Path to the Obspack index output
+#' @param verbose Logical to show more information
+#' @return A data.frame with with an index of the obspack Globalview.
+#' @rdname obs_summary
+#' @importFrom data.table fwrite ".N" ":=" setDT data.table
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' obs <- system.file("data-raw", package = "robspack")
+#' index <- obs_summary(obs)
+#' }
+#' }
+obs_index <- function(obs,
+                      categories = c("aircraft-pfp",
+                                     "aircraft-insitu",
+                                     "surface-insitu",
+                                     "tower-insitu",
+                                     "aircore",
+                                     "surface-pfp",
+                                     "shipboard-insitu",
+                                     "flask"),
+                      lnchar = 11,
+                      out = paste0(tempfile(), "_index.csv"),
+                      verbose = TRUE){
+  # cat("Please, use `obs_summary` instead")
+  #
+  #
+  # x <- list.files(obs,
+  #                 full.names = T)
+  #
+  # na <- list.files(obs,
+  #                  full.names = F)
+  #
+  # # create index ####
+  # # the idea
+  # index <- data.table::data.table(id = x, name = na)
+  # index$n <- 1:nrow(index)
+  # sector <- NULL
+  # for(i in seq_along(categories)) {
+  #   index[grepl(pattern = categories[i], x = index$na),
[…]

### 2024-03-05 README.Rmd
ranem robspack to rtorf

@@
 ---
-title: "robspack"
+title: "R Tools for Obspack, Footprints and Receptors (rtorf)"
 output: github_document
@@
 ![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/robspack)
+activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/rtorf)
 [![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
+status](https://github.com/ibarraespinosa/rtorf/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
 [![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
+status](https://github.com/ibarraespinosa/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
 
@@
 
-`robspack` only depends on `data.table`, which is basically parallel C, 
+`rtorf` only depends on `data.table`, which is basically parallel C, 
 so it can be installed in any machine.
@@
 ```{r, eval = F}
-remotes::install_github("ibarraespinosa/robspack")
+remotes::install_github("ibarraespinosa/rtorf")
 ```
@@
 ```{r, include = T, message=F, warning=F}
-library(robspack)
+library(rtorf)
 library(data.table)
@@
 The scripts are available in the path 
-`https://github.com/ibarraespinosa/robspack/tree/main/rscripts`
+`https://github.com/ibarraespinosa/rtorf/tree/main/rscripts`

### 2024-03-05 rscripts/README.R
rename robspack to rtorf

@@ -1,38 +1,0 @@
-
-# README ####
-# The directory `rscripts` contains the required files
-# to generate the list of receptors using `robspack`
-
-# The scripts will run from a defined working directory
-
-
-# 1. Run index (summary)
-
-# The script r/index.R will create the tree directory
-# to store the processed ObsPack files
-# Also, it will create summaries and store them
-# for each category in the directory `obs`
-
-# 2. Categories in the directory r
-# These scripts need to be run using Rscript command
-#examples:
-# Rscript r/aircore_year.R 2020 obs/obs_aircore.csv.gz
-# Rscript r/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv.gz
-# Rscript r/flask_year.R 2020 obs/obs_flask.csv.gz
-# Rscript r/surface_year.R 2020 obs/obs_surface-insitu.csv.gz
-# Rscript r/tower_year.R 2020 obs/obs_tower-insitu.csv.gz
-
-# Windows users can adapt the following lines
-# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/tower_insitu_year.R 2020 obs/obs_tower-insitu.csv.gz
-# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/aircore_year.R 2020 obs/obs_aircore.csv.gz
-
-
-# 3. Run Hysplit
-# Once if we have the receptor, we can
-# run Hysplit to generate footprints in NetCDF format
-
-# 4. Generate receptor info
-# Once we have the receptor list and the footprints were generated,
-# we can generate the receptor info, which
-# consists in text files.
-# The script is r/inputs_inv.R

### 2024-03-05 rscripts/r/aircore_year.R

@@ -1,173 +1,0 @@
-#example Rscript rscripts/aircore_year.R 2020 obs/obs_aircore.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-
-out <- paste0("master/master_aircore", yy)
-
-message("All missing values are filled with 999999")
-
-cat("                 _     robspackfilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(robspack)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'aircore',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "aircore",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ###
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
[…]

### 2024-03-05 rscripts/r/aircraft_year.R

@@ -1,207 +1,0 @@
-#example Rscript rscripts/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-
-out <- paste0("master/master_aircraft_insitu_", yy)
-
-message("All missing values are filled with 999999")
-
-cat("                 _     robspackfilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(robspack)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# evening <- 14 # for aircraft keep all the local hours
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'aircraft-insitu',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "aircraft-insitu",
-#               verbose = T)
-df  <- fread(ind)
-
-# 3 ECO 2020 ####
-(cat("Reading ECO 2020\n"))
-
-if(Sys.info()[["sysname"]] == "Windows") {
[…]

### 2024-03-05 rscripts/r/flask_non_noaa_year.R

@@ -1,205 +1,0 @@
-#example Rscript rscripts/flask_year.R 2020 obs/obs_flask.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_flask_no_noaa_", yy)
-
-
-
-message("All missing values are filled with 999999")
-
-cat("                 _     robspackfilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(robspack)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'flask',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "flask",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ###
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
-
[…]

### 2024-03-05 rscripts/r/index.R

@@ -1,54 +1,0 @@
-wd <- paste0(getwd(), "/rscripts") #  you can change
-# if you change, define a new path
-# wd <- "path"
-# then set working directory
-setwd(wd)
-
-cat("This is your working directory", getwd(), "\n")
-
-library(data.table)
-library(robspack)
-
-categories <- c(
-  #"aircraft-pfp",
-  "aircraft-insitu",
-  "surface-insitu",
-  "aircore",
-  #"surface-pfp",
-  "tower-insitu",
-  "shipboard-insitu",
-  "flask"
-)
-
-# ObsPack ####
-obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt/"
-
-# Create Directories ####
-# master have files with all columns
-dir.create("master", showWarnings = FALSE)
-
-# receptors have files with required columns
-dir.create("receptors", showWarnings = FALSE)
-
-# receptors have files with required columns
-dir.create("obs", showWarnings = FALSE)
-
-# 1 summary (previously named index) ####
-cat("Running obs_summary(obs)\n")
-index <- obs_summary(obs = obs)
-
-
-# 2 read ####
-
-for(i in seq_along(categories)) {
-
-cat(paste0("Running obs_read(index = index,
-               categories = ", categories[i], "\n"))
-
-df <- obs_read(index = index,
-               categories = categories[i],
-               verbose = T)
-
-fwrite(df, paste0("obs/obs_",categories[i], ".csv.gz"))
-
-}

### 2024-03-05 rscripts/r/inputs_inv.R

@@ -1,310 +1,0 @@
-
-library(colorout)
-library(data.table)
-library(robspack)
-
-dir.create("invfiles")
-
-# aircraft insitu ####
-ma1 <- paste0("/path/master/master_aircraft_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skipping ", ma1[i], "\n")
-    next
-  }
-  dx <- fread(ma1[i])
-  dx$file <- na[i]
-  print(dim(dx))
-  lx[[i]] <- dx
-}
-master <- obs_list.dt(lx)
-# filtering between dec 2017 and jan 2021
-
-master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
-                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
-
-master$nc <- obs_footname(year = master$year,
-                          month = master$month,
-                          day = master$day,
-                          hour = master$hour,
-                          minute = master$minute,
-                          lat = master$latitude,
-                          lon = master$longitude,
-                          alt = master$altitude_final)
-
-path <-  "/path/aircraft/flask/" # param
-
-master$fnc <- paste0(path, master$nc)
-
-master$is_fnc <- file.exists(master$fnc)
-
-table(master$is_fnc)
-
-master <- master[is_fnc == TRUE]
-
-master <- master[!is.na(value)]
-
-master <- master[altitude_final > 10]
-
-dfaircraft <- obs_invfiles(master = master,
-                           path = path,
-                           Type = "continuous",
-                           SubType = "aircraft")
-obs_plot(dfaircraft$master,
-         time = "timeUTC",
-         y = "value",
-         colu = "site_code",
-         yfactor = 1e9,
-         cex = 0.5,
-         type = "l"
-)
-
-
-# tower insitu ####
-ma1 <- paste0("master/master_tower_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skipping ", ma1[i], "\n")
-    next
-  }
[…]

### 2024-03-05 rscripts/r/inputs_invs.R

@@ -1,314 +1,0 @@
-
-library(colorout)
-library(data.table)
-library(robspackfilter)
-
-# dir.create("invfiles")
-
-# aircraft insitu ####
-ma1 <- paste0("master/master_aircraft_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skiiping ", ma1[i], "\n")
-    next
-  }
-  dx <- fread(ma1[i])
-  dx$file <- na[i]
-  print(dim(dx))
-  lx[[i]] <- dx
-}
-master <- obs_list.dt(lx)
-# filtering between dec 2017 and jan 2021
-
-master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
-                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
-#  footprints_hera_hysplit.txt
-master$nc <- obs_footname(year = master$year,
-                          month = master$month,
-                          day = master$day,
-                          hour = master$hour,
-                          minute = master$minute,
-                          lat = master$latitude,
-                          lon = master$longitude,
-                          alt = master$altitude_final)
-
-path <-  "/work/noaa/co2/footprint/hysplit/CPO2021/hysplit4/aircraft/flask/" # param
-
-master$fnc <- paste0(path, master$nc)
-
-master$is_fnc <- file.exists(master$fnc)
-
-table(master$is_fnc)
-
-master <- master[is_fnc == TRUE]
-
-master <- master[!is.na(value)]
-
-master <- master[altitude_final > 10]
-
-dfaircraft <- obs_invfiles(master = master,
-                           path = path,
-                           Type = "continuous",
-                           SubType = "aircraft")
-plot(dfaircraft)
-
-obs_plot(dfsurface$master,
-         time = "timeUTC",
-         y = "value",
-         colu = "site_code",
-         n = c("BRA", "EGB", "EST", "ETL", "LLB"),
-         yfactor = 1e9,
-         cex = 0.5,
-         type = "l",
-         ylim = c(1800, 3000)
-)
-
-
-# Found  760 missing `value` on master
-# It is possible that this script must be run again to explude missing value
-
-
[…]

### 2024-03-05 rscripts/r/surface_insitu_year.R

@@ -1,236 +1,0 @@
-#example Rscript rscripts/surface_year.R 2020 obs/obs_surface-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_surface_insitu_", yy)
-
-
-message("All missing values are filled with 999999")
-
-cat("                 _     robspackfilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(robspack)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14:15
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'surface-insitu',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "surface-insitu",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ####
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
[…]

### 2024-03-05 rscripts/r/tower_insitu_year.R

@@ -1,238 +1,0 @@
-#example Rscript rscripts/tower_year.R 2020 obs/obs_tower-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index",
-       call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_tower_insitu_", yy)
-
-message("All missing values are filled with 999999")
-
-n01 <- ("                  _     robspackfilter              \n")
-n02 <- ("               (`  ).                  _           \n")
-n03 <- ("              (     ).             .:(`  )`.       \n")
-n04 <- (")           _(       '`.          :(   .    )      \n")
-n05 <- ("        .=(`(      .   )     .--  `.  (    ) )      \n")
-n06 <- ("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-n07 <- ("`.     `(       ) )       (   .  )     (   )  ._   \n")
-n08 <- ("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-n09 <- (")  )  ( )       --'       `- __.'         :(      )) \n")
-n10 <- (".-'  (_.'          .')                    `(    )  ))\n")
-n11 <- ("                  (_  )                     ` __.:'          \n")
-n12 <- ("\n")
-n13 <- ("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-cat(c(n01, n02,n03, n04, n05,
-      n06, n07, n08, n09, n10,
-      n11, n12, n13))
-
-library(data.table)
-library(robspack)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14:15
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'tower-insitu',
-#               verbose = T)\n")
-# df <- obs_read(index = index,
-#               categories = "tower-insitu",
-#               verbose = T)
-df <- fread(ind)
-
-# 3 Filtering space and time ####
-cat("Filterinig\n")
[…]

### 2024-03-05 README.Rmd
rename robspack to rtorf

@@
 fwrite(master, 
-       file =  "paper/tower_insitu_2020.txt",
+       file =  paste0(tempdir(), "/tower_insitu_2020.txt"),
        sep = " ")
@@
 ```{r, eval = T}
+csvy <- paste0(tempdir(), "/tower_insitu_2020.csvy")
 obs_write_csvy(dt = master,
                notes = "tower 2020",
-          out = "paper/tower_insitu_2020.csvy")
+               out = csvy
+)
 ```
@@
 ```{r, eval = F}
-readLines("paper/tower_insitu_2020.csvy")[1:38]
+readLines(csvy)[1:38]
 ```
@@
   fwrite(x = receptor_agl,
-         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
+         file = paste0(tempdir(), "/receptor_tower_insitu_2020_AGL.txt"),
          sep = " ")}
@@
   fwrite(x = receptor_asl,
-         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
+         file = paste0(tempdir(), "/receptor_tower_insitu_2020_ASL.txt"),
          sep = " ")}

### 2024-03-05 rscripts/r/inputs_inv.R

@@ -1,0 +1,310 @@
+
+library(colorout)
+library(data.table)
+library(robspack)
+
+dir.create("invfiles")
+
+# aircraft insitu ####
+ma1 <- paste0("/path/master/master_aircraft_insitu_", 2017:2021, ".csv")
+na <- gsub(pattern = ".csv", replacement = "", x = ma1)
+na <- gsub(pattern = "master/master_", replacement = "", x = na)
+
+lx <- list()
+
+for(i in seq_along(ma1)) {
+  if(!file.exists(ma1[i])) {
+    cat("skipping ", ma1[i], "\n")
+    next
+  }
+  dx <- fread(ma1[i])
+  dx$file <- na[i]
+  print(dim(dx))
+  lx[[i]] <- dx
+}
+master <- obs_list.dt(lx)
+# filtering between dec 2017 and jan 2021
+
+master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
+                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
+
+master$nc <- obs_footname(year = master$year,
+                          month = master$month,
+                          day = master$day,
+                          hour = master$hour,
+                          minute = master$minute,
+                          lat = master$latitude,
+                          lon = master$longitude,
+                          alt = master$altitude_final)
+
+path <-  "/path/aircraft/flask/" # param
+
+master$fnc <- paste0(path, master$nc)
+
+master$is_fnc <- file.exists(master$fnc)
+
+table(master$is_fnc)
+
+master <- master[is_fnc == TRUE]
+
+master <- master[!is.na(value)]
+
+master <- master[altitude_final > 10]
+
+dfaircraft <- obs_invfiles(master = master,
+                           path = path,
+                           Type = "continuous",
+                           SubType = "aircraft")
+obs_plot(dfaircraft$master,
+         time = "timeUTC",
+         y = "value",
+         colu = "site_code",
+         yfactor = 1e9,
+         cex = 0.5,
+         type = "l"
+)
+
+
+# tower insitu ####
+ma1 <- paste0("master/master_tower_insitu_", 2017:2021, ".csv")
+na <- gsub(pattern = ".csv", replacement = "", x = ma1)
+na <- gsub(pattern = "master/master_", replacement = "", x = na)
+
+lx <- list()
+
+for(i in seq_along(ma1)) {
+  if(!file.exists(ma1[i])) {
+    cat("skipping ", ma1[i], "\n")
+    next
+  }
[…]

### 2024-03-05 rscripts/r/surface_insitu_year.R

@@ -1,0 +1,236 @@
+#example Rscript rscripts/surface_year.R 2020 obs/obs_surface-insitu.csv
+
+#!/usr/bin/env Rscript
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) != 2) {
+  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
+} else {
+  yy <-  as.numeric(args[1] )
+  print(yy)
+
+  ind  <- as.character(args[2])
+  print(ind)
+
+}
+out <- paste0("master/master_surface_insitu_", yy)
+
+
+message("All missing values are filled with 999999")
+
+cat("                 _     robspackfilter              \n")
+cat("               (`  ).                  _           \n")
+cat("              (     ).             .:(`  )`.       \n")
+cat(")           _(       '`.          :(   .    )      \n")
+cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
+cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
+cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
+cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
+cat(")  )  ( )       --'       `- __.'         :(      )) \n")
+cat(".-'  (_.'          .')                    `(    )  ))\n")
+cat("                  (_  )                     ` __.:'          \n")
+cat("\n")
+cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
+
+library(data.table)
+library(robspack)
+
+# 0 filters ####
+north <- 80
+south <- 10
+west <- -170
+east <- -50
+max_altitude <- 8000
+evening <- 14:15
+
+# 1 Index ####
+#cat("Running obs_index(obs)\n")
+#index <- obs_index(obs = obs)
+
+# 2 Read ####
+#cat("Running obs_read(index = index,
+#               categories = 'surface-insitu',
+#               verbose = T)\n")
+#df <- obs_read(index = index,
+#               categories = "surface-insitu",
+#               verbose = T)
+
+df <- fread(ind)
+
+# 3 Filtering space and time ####
+cat("Filtering\n")
+df <- rbind(df[year == yy - 1 & month == 12],
+            df[year == yy],
+            df[year == yy + 1 & month == 1])
+
+df <- df[altitude_final < max_altitude &
+           latitude < north &
+           latitude > south &
[…]

### 2024-03-06 README.Rmd
rename robspack to rtorf
add actions

@@
 activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/rtorf)
-[![R build
-status](https://github.com/ibarraespinosa/rtorf/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
+
 [![R build
 status](https://github.com/ibarraespinosa/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
+
+[![R-CMD-check](https://github.com/ibarraespinosa/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ibarraespinosa/torf/actions/workflows/R-CMD-check.yaml)
 

### 2024-03-06 paper_elsevier/paper_elsevier.Rmd
robspack to rtorf

@@
 ---
-title: robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
+title: rtorf An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
 author:
@@
 abstract: |
-  In this study, we present a new open-source R package `robspack`, 
+  In this study, we present a new open-source R package `rtorf`, 
   to read, process, select, and plot NOAA Observation Package (ObsPack)
@@
   Python, Julia and R implementations for data-science, 
-  providing a strong basis for robspack. robspack provides functions 
+  providing a strong basis for rtorf. rtorf provides functions 
   to perform these tasks in a transparent and efficient way, 
@@
 can be cumbersome.
-The `robspack` package provides the 
+The `rtorf` package provides the 
 GHG science and research community a transparent and efficient tool to 
@@
 
-In this manuscript we present `robspack`, an R package to read,
+In this manuscript we present `rtorf`, an R package to read,
 process and plot NOAA ObsPack data. For this release, we are 
@@
 
-To install `robspack`, the user must have installed the R package 
+To install `rtorf`, the user must have installed the R package 
 remotes and run the following script. This process will install 
@@
 ```{r install, eval = F}
-remotes::install_github("ibarraespinosa/robspack")
-library(robspack)
+remotes::install_github("ibarraespinosa/rtorf")
+library(rtorf)
 ```
@@
 
-`robspack` is a collection of function organized together to read and 
+`rtorf` is a collection of function organized together to read and 
 process ObsPack files [@masarie2014obspack]. The general process 
@@
 
-https://github.com/ibarraespinosa/robspack/tree/main/rscripts
+https://github.com/ibarraespinosa/rtorf/tree/main/rscripts
 
@@
 ```{r libs, include = F, message=F, warning=F}
-library(robspack)
+library(rtorf)
 library(data.table)
@@
 
-The `robspack` functions are shown in table 1.
+The `rtorf` functions are shown in table 1.
 
-Table 1. Functions and classes in `robspack`.
+Table 1. Functions and classes in `rtorf`.
 
@@
 Monitoring Laboratory (GML) from the National Oceanographic and 
-Atmospheric Administration (NOAA). robspack reads the text data which 
+Atmospheric Administration (NOAA). rtorf reads the text data which 
 have different headers and organizes them in a common format. Then, 

### 2024-03-06 paper_elsevier/paper_elsevier.tex
robspack to rtorf

@@
             pdfauthor={},
-            pdftitle={robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack},
+            pdftitle={rtorf An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack},
             colorlinks=false,
@@
 
-  \title{robspack An R package for processing atmospheric greenhouse gas
+  \title{rtorf An R package for processing atmospheric greenhouse gas
 observations in NOAA's ObsPack}
@@
   In this study, we present a new open-source R package
-  \texttt{robspack}, to read, process, select, and plot NOAA Observation
+  \texttt{rtorf}, to read, process, select, and plot NOAA Observation
   Package (ObsPack) data products. We use a methane ObsPack data product
@@
   data.table is faster than other Python, Julia and R implementations
-  for data-science, providing a strong basis for robspack. robspack
+  for data-science, providing a strong basis for rtorf. rtorf
   provides functions to perform these tasks in a transparent and
@@
 Given the complexity of ObsPack format, reading and analyzing the data
-can be cumbersome. The \texttt{robspack} package provides the GHG
+can be cumbersome. The \texttt{rtorf} package provides the GHG
 science and research community a transparent and efficient tool to
@@
 
-In this manuscript we present \texttt{robspack}, an R package to read,
+In this manuscript we present \texttt{rtorf}, an R package to read,
 process and plot NOAA ObsPack data. For this release, we are focued on
@@
 
-To install \texttt{robspack}, the user must have installed the R package
+To install \texttt{rtorf}, the user must have installed the R package
 remotes and run the following script. This process will install all the
@@
 \begin{Highlighting}[]
-\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"ibarraespinosa/robspack"}\NormalTok{)}
-\FunctionTok{library}\NormalTok{(robspack)}
+\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"ibarraespinosa/rtorf"}\NormalTok{)}
+\FunctionTok{library}\NormalTok{(rtorf)}
 \end{Highlighting}
@@
 
-\texttt{robspack} is a collection of function organized together to read
+\texttt{rtorf} is a collection of function organized together to read
 and process ObsPack files \citep{masarie2014obspack}. The general
@@
 
-The \texttt{robspack} functions are shown in table 1.
+The \texttt{rtorf} functions are shown in table 1.
 
-Table 1. Functions and classes in \texttt{robspack}.
+Table 1. Functions and classes in \texttt{rtorf}.
 
@@
 Laboratory (GML) from the National Oceanographic and Atmospheric
-Administration (NOAA). robspack reads the text data which have different
+Administration (NOAA). rtorf reads the text data which have different
 headers and organizes them in a common format. Then, this software

### 2024-03-06 rscripts/README.Rmd
robspack to rtorf

--- 2023-06-12 rscripts/00_README.Rmd
@@
 The directory `rscripts` contains the required files
-to generate the list of receptors using `robspack`
+to generate the list of receptors using `rtorf`
 The scripts will run from a defined working directory

### 2024-03-06 R/01_summary_obs.R

@@ -20,7 +20,7 @@
 #' @examples {
 #' \dontrun{
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' }
 #' }

### 2024-03-06 R/02_read_obs.R

@@ -48,7 +48,7 @@
 #'
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' dt <- obs_read(index)
 #' }
@@ -325,7 +325,7 @@ obs_read <- function(index,
 #'
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' dt <- obs_read(index)
 #' }

### 2024-03-06 R/03_addtime_obs.R

@@ -18,7 +18,7 @@
 #' @export
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <-  obs_summary(obs)
 #' dt <- obs_read(index)
 #' dt <- obs_addtime(dt)
@@ -150,7 +150,7 @@ obs_addltime <- function(dt,
 #' @export
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <-  obs_summary(obs)
 #' dt <- obs_read(index)
 #' dt <- obs_addtime(dt)

### 2024-03-06 R/07_meta_obs.R

@@ -20,7 +20,7 @@
 #' @export
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' dt <- obs_meta(index)
 #' }

### 2024-03-06 R/08_plot.R

@@ -24,7 +24,7 @@
 #' @export
 #' @examples \dontrun{
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' dt <- obs_read(index)
 #' obs_plot(dt, time = "time")

### 2024-03-06 R/09_obs_table.R

@@ -32,7 +32,7 @@
 #' then `obs_table` will calculate stats for altitude_final.
 #' @examples {
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs, "aircraft-flask")
 #' dx <- obs_table(index, categories = "aircraft-flask")
 #' }

### 2024-03-06 R/robspack-deprecated.R

@@ -1,10 +1,10 @@
-## robspack-deprecated.r
-#' @title Deprecated functions in package \pkg{robspack}.
+## rtorf-deprecated.r
+#' @title Deprecated functions in package \pkg{rtorf}.
 #' @description The functions listed below are deprecated and will be defunct in
 #'   the near future. When possible, alternative functions with similar
 #'   functionality are also mentioned. Help pages for deprecated functions are
 #'   available at \code{help("-deprecated")}.
-#' @name robspack-deprecated
+#' @name rtorf-deprecated
 #' @keywords internal
 NULL
 
@@ -32,7 +32,7 @@ NULL
 #' @examples {
 #' \dontrun{
 #' # Do not run
-#' obs <- system.file("data-raw", package = "robspack")
+#' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' }
 #' }
@@ -123,8 +123,8 @@ obs_index <- function(obs,
 #' @family helpers
 #' @description transform integer to character and add zero at left
 #' @param x integer
-#' @rdname robspack-deprecated
-#' @seealso \code{\link{robspack-deprecated}}
+#' @rdname rtorf-deprecated
+#' @seealso \code{\link{rtorf-deprecated}}
 #' @export
 #' @examples {\dontrun{
 #' #do not run

### 2024-03-06 rscripts/README.R

@@ -1,7 +1,7 @@
 
 # README ####
 # The directory `rscripts` contains the required files
-# to generate the list of receptors using `robspack`
+# to generate the list of receptors using `rtorf`
 
 # The scripts will run from a defined working directory
 

### 2024-03-06 rscripts/r/aircore_year.R

@@ -18,7 +18,7 @@ out <- paste0("master/master_aircore", yy)
 
 message("All missing values are filled with 999999")
 
-cat("                 _     robspackfilter              \n")
+cat("                 _     rtorffilter              \n")
 cat("               (`  ).                  _           \n")
 cat("              (     ).             .:(`  )`.       \n")
 cat(")           _(       '`.          :(   .    )      \n")
@@ -33,7 +33,7 @@ cat("\n")
 cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 # 0 filters ####
 north <- 80

### 2024-03-06 rscripts/r/aircraft_year.R

@@ -18,7 +18,7 @@ out <- paste0("master/master_aircraft_insitu_", yy)
 
 message("All missing values are filled with 999999")
 
-cat("                 _     robspackfilter              \n")
+cat("                 _     rtorffilter              \n")
 cat("               (`  ).                  _           \n")
 cat("              (     ).             .:(`  )`.       \n")
 cat(")           _(       '`.          :(   .    )      \n")
@@ -33,7 +33,7 @@ cat("\n")
 cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 # 0 filters ####
 north <- 80

### 2024-03-06 rscripts/r/flask_non_noaa_year.R

@@ -19,7 +19,7 @@ out <- paste0("master/master_flask_no_noaa_", yy)
 
 message("All missing values are filled with 999999")
 
-cat("                 _     robspackfilter              \n")
+cat("                 _     rtorffilter              \n")
 cat("               (`  ).                  _           \n")
 cat("              (     ).             .:(`  )`.       \n")
 cat(")           _(       '`.          :(   .    )      \n")
@@ -34,7 +34,7 @@ cat("\n")
 cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 # 0 filters ####
 north <- 80

### 2024-03-06 rscripts/r/index.R

@@ -7,7 +7,7 @@ setwd(wd)
 cat("This is your working directory", getwd(), "\n")
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 categories <- c(
   #"aircraft-pfp",

### 2024-03-06 rscripts/r/inputs_inv.R

@@ -1,7 +1,7 @@
 
 library(colorout)
 library(data.table)
-library(robspack)
+library(rtorf)
 
 dir.create("invfiles")
 

### 2024-03-06 rscripts/r/inputs_invs.R

@@ -1,7 +1,7 @@
 
 library(colorout)
 library(data.table)
-library(robspackfilter)
+library(rtorffilter)
 
 # dir.create("invfiles")
 

### 2024-03-06 rscripts/r/surface_insitu_year.R

@@ -18,7 +18,7 @@ out <- paste0("master/master_surface_insitu_", yy)
 
 message("All missing values are filled with 999999")
 
-cat("                 _     robspackfilter              \n")
+cat("                 _     rtorffilter              \n")
 cat("               (`  ).                  _           \n")
 cat("              (     ).             .:(`  )`.       \n")
 cat(")           _(       '`.          :(   .    )      \n")
@@ -33,7 +33,7 @@ cat("\n")
 cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 # 0 filters ####
 north <- 80

### 2024-03-06 rscripts/r/tower_insitu_year.R

@@ -18,7 +18,7 @@ out <- paste0("master/master_tower_insitu_", yy)
 
 message("All missing values are filled with 999999")
 
-n01 <- ("                  _     robspackfilter              \n")
+n01 <- ("                  _     rtorffilter              \n")
 n02 <- ("               (`  ).                  _           \n")
 n03 <- ("              (     ).             .:(`  )`.       \n")
 n04 <- (")           _(       '`.          :(   .    )      \n")
@@ -36,7 +36,7 @@ cat(c(n01, n02,n03, n04, n05,
       n11, n12, n13))
 
 library(data.table)
-library(robspack)
+library(rtorf)
 
 # 0 filters ####
 north <- 80

### 2024-03-07 R/01_summary_obs.R
order index

@@ -47,6 +47,8 @@ obs_summary <- function(obs,
   # create index ####
   # the idea
   index <- data.table::data.table(id = x, name = na)
+  data.table::setorderv(index, "name")
+
   index$n <- 1:nrow(index)
   sector <- NULL
   for(i in seq_along(categories)) {

### 2024-03-18 README.Rmd
order index
readme

@@
 ![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/rtorf)
+activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
 
@@
 
-[![R-CMD-check](https://github.com/ibarraespinosa/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ibarraespinosa/torf/actions/workflows/R-CMD-check.yaml)
+[![R-CMD-check](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml)
 
@@
 ```{r, eval = F}
-remotes::install_github("ibarraespinosa/rtorf")
+remotes::install_github("noaa-gml/rtorf")
 ```
@@
 
-```{r}
+## This exampel reads text files but it will be updated to read NetCDF
+
+```{r, eval = F}
 
@@
 index <- obs_summary(obs = obs)
+
+```
+
+```{r, eval = T, include = F}
+
+categories <- c(
+  "aircraft-pfp",
+  "aircraft-insitu",
+  "surface-insitu",
+  "aircore",
+  "surface-pfp",
+  "tower-insitu",
+  "shipboard-insitu",
+  "flask"
+)
+obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/txt/"
+index <- obs_summary(obs = obs, verbose = T)
 
@@
 
-## Recommendation for other applications
-The approach to generate receptors depends on each type of 
-observation and other considerations. For instance, aircraft with 
-continuous observations at each second can be filtered and averaged 
-every 20 seconds. In that way, the footprints are still 
-representative and it would not be necessary to run HYSPLIT 
-every second. Of course, it depends on the application 
-and objective of the study. For this manuscript, we are 
-presenting the generation of receptors based on tower observations.
-
 ## Application for other sectors

### 2024-03-25 README.Rmd
readme
readme mentioning python

@@
 
-## This exampel reads text files but it will be updated to read NetCDF
+>  This example reads text files but it will be updated to read NetCDF
 
@@
 `https://github.com/ibarraespinosa/rtorf/tree/main/rscripts`
+
+## Implementation in python:
+
+I'm currently implementing a verison in python:
+
+```{r, include=F}
+data.frame(
+  stringsAsFactors = FALSE,
+                 R = c("fex","invfile",
+                       "obs_addltime","obs_addstime","obs_addtime","obs_agg",
+                       "obs_find_receptors","obs_footname","obs_format","obs_freq",
+                       "obs_index","obs_invfiles","obs_list.dt","obs_meta",
+                       "obs_out","obs_plot","obs_rbind","obs_read","obs_read_csvy",
+                       "obs_read_nc","obs_roundtime","obs_summary",
+                       "obs_table","obs_trunc","obs_write_csvy","plot.invfile",
+                       "print.invfile","sr","summary.invfile"),
+       description = c("File extension",
+                       "Methods for objects with class 'invfile'",
+                       "local hour (bsed on longitude and time)","Add solar time into obspack",
+                       "Add times into obspack","Aggregates ObsPack by time",
+                       "Compares expected receptors","Expected footprint name",
+                       "Formatting data","return numeric vector in intervals",
+                       "Summary of the ObsPack files (.txt)",
+                       "Generate files to perform inverse modeling","list.dt","Read obspack metadata",
+                       "outersect","Read obspack metadata","rbind obspack",
+                       "Read obspack (.txt)","reads CSVY","Read obspack (.nc)",
+                       "round seconds from \"POSIXct\" \"POSIXt\" classes",
+                       "Summary of the ObsPack files (.txt)","Obspack Table",
+                       "Trunc numbers with a desired number of decimals",
+                       "Generates YAML and write data.frame",
+                       "Methods for objects with class 'invfile'","Methods for objects with class 'invfile'",
+                       "Extacts n last characters",
+                       "Methods for objects with class 'invfile"),
+            Python = c("","","","","","","","",
+                       "","","OK","","","","","","","","","OK","",
+                       "","","","","","","","")
+) -> df
+
+```
+
+```{r, eval=TRUE, echo=FALSE}
+knitr::kable(df)
+
+```

### 2024-04-24 examples/README_nc.Rmd
documentation and details ins NEWS.md

--- 2024-03-25 README.Rmd
@@
 
-```{r, eval = F}
+```{r, eval = T}
 
-categories <- c(
-  "aircraft-pfp",
+cate = c("aircraft-pfp",
          "aircraft-insitu",
+         "aircraft-flask",
          "surface-insitu",
-  "aircore",
+         "surface-flask", 
          "surface-pfp",   
          "tower-insitu",  
+         "aircore",       
          "shipboard-insitu",
-  "flask"
-)
-obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
-index <- obs_summary(obs = obs)
+         "shipboard-flask")
+
+obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
+index <- obs_summary(obs = obs, categories = cate)
 
@@
 
-```{r, eval = T, include = F}
-
-categories <- c(
-  "aircraft-pfp",
-  "aircraft-insitu",
-  "surface-insitu",
-  "aircore",
-  "surface-pfp",
-  "tower-insitu",
-  "shipboard-insitu",
-  "flask"
-)
-obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/txt/"
-index <- obs_summary(obs = obs, verbose = T)
-
-```
-
-There are 362 files in the ObsPack directory.
+There are `r nrow(index)` files in the ObsPack directory.
 The printed information also shows the total at the bottom,
@@
 furthermore, the printed information
-also shows that there are 136 files with the
+also shows that there are 109 files with the
 `agl` explicitly mentioned in the name of the file.
-
-Sometimes we need more information about the site. For instance,
-what do the observations start and end. Then, we added 
-the function `obs_table`, which calculates statistics
-summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode. For instance,
-the observations in the site "SCT" in
-South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
-"2020-12-31 23:30:00 UTC".
-```{r}
-dft <- obs_table(index = index,
-                 categories = "tower-insitu",
-                 verbose = FALSE)
-dft[site_code == "SCT", ]$timeUTC |> 
-  range()
-```
 
@@
 ```{r}
-df <- obs_read(index = index,
-               categories = "tower-insitu",
-               verbose = FALSE)
+df <- obs_read_nc(index = index,
+               categories = "tower-insitu")
 
+```
+
+Sometimes we need more information about the site. For instance,
+what do the observations start and end. Then, we added 
+the function `obs_table`, which calculates statistics
+summary of "time" and other numeric variables by file name,
+sector, site, altitude and mode. For instance,
+the observations in the site "SCT" in
+South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
+"2020-12-31 23:30:00 UTC".
+
+```{r}
+dft <- obs_table(df = df)
+dft[site_name %in% unique(site_name)[sample(1:length(unique(dft$site_code)), 2)]]
 ```
@@
 ```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
-obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
+obs_plot(dt = df, 
+         time = "time", 
+         yfactor = 1e9, 
+         cex = 0.5)
 ```
@@
 max_altitude <- 8000
-evening <- 14:15
+evening <- 14
 
@@
 
+if(any(names(df) %in% "intake_height")) {
+  df <- df[intake_height < max_altitude &
+           latitude < north &
+           latitude > south &
+           longitude < east &
+           longitude > west]
+} else {
   df <- df[altitude_final < max_altitude &
@@
            longitude > west]
+}
+
+unique(df[, c("intake_height", "site_elevation", "elevation",
+              "dataset_selection_tag",
+              "site_name")])
+
 ```
@@
 ```{r}
+
+if(any(names(df) %in% "intake_height")) {
+dfa <- df[,
+          max(intake_height),
+          by = site_code] |> unique()
+
+names(dfa)[2] <- "max_altitude"
+} else {
 dfa <- df[,
@@
 names(dfa)[2] <- "max_altitude"
+}
+dfa
 ```
@@
 df2 <- obs_addtime(df)
-df2$timeUTC <- cut(x = df2$timeUTC+3600,
-                   breaks = "2 hour") |>
-  as.character() |>
-  as.POSIXct(tz = "UTC")
-df3 <- obs_addltime(df2)
-df3 <- df3[lh %in% evening]
+```
+
+"_st".
+
+```{r}
+df2 <- obs_addstime(df2)
+
+```
+solar time is not exact, then I need to cut every hour and aggregate observations
+by hour. At the end I want one daily observation per day and site
+
+# TODO ...
+```{r}
+df2 <- df2[hour_st %in% evening]
+
 ```
@@
 
-I'm currently implementing a verison in python:
+I'm currently implementing a version in python:
 

### 2024-04-24 examples/README_text.Rmd
documentation and details ins NEWS.md

--- 2024-03-25 README.Rmd
@@
 
->  This example reads text files but it will be updated to read NetCDF
-
-```{r, eval = F}
-
-categories <- c(
-  "aircraft-pfp",
-  "aircraft-insitu",
-  "surface-insitu",
-  "aircore",
-  "surface-pfp",
-  "tower-insitu",
-  "shipboard-insitu",
-  "flask"
-)
-obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
-index <- obs_summary(obs = obs)
-
-```
+>  This example reads text files but it will be updated to read text file
 
@@
 
-Sometimes we need more information about the site. For instance,
-what do the observations start and end. Then, we added 
-the function `obs_table`, which calculates statistics
-summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode. For instance,
-the observations in the site "SCT" in
-South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
-"2020-12-31 23:30:00 UTC".
-```{r}
-dft <- obs_table(index = index,
-                 categories = "tower-insitu",
-                 verbose = FALSE)
-dft[site_code == "SCT", ]$timeUTC |> 
-  range()
-```
-
 ### Read data
@@
                verbose = FALSE)
+df[1:5]
 
+```
+
+Sometimes we need more information about the site. For instance,
+what do the observations start and end. Then, we added 
+the function `obs_table`, which calculates statistics
+summary of "time" and other numeric variables by file name,
+sector, site, altitude and mode. For instance,
+the observations in the site "SCT" in
+South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
+"2020-12-31 23:30:00 UTC".
+
+```{r}
+dft <- obs_table(df = df)
+dft[site_name %in% unique(site_name)[1:2]]
+# dft[site_code == "SCT", ]$timeUTC |> 
+# range()
 ```
@@
            longitude > west]
+}
+
 ```
@@
 ```{r}
-dfa <- df[,
-          max(altitude_final),
-          by = site_code] |> unique()
+
+df5[,
+    max_altitude := max(altitude_final),
+    by = site_code]
 
@@
 df2 <- obs_addtime(df)
+```
+
+```{r}
+
 df2$timeUTC <- cut(x = df2$timeUTC+3600,
@@
   as.POSIXct(tz = "UTC")
+
+```
+
+```{r}
 df3 <- obs_addltime(df2)
+
+```
+
+```{r}
 df3 <- df3[lh %in% evening]
+
 ```

### 2024-04-24 R/02_read_obs.R

@@ -13,6 +13,8 @@
 #' "flask").
 #' @param verbose Logical to show more information
 #' @param n_site_code number of characters extratced from metadata after search
+#' @param n_site_latitude number of characters extracted from metadata after search
+#' @param n_site_longitude number of characters extracted from metadata after search
 #' @param n_site_name number of characters extracted from metadata after search
 #' @param n_site_country number of characters extracted from metadata after search
 #' @param n_dataset_project number of characters extracted from metadata after search
@@ -56,6 +58,8 @@ obs_read <- function(index,
                      categories = "flask",
                      verbose = TRUE,
                      n_site_code = 15,
+                     n_site_latitude = 18,
+                     n_site_longitude = 19,
                      n_site_name = 15,
                      n_site_country = 18,
                      n_dataset_project = 21,
@@ -104,17 +108,34 @@ obs_read <- function(index,
                     x = att,
                     value = T)
 
-    # everything after char 15
     site_name <- substr(x = pattern,
                         start = n_site_name,
                         stop = nchar(pattern))
+
+
+    # site_latitude ####
+    pattern <- grep(pattern = "site_latitude",
+                    x = att,
+                    value = T)[1]
+
+    (site_latitude <- substr(x = pattern,
+                             start = n_site_latitude,
+                             stop = nchar(pattern)))
+
+    # site_longitude ####
+    pattern <- grep(pattern = "site_longitude",
+                    x = att,
+                    value = T)[1]
+
+    (site_longitude <- substr(x = pattern,
+                              start = n_site_longitude,
+                              stop = nchar(pattern)))
 
     # site_country ####
     pattern <- grep(pattern = "site_country",
                     x = att,
                     value = T)[1]
 
-    # everything after char 18
     site_country <- substr(x = pattern,
                            start = n_site_country,
                            stop = nchar(pattern))
@@ -124,7 +145,6 @@ obs_read <- function(index,
                     x = att,
                     value = T)[1]
 
-    # everything after char 21
     dataset_project <- substr(x = pattern,
                               start = n_dataset_project,
                               stop = nchar(pattern))
@@ -134,10 +154,10 @@ obs_read <- function(index,
                     x = att,
[…]

### 2024-04-24 R/09_obs_table.R

@@ -1,4 +1,4 @@
-#' Obspack Table
+#' Obspack Table Summary
 #'
 #' This function reads the obsPack directory
 #' providing a summary for the columns: "value", "time",
@@ -7,321 +7,40 @@
 #' "site_country", "type_altitude", "lab_1_abbr" and
 #' "site_utc2lst"
 #'
-#' @param index data.table
-#' @param categories character; ONE category
-#' : of c("aircraft-pfp", "aircraft-insitu",
-#' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
-#' "flask").
-#' @param verbose Logical to show more information
-#' @param n_site_code number of characters extraced from metadata after search
-#' @param n_site_lat number of characters for latitude
-#' @param n_site_lon number of characters for longitude
-#' @param n_site_name number of characters extraced from metadata after search
-#' @param n_site_country number of characters extraced from metadata after search
-#' @param n_dataset_project number of characters extraced from metadata after search
-#' @param n_lab number of characters extraced from metadata after search
-#' @param n_scales number of characters extraced from metadata after search
-#' @param n_site_elevation number of characters extraced from metadata after search
-#' @param n_altitude_comment number of characters extraced from metadata after search
-#' @param n_utc number of characters extraced from metadata after search
-#' @param fill_value fill value. Appeared in aoa_aircraft-flask_19_allvalid.txt
+#' @param df data.table
+#' @param cols String of columns to be summarized.
 #' @return A data.frame with with an index obspack.
 #' @importFrom data.table fwrite ".N" ":=" rbindlist "%chin%"
 #' @export
-#' @note Categories are defined in `obs_summary`. If aircraft* is used,
-#' then `obs_table` will calculate stats for altitude_final.
+
 #' @examples {
 #' # Do not run
 #' obs <- system.file("data-raw", package = "rtorf")
-#' index <- obs_summary(obs, "aircraft-flask")
-#' dx <- obs_table(index, categories = "aircraft-flask")
+#' index <- obs_summary(obs)
+#' dt <- obs_read(index)
+#' dx <- obs_table(dt)
 #' }
-obs_table <- function(index,
-                      categories = "tower-insitu",
-                      verbose = FALSE,
-                      n_site_code = 15,
-                      n_site_lat = 18,
-                      n_site_lon = 19,
-                      n_site_name = 15,
-                      n_site_country = 18,
-                      n_dataset_project = 21,
-                      n_lab = 16,
[…]

### 2024-04-25 examples/README_nc.Rmd
documentation and change in obs_agg

@@
 summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode. For instance,
-the observations in the site "SCT" in
-South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
-"2020-12-31 23:30:00 UTC".
+sector, site, altitude and mode.
 
@@
 dft <- obs_table(df = df)
-dft[site_name %in% unique(site_name)[sample(1:length(unique(dft$site_code)), 2)]]
+```
+
+Now we can visualize the average of observations by site.
+
+```{r sf, echo = TRUE, message=FALSE}
+library(sf)
+sdft <- st_as_sf(dft[stat == "mean"], 
+                 coords = c("longitude", "latitude"), 
+                 crs = 4326)
+sdft$value <- sdft$value*1e09
+plot(sdft["value"], 
+     reset = FALSE, 
+     axes = TRUE, 
+     graticule = TRUE,
+     pch = 16, 
+     cex = 1.5, 
+     pal = cptcity::lucky(colorRampPalette = T), 
+     main = NULL)
+maps::map(add = T)
+```
+
+Let us randomly select a couples of sites from the database.
+
+```{r}
+usites <- unique(df$site_name)[sample(seq_along(unique(df$site_name)), 2)]
+usites
+```
+
+```{r}
+knitr::kable(dft[site_name %in% usites])
 ```
@@
 ```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
-obs_plot(dt = df, 
+obs_plot(dt = df[site_name %in% usites], 
          time = "time", 
@@
 
-"_st".
-
-```{r}
-df2 <- obs_addstime(df2)
-
-```
-solar time is not exact, then I need to cut every hour and aggregate observations
-by hour. At the end I want one daily observation per day and site
-
-# TODO ...
 ```{r}
@@
 
-Now there are `r nrow(df3)` observations.
+Now there are `r nrow(df2)` observations.
 At this point we can calculate the averages of several columns by
@@
 in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`, in this case
-the function mean by time and altitude. Finally, we
-add local time again.
+establish the function used to aggregate `cols`. I need to aggregate the data
+by date (year, month, date), because it is already filtered by
+the hours of interest. Then, I would have 1 observation per day.
 
 ```{r}
-df4 <- obs_agg(dt = df3,
+df3 <- obs_agg(dt = df2,
                gby = "mean",

### 2024-04-25 R/03_addtime_obs.R

@@ -7,6 +7,7 @@
 #' @family time
 #' @param dt obspack data.table
 #' @param verbose obspack data.table
+#' @param tz Timezone, default "UTC"
 #' @return A data.frame with with an index obspack.
 #' @importFrom data.table year month hour minute second setDT fifelse uniqueN
 #' @note timeUTC is calculated based on the field column start_time,
@@ -24,7 +25,8 @@
 #' dt <- obs_addtime(dt)
 #' }
 obs_addtime <- function(dt,
-                        verbose = TRUE){
+                        verbose = TRUE,
+                        tz = "UTC"){
 
   if(!inherits(dt, "data.table")) data.table::setDT(dt)
 
@@ -38,7 +40,7 @@ obs_addtime <- function(dt,
   # Time is already UTC
   dt$timeUTC_start <- as.POSIXct(dt$start_time,
                                  origin = "1970-01-01",
-                                 tz = "UTC")
+                                 tz = tz)
 
   if(verbose) cat("Adding timeUTC_end\n")
   # adding timeUTC_end
@@ -98,6 +100,7 @@ obs_addtime <- function(dt,
 #' @param utc2lt Character indicating the integer column to convert to local time
 #' if available
 #' @param longitude Character indicating the column with the lingitude
+#' @param tz Timezone, default "UTC"
 #' @importFrom data.table hour
 #' @return data.table with local time columns
 #' @note time depending n longitude is by John Miller (GML/NOAA)
@@ -110,7 +113,8 @@ obs_addtime <- function(dt,
 obs_addltime <- function(dt,
                          timeUTC = "timeUTC",
                          utc2lt = "site_utc2lst",
-                         longitude = "longitude") {
+                         longitude = "longitude",
+                         tz = "UTC") {
 
   # Priority site_utc2lst and if it is not available,
   if(any(utc2lt == names(dt))) {
@@ -128,7 +132,7 @@ obs_addltime <- function(dt,
   if(inherits(dt$local_time, "numeric")) {
     dt$local_time <- as.POSIXct(dt$local_time,
                                 origin = "1970-01-01",
-                                tz = "UTC")
+                                tz = tz)
   }
 
   dt$lh <- data.table::hour(dt$local_time)
@@ -143,7 +147,7 @@ obs_addltime <- function(dt,
 #'
 #' @family time
 #' @param dt obspack data.table
-#' @param verbose obspack data.table
+#' @param tz Timezone, default "UTC"
[…]

### 2024-04-25 R/04_aggregate_obs.R

@@ -10,39 +10,41 @@
 #' data. Default is "mean" meaning that all numeric variables
 #' will be averaged by the new_second column established in the `idcol`
 #' @param cols Character which defines columns to be aggregated
+#' @param time Character to aggregate time (solar, local, UTC)
 #' @param byalt Logical, to aggregate by altitude or not (used in tower)
 #' @param verbose logical to show more information
 #' @return A data.frame with with an index obspack.
 #' @importFrom data.table  second setDT setorderv mday
 #' @importFrom lubridate decimal_date
-
+#' @note By default add column timeUTC with parameters from time loca, solar or UTC
 #' @export
 #' @examples \dontrun{
 #' # Do not run
 #' dt <- obs_addtime(dt)
 #' }
 obs_agg <- function(dt,
-                     gby = "mean",
-                     cols = c("year",
-                               "month",
-                               "day",
-                               "hour",
-                               "minute",
-                               "second",
-                               "time",
-                               "time_decimal",
-                               "value",
-                               "latitude",
-                               "longitude",
-                               "altitude",
-                               "pressure",
-                               "u",
-                               "v",
-                               "temperature",
-                               "type_altitude",
-                               "altitude_final"),
-                     byalt = FALSE,
-                     verbose = TRUE){
+                    gby = "mean",
+                    cols = c("year",
+                             "month",
+                             "day",
+                             "hour",
+                             "minute",
+                             "second",
+                             "time",
+                             "time_decimal",
+                             "value",
+                             "latitude",
+                             "longitude",
+                             "altitude",
+                             "pressure",
+                             "u",
+                             "v",
+                             "temperature",
+                             "type_altitude",
+                             "altitude_final"),
+                    time = c("UTC", "solar", "local"),
+                    byalt = FALSE,
+                    verbose = TRUE){
 
   if(!inherits(dt, "data.table")){
     if(verbose) cat("Transforming your data into data.table\n")
@@ -64,45 +66,252 @@ obs_agg <- function(dt,
   timeUTC <- NULL
   lab_1_abbr <- NULL
   dataset_calibration_scale <- NULL
+  year_st <- NULL
+  month_st  <- NULL
+  day_st  <- NULL
+  intake_height  <- NULL
+  local_time  <- NULL
+  day <- NULL
[…]

### 2024-05-01 README.Rmd
typo
update obs_agg. TODO, add type altitude identifier from NetCDF< prob rename intake_height to altitude_final fix #6 #5 #4

@@
 ```{r}
-df <- obs_read(index = index,
-               categories = "tower-insitu",
-               verbose = FALSE)
+df <- obs_read_nc(index = index,
+                  categories = "tower-insitu")
 
+```
+
+Sometimes we need more information about the site. For instance,
+what do the observations start and end. Then, we added 
+the function `obs_table`, which calculates statistics
+summary of "time" and other numeric variables by file name,
+sector, site, altitude and mode.
+
+```{r}
+dft <- obs_table(df = df)
+```
+
+Now we can visualize the average of observations by site.
+
+```{r sf, echo = TRUE, message=FALSE}
+library(sf)
+sdft <- st_as_sf(dft[stat == "mean"], 
+                 coords = c("longitude", "latitude"), 
+                 crs = 4326)
+sdft$value <- sdft$value*1e09
+plot(sdft["value"], 
+     reset = FALSE, 
+     axes = TRUE, 
+     graticule = TRUE,
+     pch = 16, 
+     cex = 1.5, 
+     pal = cptcity::lucky(colorRampPalette = T), 
+     main = NULL)
+maps::map(add = T)
+```
+
+Let us randomly select a couples of sites from the database.
+
+```{r}
+usites <- unique(df$site_name)[sample(seq_along(unique(df$site_name)), 2)]
+usites
+```
+
+```{r}
+knitr::kable(dft[site_name %in% usites])
 ```
@@
 ```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
-obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
+obs_plot(dt = df[site_name %in% usites], 
+         time = "time", 
+         yfactor = 1e9, 
+         cex = 0.5)
 ```
@@
 
+# TODO ADD identifier of type of altitude of intake_height
+
 ```{r}
@@
 
+if(any(names(df) %in% "intake_height")) {
+  df <- df[intake_height < max_altitude &
+             latitude < north &
+             latitude > south &
+             longitude < east &
+             longitude > west]
+} else {
   df <- df[altitude_final < max_altitude &
@@
              longitude > west]
+}
+
+unique(df[, c("intake_height", "site_elevation", "elevation",
+              "dataset_selection_tag",
+              "site_name")])
+
 ```
@@
 ```{r}
+
+if(any(names(df) %in% "intake_height")) {
+  dfa <- df[,
+            max(intake_height),
+            by = site_code] |> unique()
+
+  names(dfa)[2] <- "max_altitude"
+} else {
   dfa <- df[,
@@
   names(dfa)[2] <- "max_altitude"
+}
+dfa
 ```
 
-### Time
+### Key Time
 
@@
 which shows the end time for each observation.
+
+```{r}
+df2 <- obs_addtime(df)
+```
+Then we need a *key_time* to aggregate data.
+This can be done using UTC, solar, or local time.
+The normal approach is using afteroon solar or local time.
+
+#### Hierarchy of solar or local time
+
+1) solar time
+2) local time with columns `site_utc2lst`
+3) local time longitude
+
+1) solar time (default)
+
+Here we select the hours of interest and then
+aggregate data by year, month and day of solar time.
+In this way, we will have one information per day. however
+this approach is not appropiate for aircraft
+which are aggregated every 10 or 20 seconds. Hence
+we need to aggregate data by one time column.
+Also, this helps to generate the receptor info 
+files including hour, minute and second.
+Hence, *we need to add solar or local time column*.
+
+```{r}
+df2$solar_time <- obs_addstime(df2)
+
+```
+
+2) local time with column `site_utc2lst`
+
 Then we need to identify the local time with the function
@@
 the time difference with utc by identifying the
-metadata column "site_utc2lst". If this information
+metadata column "site_utc2lst". 
+If solar time is not available
+#now we need to cut solar time for the frequency needed. 
+As we will work with
+
+3) local time longitude
+
+If this information
 is not available, with the aircrafts for instance,
@@
 
+#### Cut time
+
+Now we have they key column time, we can cut it accordingly.
+
 ```{r}
-df2 <- obs_addtime(df)
-df2$timeUTC <- cut(x = df2$timeUTC+3600,
-                   breaks = "2 hour") |>
-  as.character() |>
-  as.POSIXct(tz = "UTC")
-df3 <- obs_addltime(df2)
-df3 <- df3[lh %in% evening]
+df2$solar_time_cut <- cut(x = df2$solar_time,
+                   breaks = "1 hour") |>
+  as.character()
+
 ```
 
-Now there are `r nrow(df3)` observations.
+How we can check the solar time and the cut solar time.
+Please note that solar_time_cut,
+the column that it will be used to aggregate data
+
+```{r}
+df2[, c("solar_time", "solar_time_cut")]
+```
+
+How we filter for the required solar time, in this case `r evening`.
+
+```{r}
+df3 <- df2[hour_st %in% evening]
+
+```
+
+Now there are `r nrow(df3)` observations and before filtering
+`r nrow(df2)`.
 At this point we can calculate the averages of several columns by
@@
 in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`, in this case
-the function mean by time and altitude. Finally, we
-add local time again.
+establish the function used to aggregate `cols`. I need to aggregate the data
+by date (year, month, date), because it is already filtered by
+the hours of interest. Then, I would have 1 observation per day.
+
+As standard, let us define `key_time` as `solar_time`. 
+The `obs_agg` function will aggregate the desired data by that column.
+
+```{r}
+df3$key_time <- df3$solar_time_cut
+```
 
@@
                         "longitude",
-                        "type_altitude",
-                        "dif_time",
+                        # "type_altitude",
                         "year_end",
                         "site_utc2lst"),
-               verbose = FALSE,
+               verbose = T,
                byalt = TRUE)
-df5 <- obs_addltime(df4)
 ```
 
-Now there are `r nrow(df5)` observations,
-`r nrow(df3) - nrow(df5)` less observations.
+Now there are `r nrow(df4)` observations,
+`r nrow(df) - nrow(df4)` less observations.
 Here we add the column `max_altitude` to identify the max altitude
@@
 ```{r}
-df5[,
+if(any(names(df) %in% "intake_height")) {
+  df4[,
+      max_altitude := max(intake_height),
+      by = site_code]
+  df4[,
+      c("site_code",
+        "intake_height",
+        "max_altitude")] |> unique()
+} else {
+  df4[,
       max_altitude := max(altitude_final),
       by = site_code]
-df5[,
+  df4[,
       c("site_code",
@@
         "max_altitude")] |> unique()
+}
+
 ```
@@
 ```{r, eval = T}
-master <- df5
+master <- df3
 master$timeUTC <- as.character(master$timeUTC)
-master$timeUTC_end <- as.character(master$timeUTC_end)
-master$local_time <- as.character(master$local_time)
-
 fwrite(master, 
@@
 ```{r, eval = F}
-readLines(csvy)[1:38]
+readLines(csvy)[1:33]
 ```
@@
 
+if(any(names(df) %in% "intake_height")) {
+receptor <- master[intake_height == max_altitude,
+                   c("site_code",
+                     "year",
+                     "month",
+                     "day",
+                     "hour", 
+                     "minute",
+                     "second",
+                     "latitude",
+                     "longitude",
+                     "intake_height",
+                     "type_altitude",
+                     "year_end",
+                     "month_end",
+                     "day_end",
+                     "hour_end",
+                     "minute_end",
+                     "second_end")]
+receptor$intake_height <- round(receptor$intake_height)
+
+} else {
 receptor <- master[altitude_final == max_altitude,
@@
                      "second_end")]
+
 receptor$altitude_final <- round(receptor$altitude_final)
+}
+
 receptor <- obs_format(receptor)

### 2024-05-01 R/04_aggregate_obs.R

@@ -10,13 +10,12 @@
 #' data. Default is "mean" meaning that all numeric variables
 #' will be averaged by the new_second column established in the `idcol`
 #' @param cols Character which defines columns to be aggregated
-#' @param time Character to aggregate time (solar, local, UTC)
 #' @param byalt Logical, to aggregate by altitude or not (used in tower)
 #' @param verbose logical to show more information
 #' @return A data.frame with with an index obspack.
-#' @importFrom data.table  second setDT setorderv mday
+#' @importFrom data.table  second minute hour month year setDT setorderv mday
 #' @importFrom lubridate decimal_date
-#' @note By default add column timeUTC with parameters from time loca, solar or UTC
+#' @note By default add column timeUTC based on the input column `key_time`
 #' @export
 #' @examples \dontrun{
 #' # Do not run
@@ -40,9 +39,7 @@ obs_agg <- function(dt,
                              "u",
                              "v",
                              "temperature",
-                             "type_altitude",
-                             "altitude_final"),
-                    time = c("UTC", "solar", "local"),
+                             "type_altitude"),
                     byalt = FALSE,
                     verbose = TRUE){
 
@@ -52,9 +49,9 @@ obs_agg <- function(dt,
   }
 
 
-  if(!any(names(dt) %in% "timeUTC")) stop("missing `timeUTC``")
-  if(!any(names(dt) %in% "site_code")) stop("missing `site_code``")
-  if(!any(names(dt) %in% "altitude_final")) stop("missing `altitude_final``")
+  if(!any(names(dt) %in% "key_time")) stop("missing `key_time`")
+  if(!any(names(dt) %in% "site_code")) stop("missing `site_code`")
+  # if(!any(names(dt) %in% "altitude_final")) stop("missing `altitude_final``")
 
   # Remember aircraft
   # Only aircraft need to be aggregated every 20 seconds
@@ -64,6 +61,7 @@ obs_agg <- function(dt,
   site_code <- NULL
   altitude_final <- NULL
   timeUTC <- NULL
+  key_time <- NULL
   lab_1_abbr <- NULL
   dataset_calibration_scale <- NULL
   year_st <- NULL
@@ -89,244 +87,75 @@ obs_agg <- function(dt,
     # Remember tower
     # A site can have different altitudes, then
     # each group must be processed differently
-
-
-    if(time == "solar") {
-
-      if(verbose) cat("Selecting time `solar`\n")
-
[…]

### 2024-05-03 examples/README_nc.Rmd
readme nc?

@@
 
-`rtorf` only depends on `data.table`, which is basically parallel C, 
+`rtorf` only depends on `data.table` and `ncdf4`, which is basically parallel C, 
 so it can be installed in any machine.
 
-## installation
+## Installation
 
@@
 
->  This example reads text files but it will be updated to read NetCDF
+>  We first define the categories
 
@@
 
+> intake_height or altitude_final
+
+The identification
+of the altitude and type is critical. The approach used here 
+consists of:
+
+1. Use intake_height
+2. Identify `agl` from the name of the tile.
+3. If `agl` is not present, `agl = altitude - site_elevation`.
+4. If there are some NA in elevation, will result some NA in `agl`
+5. A new column is added named `altitude_final` to store `agl` or `asl`
+6. Another column named `type_altitude` is added to identify `agl` as 0, or `asl` as 1.
+
+To maintain a coherent approach among text and NetCDF files, 
+we will add a column altitude_final and type_altitude
+
+```{r}
+df$altitude_final <- df$intake_height
+df$type_altitude <- ifelse(df$dataset_intake_ht_unit == "magl", 0, 1)
+```
+
 Sometimes we need more information about the site. For instance,
@@
      cex = 1.5, 
-     pal = cptcity::lucky(colorRampPalette = T), 
+     pal = cptcity::lucky(colorRampPalette = T), #random
      main = NULL)
@@
          yfactor = 1e9, 
-         cex = 0.5)
+         cex = 0.5,
+         verbose = FALSE)
 ```
@@
 for `tower-insitu`. These observations are made between 
-`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
-of the altitude and type is critical. The approach used here 
-consists of:
-
-1. Identify `agl` from the name of the tile.
-2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
-3. If `agl` is not present, `agl = altitude - elevation`.
-4. If there are some NA in elevation, will result some NA in `agl`
-5. A new column is added named `altitude_final` to store `agl` or `asl`
-6. Another column named `type_altitude` is added to identify `agl` or `asl`.
-7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"
+`r range(df$year)[1]` and `r range(df$year)[2]`.
 
@@
 
-if(any(names(df) %in% "intake_height")) {
-  df <- df[intake_height < max_altitude &
-             latitude < north &
-             latitude > south &
-             longitude < east &
-             longitude > west]
-} else {
 df <- df[altitude_final < max_altitude &
@@
            longitude > west]
-}
 
-unique(df[, c("intake_height", "site_elevation", "elevation",
+unique(df[, c("altitude_final", "site_elevation", "elevation",
               "dataset_selection_tag",
@@
 ```{r}
-
-if(any(names(df) %in% "intake_height")) {
-  dfa <- df[,
-            max(intake_height),
-            by = site_code] |> unique()
-
-  names(dfa)[2] <- "max_altitude"
-} else {
 dfa <- df[,
@@
 names(dfa)[2] <- "max_altitude"
-}
 dfa
@@
 
-### Time
+### Key Time
 
@@
 which shows the end time for each observation.
+
+```{r}
+df2 <- obs_addtime(df)
+```
+Then we need a *key_time* to aggregate data.
+This can be done using UTC, solar, or local time.
+The normal approach is using afternoon solar or local time.
+
+#### Hierarchy of solar or local time
+
+1) Solar time
+2) Local time with columns `site_utc2lst`
+3) Local time longitude
+
+> solar time (default)
+
+Here we select the hours of interest and then
+aggregate data by year, month and day of solar time.
+In this way, we will have one information per day. however
+this approach is not appropriate for aircraft
+which are aggregated every 10 or 20 seconds. Hence
+we need to aggregate data by one time column.
+Also, this helps to generate the receptor info 
+files including hour, minute and second.
+Hence, *we need to add solar or local time column*.
+
+```{r}
+df2$solar_time <- obs_addstime(df2)
+
+```
+
+> local time with column `site_utc2lst`
+
 Then we need to identify the local time with the function
@@
 the time difference with utc by identifying the
-metadata column "site_utc2lst". If this information
+metadata column "site_utc2lst". 
+If solar time is not available
+#now we need to cut solar time for the frequency needed. 
+As we will work with
+
+> local time longitude
+
+If this information
 is not available, with the aircrafts for instance,
@@
 
-```{r}
-df2 <- obs_addtime(df)
-```
+#### Cut time
+
+Now we have they key column time, we can cut it accordingly.
 
 ```{r}
-df2 <- df2[hour_st %in% evening]
+df2$solar_time_cut <- cut(x = df2$solar_time,
+                          breaks = "1 hour") |>
+  as.character()
 
@@
 
-Now there are `r nrow(df2)` observations.
+How we can check the solar time and the cut solar time.
+Please note that solar_time_cut,
+the column that it will be used to aggregate data
+
+How we filter for the required solar time, in this case `r evening`.
+
+```{r}
+df3 <- df2[hour_st %in% evening]
+df3[, c("solar_time", "solar_time_cut")]
+
+```
+
+Now there are `r nrow(df3)` observations and before filtering
+`r nrow(df2)`.
 At this point we can calculate the averages of several columns by
@@
 
+As standard, let us define `key_time` as `solar_time`. 
+The `obs_agg` function will aggregate the desired data by that column.
+
 ```{r}
-df3 <- obs_agg(dt = df2,
+df3$key_time <- df3$solar_time_cut
+```
+
+```{r}
+df4 <- obs_agg(dt = df3,
                gby = "mean",
@@
                         "longitude",
-                        "type_altitude",
-                        "dif_time",
-                        "year_end",
                         "site_utc2lst"),
-               verbose = FALSE,
+               verbose = T,
                byalt = TRUE)
-df5 <- obs_addltime(df4)
 ```
 
-Now there are `r nrow(df5)` observations,
-`r nrow(df3) - nrow(df5)` less observations.
+Now there are `r nrow(df4)` observations,
+`r nrow(df) - nrow(df4)` less observations.
 Here we add the column `max_altitude` to identify the max altitude
@@
 ```{r}
-df5[,
+df4[,
     max_altitude := max(altitude_final),
     by = site_code]
-df5[,
+df4[,
     c("site_code",
@@
       "max_altitude")] |> unique()
+
 ```
@@
 ```{r, eval = F}
-readLines(csvy)[1:38]
+readLines(csvy)[1:34] # and more
 ```
@@
                      "minute_end",
-                     "second_end")]
+                     "second_end",
+                     "time_decimal")]
+
 receptor$altitude_final <- round(receptor$altitude_final)
+
 receptor <- obs_format(receptor)

### 2024-05-03 R/04_aggregate_obs.R

@@ -51,7 +51,7 @@ obs_agg <- function(dt,
 
   if(!any(names(dt) %in% "key_time")) stop("missing `key_time`")
   if(!any(names(dt) %in% "site_code")) stop("missing `site_code`")
-  # if(!any(names(dt) %in% "altitude_final")) stop("missing `altitude_final``")
+  if(!any(names(dt) %in% "altitude_final")) stop("missing `altitude_final`")
 
   # Remember aircraft
   # Only aircraft need to be aggregated every 20 seconds
@@ -87,34 +87,17 @@ obs_agg <- function(dt,
     # Remember tower
     # A site can have different altitudes, then
     # each group must be processed differently
-    if(any(names(dt) %in% "intake_height")) { #solar + alt
-      if(verbose) cat("Identified intake_height\n")
-
-      dt <- dt[,
-               lapply(.SD,
-                      mean,
-                      na.rm = T),
-               .SDcols = cols,
-               by = list(timeUTC = key_time,
-                         site_code,
-                         intake_height,
-                         lab_1_abbr,
-                         dataset_calibration_scale)]
-
-    } else { #solar - alt
-      if(verbose) cat("Identified altitude_final\n")
-
-      dt <- dt[,
-               lapply(.SD,
-                      mean,
-                      na.rm = T),
-               .SDcols = cols,
-               by = list(timeUTC = key_time,
-                         site_code,
-                         altitude_final,
-                         lab_1_abbr,
-                         dataset_calibration_scale)]
-    }
+    dt <- dt[,
+             lapply(.SD,
+                    mean,
+                    na.rm = T),
+             .SDcols = cols,
+             by = list(timeUTC = key_time,
+                       site_code,
+                       altitude_final,
+                       type_altitude,
+                       lab_1_abbr,
+                       dataset_calibration_scale)]
   } else {
     dt <- dt[,
              lapply(.SD,
@@ -136,8 +119,8 @@ obs_agg <- function(dt,
   dt$year <- data.table::year(dt$timeUTC)
   dt$month <- data.table::month(dt$timeUTC)
   dt$day <- strftime(dt$timeUTC,
-                             "%d",
-                             tz = "UTC")
+                     "%d",
+                     tz = "UTC")
   dt$hour <- data.table::hour(dt$timeUTC)
   dt$minute <- data.table::minute(dt$timeUTC)
   dt$second <- data.table::second(dt$timeUTC)

### 2024-05-03 README.Rmd
update docs

@@
 index file previously generated.
-Here we selected the category "tower-insitu".
+Here we
+selected the category "tower-insitu".
 The argument verbose prints which files are being read each time,
@@
 ```
+
+> intake_height or altitude_final
+
+The identification
+of the altitude and type is critical. The approach used here 
+consists of:
+
+1. Use intake_height
+2. Identify `agl` from the name of the tile.
+3. If `agl` is not present, `agl = altitude - site_elevation`.
+4. If there are some NA in elevation, will result some NA in `agl`
+5. A new column is added named `altitude_final` to store `agl` or `asl`
+6. Another column named `type_altitude` is added to identify `agl` as 0, or `asl` as 1.
+
+To maintain a coherent approach among text and NetCDF files, 
+we will add a column altitude_final and type_altitude. This
+is done in `obs_read_nc` or `obs_read`.
 
@@
 by using the coordinates.
-
-# TODO ADD identifier of type of altitude of intake_height
 
@@
 This can be done using UTC, solar, or local time.
-The normal approach is using afteroon solar or local time.
+The normal approach is using afternoon solar or local time.
 
@@
 
-1) solar time
-2) local time with columns `site_utc2lst`
-3) local time longitude
+1) Solar time
+2) Local time with columns `site_utc2lst`
+3) Local time longitude
 
-1) solar time (default)
+> solar time (default)
 
@@
 In this way, we will have one information per day. however
-this approach is not appropiate for aircraft
+this approach is not appropriate for aircraft
 which are aggregated every 10 or 20 seconds. Hence
@@
 
-2) local time with column `site_utc2lst`
+> local time with column `site_utc2lst`
 
@@
 
-3) local time longitude
+> local time longitude
 
@@
 
-```{r}
-df2[, c("solar_time", "solar_time_cut")]
-```
-
 How we filter for the required solar time, in this case `r evening`.
@@
 df3 <- df2[hour_st %in% evening]
+df3[, c("solar_time", "solar_time_cut")]
 
@@
                         "longitude",
-                        # "type_altitude",
-                        "year_end",
                         "site_utc2lst"),
@@
 ```{r}
-if(any(names(df) %in% "intake_height")) {
-  df4[,
-      max_altitude := max(intake_height),
-      by = site_code]
-  df4[,
-      c("site_code",
-        "intake_height",
-        "max_altitude")] |> unique()
-} else {
 df4[,
@@
       "max_altitude")] |> unique()
-}
 
@@
 ```{r, eval = F}
-readLines(csvy)[1:33]
+readLines(csvy)[1:34] # and more
 ```
@@
 
-if(any(names(df) %in% "intake_height")) {
-receptor <- master[intake_height == max_altitude,
-                   c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour", 
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "intake_height",
-                     "type_altitude",
-                     "year_end",
-                     "month_end",
-                     "day_end",
-                     "hour_end",
-                     "minute_end",
-                     "second_end")]
-receptor$intake_height <- round(receptor$intake_height)
-
-} else {
 receptor <- master[altitude_final == max_altitude,
@@
                      "minute_end",
-                     "second_end")]
+                     "second_end",
+                     "time_decimal")]
 
 receptor$altitude_final <- round(receptor$altitude_final)
-}
 

### 2024-05-03 R/02_read_obs.R

@@ -458,6 +458,10 @@ obs_read_nc <- function(index,
 
     dt$obspack_citation <- NULL
     ncdf4::nc_close(nc)
+
+    # if exist, cool, if not, NA
+    dt$altitude_final <- dt$intake_height
+    dt$type_altitude <- ifelse(dt$dataset_intake_ht_unit == "magl", 0, 1)
     dt
   }) -> lx
 

### 2024-05-03 R/04_aggregate_obs.R

@@ -60,6 +60,7 @@ obs_agg <- function(dt,
   timecol <- NULL
   site_code <- NULL
   altitude_final <- NULL
+  type_altitude <- NULL
   timeUTC <- NULL
   key_time <- NULL
   lab_1_abbr <- NULL

### 2024-07-08 README.Rmd
web page

@@
 
+ <!-- badges: start -->
+
 ![GitHub commit
 activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
-
-[![R build
-status](https://github.com/ibarraespinosa/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
-
-[![R-CMD-check](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml)
+  [![R-CMD-check](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml)
+  <!-- badges: end -->
 

### 2024-07-08 README.Rmd
docs

@@
 
- <!-- badges: start -->
-
-![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
+![GitHub commit activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
 [![R-CMD-check](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml)
-  <!-- badges: end -->
+![GitHub Repo stars](https://img.shields.io/github/stars/noaa-gml/rtorf)
 

### 2024-07-10 README.Rmd
remove cols arguemnt from obs_Read_nc

@@
 ```{r}
-df <- obs_read_nc(index = index,
+df <- obs_read_nc(
+  index = index,
   categories = "tower-insitu")
@@
 ```{r, eval = T}
-master <- df3
+master <- df4
 master$timeUTC <- as.character(master$timeUTC)

### 2024-07-10 paper_elsevier/paper_elsevier.Rmd
remove cols arguemnt from obs_Read_nc

@@
 ```{r install, eval = F}
-remotes::install_github("ibarraespinosa/rtorf")
+remotes::install_github("noaa-gml/rtorf")
 library(rtorf)
@@
 
-https://github.com/ibarraespinosa/rtorf/tree/main/rscripts
+https://github.com/noaa-gml/rtorf/tree/main/rscripts
 
@@
     |-- inputs_inv.R
-```
-
-```{r flow, echo = FALSE, message=FALSE, fig.cap="Methane observations (ppb) by towers-insitu", out.height="50%", out.width="100%"}
-knitr::include_graphics("flow/Slide1.PNG")
 ```

### 2024-07-10 R/02_read_obs.R

@@ -381,7 +381,8 @@ obs_read_nc <- function(index,
     names(nc$var)
 
     na <- names(nc$var)
-    na <- data.frame(vars = names(nc$var), stringsAsFactors = FALSE)
+    na <- data.frame(vars = names(nc$var),
+                     stringsAsFactors = FALSE)
 
     la <- lapply(1:nrow(na), function(i) {
       suppressWarnings(
@@ -456,13 +457,18 @@ obs_read_nc <- function(index,
     # }
 
 
-    dt$obspack_citation <- NULL
+    # dt$obspack_citation <- NULL
     ncdf4::nc_close(nc)
 
     # if exist, cool, if not, NA
     dt$altitude_final <- dt$intake_height
     dt$type_altitude <- ifelse(dt$dataset_intake_ht_unit == "magl", 0, 1)
-    dt
+
+    # if(!is.null(cols)) {
+      # dt <- dt[, cols, with = FALSE]
+    # } else {
+      dt
+    # }
   }) -> lx
 
 

### 2024-07-11 R/02_read_obs.R
altituyde for obs_read_nc

@@ -461,8 +461,15 @@ obs_read_nc <- function(index,
     ncdf4::nc_close(nc)
 
     # if exist, cool, if not, NA
-    dt$altitude_final <- dt$intake_height
-    dt$type_altitude <- ifelse(dt$dataset_intake_ht_unit == "magl", 0, 1)
+    if(any(grepl("aircraft", unique(dt$dataset_project)))){
+      dt$altitude_final <- dt$altitude
+      dt$type_altitude <- 1
+
+    } else {
+      dt$altitude_final <- dt$intake_height
+      dt$type_altitude <- ifelse(dt$dataset_intake_ht_unit == "magl", 0, 1)
+
+    }
 
     # if(!is.null(cols)) {
       # dt <- dt[, cols, with = FALSE]

### 2024-07-11 R/02_read_obs.R
obs_read_nc reads `altitude` for aicraft and aircore

@@ -461,7 +461,8 @@ obs_read_nc <- function(index,
     ncdf4::nc_close(nc)
 
     # if exist, cool, if not, NA
-    if(any(grepl("aircraft", unique(dt$dataset_project)))){
+    #aircore or aircraft
+    if(any(grepl("airc", unique(dt$dataset_project)))){
       dt$altitude_final <- dt$altitude
       dt$type_altitude <- 1
 

### 2024-07-15 README.Rmd
fix #8

@@
 
-```{r, eval = F}
+```{r install, eval = F}
 remotes::install_github("noaa-gml/rtorf")
@@
 
-```{r, include = T, message=F, warning=F}
+```{r call, include = T, message=F, warning=F}
 library(rtorf)
@@
 
-```{r, eval = T}
+```{r read1, eval = T}
 
@@
 
-```{r}
+```{r read2}
 df <- obs_read_nc(
@@
 
-```{r}
+```{r table}
 dft <- obs_table(df = df)
@@
 
-```{r}
+```{r uniques}
 usites <- unique(df$site_name)[sample(seq_along(unique(df$site_name)), 2)]
@@
 
-```{r}
+```{r table1}
 knitr::kable(dft[site_name %in% usites])
@@
 
-```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
+```{r plot1, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
 obs_plot(dt = df[site_name %in% usites], 
@@
 
-```{r}
+```{r filter}
 north <- 80
@@
 
-```{r}
+```{r altitude}
 dfa <- df[,
@@
 
-```{r}
+```{r addtime}
 df2 <- obs_addtime(df)
@@
 
-```{r}
+```{r st}
 df2$solar_time <- obs_addstime(df2)
@@
 
-```{r}
+```{r cut}
 df2$solar_time_cut <- cut(x = df2$solar_time,
@@
 
-```{r}
+```{r checkst}
 df3 <- df2[hour_st %in% evening]
@@
 
-```{r}
+```{r keytime}
 df3$key_time <- df3$solar_time_cut
@@
 
-```{r}
+```{r agg}
 df4 <- obs_agg(dt = df3,
-               gby = "mean",
                cols = c("value",

### 2024-07-15 R/04_aggregate_obs.R

@@ -6,9 +6,6 @@
 #'
 #'
 #' @param dt obspack data.table
-#' @param gby (groupby) Character indicating the function to group the
-#' data. Default is "mean" meaning that all numeric variables
-#' will be averaged by the new_second column established in the `idcol`
 #' @param cols Character which defines columns to be aggregated
 #' @param byalt Logical, to aggregate by altitude or not (used in tower)
 #' @param verbose logical to show more information
@@ -19,10 +16,9 @@
 #' @export
 #' @examples \dontrun{
 #' # Do not run
-#' dt <- obs_addtime(dt)
+#' # dt <- obs_addtime(dt)
 #' }
 obs_agg <- function(dt,
-                    gby = "mean",
                     cols = c("year",
                              "month",
                              "day",
@@ -71,7 +67,9 @@ obs_agg <- function(dt,
   intake_height  <- NULL
   local_time  <- NULL
   day <- NULL
+  gby <- NULL
   # mean <- NULL
+
 
   # Here we need daily information
   # as this function assumes that the input is already preselected data
@@ -90,7 +88,7 @@ obs_agg <- function(dt,
     # each group must be processed differently
     dt <- dt[,
              lapply(.SD,
-                    mean,
+                    FUN = mean,
                     na.rm = T),
              .SDcols = cols,
              by = list(timeUTC = key_time,
@@ -102,7 +100,7 @@ obs_agg <- function(dt,
   } else {
     dt <- dt[,
              lapply(.SD,
-                    mean,
+                    FUN = mean,
                     na.rm = T),
              .SDcols = cols,
              by = list(timeUTC = key_time,

### 2024-07-16 R/02_read_obs.R
fix #10, rmarkdown runs README but pkgdown fails generate site

@@ -11,6 +11,7 @@
 #' : of c("aircraft-pfp", "aircraft-insitu",
 #' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
 #' "flask").
+#' @param expr String expressions to filter data.table internally
 #' @param verbose Logical to show more information
 #' @param n_site_code number of characters extratced from metadata after search
 #' @param n_site_latitude number of characters extracted from metadata after search
@@ -53,9 +54,11 @@
 #' obs <- system.file("data-raw", package = "rtorf")
 #' index <- obs_summary(obs)
 #' dt <- obs_read(index)
+#' obs_read(index, expr = "altitude_final == '5800'")
 #' }
 obs_read <- function(index,
                      categories = "flask",
+                     expr = NULL,
                      verbose = TRUE,
                      n_site_code = 15,
                      n_site_latitude = 18,
@@ -288,6 +291,7 @@ obs_read <- function(index,
                                dt$type_altitude)
     dt$id <- i
 
+    if(!is.null(expr)) dt <- dt[eval(parse(text = expr))]
     dt
 
   }) -> lx
@@ -336,6 +340,7 @@ obs_read <- function(index,
 #' : of c("aircraft-pfp", "aircraft-insitu",
 #' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
 #' "flask").
+#' @param expr String expressions to filter data.table internally
 #' @param solar_time Logical, add solar time?
 #' @param as_list Logical to return as list
 #' @param verbose Logical to show more information
@@ -354,6 +359,7 @@ obs_read <- function(index,
 #' }
 obs_read_nc <- function(index,
                         categories = "flask",
+                        expr = NULL,
                         solar_time = TRUE,
                         as_list = FALSE,
                         verbose = FALSE,
@@ -475,7 +481,8 @@ obs_read_nc <- function(index,
     # if(!is.null(cols)) {
       # dt <- dt[, cols, with = FALSE]
     # } else {
-      dt
+    if(!is.null(expr)) dt <- dt[eval(parse(text = expr))]
+    dt
     # }
   }) -> lx
 

### 2024-08-19 R/02_read_obs.R
fix #11

@@ -340,6 +340,7 @@ obs_read <- function(index,
 #' : of c("aircraft-pfp", "aircraft-insitu",
 #' "surface-insitu", "tower-insitu", "aircore", "surface-pfp", "shipboard-insitu",
 #' "flask").
+#' @param att Logical, if global attributes should be added to data.table
 #' @param expr String expressions to filter data.table internally
 #' @param solar_time Logical, add solar time?
 #' @param as_list Logical to return as list
@@ -359,13 +360,14 @@ obs_read <- function(index,
 #' }
 obs_read_nc <- function(index,
                         categories = "flask",
+                        att = FALSE,
                         expr = NULL,
                         solar_time = TRUE,
                         as_list = FALSE,
                         verbose = FALSE,
                         warnings = FALSE){
 
-  if(nrow(index) == 0) stop("empty index")
+    if(nrow(index) == 0) stop("empty index")
 
   if(verbose) cat(paste0("Searching ", categories, "...\n"))
   sector <- NULL
@@ -444,24 +446,26 @@ obs_read_nc <- function(index,
 
     dt$scale <- la$value[["scale_comment"]]
 
+    if(att) {
 
-    log <- utils::capture.output(
-      global <- ncdf4::ncatt_get(nc = nc,
-                                 varid = 0, verbose = F)
-    )
-    if(warnings) print(log)
+      log <- utils::capture.output(
+        global <- ncdf4::ncatt_get(nc = nc,
+                                   varid = 0, verbose = F)
+      )
+      if(warnings) print(log)
 
-    x <- do.call("cbind", global)
-    x <- as.data.frame(x)
-    for(i in 1:ncol(x)) {
-      dt[[names(x)[i]]] <- global[[names(x)[i]]]
+      x <- do.call("cbind", global)
+      x <- as.data.frame(x)
+      for(i in 1:ncol(x)) {
+        dt[[names(x)[i]]] <- global[[names(x)[i]]]
+      }
+
+      # for(i in seq_along(global)) {
+      #   # print(length(global[[i]]))
+      #   print(global[[names(global)[i]]])
+      # }
+
     }
-
-    # for(i in seq_along(global)) {
-    #   # print(length(global[[i]]))
-    #   print(global[[names(global)[i]]])
-    # }
-
 
     # dt$obspack_citation <- NULL
     ncdf4::nc_close(nc)
[…]

### 2024-08-19 R/02_read_obs.R
add more info for article/vignette aircraft-pfp, more coming

@@ -367,7 +367,7 @@ obs_read_nc <- function(index,
                         verbose = FALSE,
                         warnings = FALSE){
 
-    if(nrow(index) == 0) stop("empty index")
+  if(nrow(index) == 0) stop("empty index")
 
   if(verbose) cat(paste0("Searching ", categories, "...\n"))
   sector <- NULL
@@ -446,14 +446,19 @@ obs_read_nc <- function(index,
 
     dt$scale <- la$value[["scale_comment"]]
 
+
+    log <- utils::capture.output(
+      global <- ncdf4::ncatt_get(nc = nc,
+                                 varid = 0, verbose = F)
+    )
+    if(warnings) print(log)
+
+    #here we manually add intake and site unit
+    suppressWarnings(dt$dataset_intake_ht_unit <- global[["dataset_intake_ht_unit"]])
+    suppressWarnings(dt$site_elevation_unit <- global[["site_elevation_unit"]])
+    suppressWarnings(dt$dataset_project <- global[["dataset_project"]])
+
     if(att) {
-
-      log <- utils::capture.output(
-        global <- ncdf4::ncatt_get(nc = nc,
-                                   varid = 0, verbose = F)
-      )
-      if(warnings) print(log)
-
       x <- do.call("cbind", global)
       x <- as.data.frame(x)
       for(i in 1:ncol(x)) {

### 2024-08-20 R/02_read_obs.R
fix readme

@@ -457,6 +457,22 @@ obs_read_nc <- function(index,
     suppressWarnings(dt$dataset_intake_ht_unit <- global[["dataset_intake_ht_unit"]])
     suppressWarnings(dt$site_elevation_unit <- global[["site_elevation_unit"]])
     suppressWarnings(dt$dataset_project <- global[["dataset_project"]])
+    suppressWarnings(dt$dataset_selection_tag <- global[["dataset_selection_tag"]])
+
+    # obs_table
+    suppressWarnings(dt$site_name <- global[["site_name"]])
+    suppressWarnings(dt$site_elevation <- global[["site_elevation"]])
+    suppressWarnings(dt$site_latitude <- global[["site_latitude"]])
+    suppressWarnings(dt$site_longitude <- global[["site_longitude"]])
+    suppressWarnings(dt$site_country <- global[["site_country"]])
+    suppressWarnings(dt$site_code <- global[["site_code"]])
+    suppressWarnings(dt$site_utc2lst <- global[["site_utc2lst"]])
+
+    # obs_agg
+    suppressWarnings(dt$lab_1_abbr <- global[["lab_1_abbr"]])
+    suppressWarnings(dt$dataset_calibration_scale <- global[["dataset_calibration_scale"]])
+
+
 
     if(att) {
       x <- do.call("cbind", global)

### 2024-08-20 R/03_addtime_obs.R
added article aircraft-pfp

@@ -119,15 +119,16 @@ obs_addltime <- function(dt,
   # Priority site_utc2lst and if it is not available,
   if(any(utc2lt == names(dt))) {
 
-    cat("Found site_utc2lst\n")
+    cat(paste("Found ", utc2lt, "\n"))
     dt$local_time <- dt[[timeUTC]] + as.numeric(dt[[utc2lt]])*60*60
+    dt$local_time <- ifelse(
+      is.na(dt$local_time),
+      dt[[timeUTC]] + dt[[longitude]]/15*60*60, #john miller approach
+      dt$local_time)
+  } else {
+    dt$local_time <- dt[[timeUTC]] + dt[[longitude]]/15*60*60 #john miller approach
   }
 
-
-  dt$local_time <- ifelse(
-    is.na(dt$local_time),
-    dt[[timeUTC]] + dt[[longitude]]/15*60*60, #john miller approach
-    dt$local_time)
 
   if(inherits(dt$local_time, "numeric")) {
     dt$local_time <- as.POSIXct(dt$local_time,

### 2024-08-20 README.Rmd
fix readme
site

@@
 ---
-title: "R Tools for Obspack, Footprints and Receptors (rtorf)"
+title: "R Tools for Obspack, Receptors and Footprints (rtorf)"
 output: github_document
@@
 
-```{r sf, echo = TRUE, message=FALSE}
+```{r sfsave, echo = FALSE, message=FALSE}
+library(sf)
+sdft <- st_as_sf(dft[stat == "mean"], 
+                 coords = c("longitude", "latitude"), 
+                 crs = 4326)
+sdft$value <- sdft$value*1e09
+png("man/figures/plotsf.png", width = 2000, height = 1500, res = 200)
+plot(sdft["value"], 
+     reset = FALSE, 
+     axes = TRUE, 
+     graticule = TRUE,
+     pch = 16, 
+     cex = 1.5, 
+     pal = cptcity::lucky(colorRampPalette = T), #random
+     main = NULL)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, echo = TRUE, message=FALSE, eval = FALSE}
 library(sf)
@@
 ```
+
+![](man/figures/plotsf.png)
 
@@
 
-unique(df[, c("altitude_final", "site_elevation", "elevation",
+unique(df[, c("altitude_final", 
+              "site_elevation",
+              "elevation",
               "dataset_selection_tag",
@@
                         "latitude",
-                        "longitude",
-                        "site_utc2lst"),
+                        "longitude"),
                verbose = T,

### 2024-08-20 README.Rmd
site

@@
 
-```{r sfsave, echo = FALSE, message=FALSE}
-library(sf)
-sdft <- st_as_sf(dft[stat == "mean"], 
-                 coords = c("longitude", "latitude"), 
-                 crs = 4326)
-sdft$value <- sdft$value*1e09
-png("man/figures/plotsf.png", width = 2000, height = 1500, res = 200)
-plot(sdft["value"], 
-     reset = FALSE, 
-     axes = TRUE, 
-     graticule = TRUE,
-     pch = 16, 
-     cex = 1.5, 
-     pal = cptcity::lucky(colorRampPalette = T), #random
-     main = NULL)
-maps::map(add = T)
-dev.off()
-```
+(run it locally)
 
@@
 ```
-
-![](man/figures/plotsf.png)
 
@@
 
-```{r plot1, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
+(run it locally)
+
+```{r plot1, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack", eval = F}
 obs_plot(dt = df[site_name %in% usites], 

### 2024-08-22 examples/aircraft-flask.Rmd
add plots to articles

---
title: "aircraft-flask"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aircraft-pfp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rtorf)
library(data.table)
```

```{r read1, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "/media/sergio/ext5/obspack/obspack_ch4_1_GLOBALVIEWplus_v6.0_2023-12-01/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

Now we read the aircraft-pfp using the function `obs_read_nc`. To this date,
`solar_time` is not included for aircraft, so we `FALSE` that argument.

```{r readnc, eval = T}
datasetid <- "aircraft-flask"
df <- obs_read_nc(index = index,
                  categories = datasetid,
                  solar_time = FALSE,
                  verbose = T)

```

Now we check the data

```{r checkdf, eval = T}
df
```

Now we can process the data. We first filter for observations
within our spatial domain:

## Checks and definitions

```{r spatial, eval = T}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
yy <- 2020

```

We check altitude, intake_height, altitude_final and elevation.
altitude_final is a column from intake_height, added to match
column from obs_read text files.

```{r checkcols, eval = T}
df[, c("altitude", "altitude_final", "intake_height", "elevation")]
```

The temporal range of data is
```{r range_year, eval = T}
range(df$year)
```

We also check for dimensions of data

```{r dim_df, eval = T}
dim(df)
```

## Filters

```{r spatial_temporal_filter, eval = T}
df <- df[year == yy]

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
dim(df)
```

## Adding time as POSIXct class

```{r add_time1, eval = T}
df <- obs_addtime(df)
df[, "timeUTC"]
```

## Cut time

now we can cut time every 20 seconds. We can chosse other frequency as well.

```{r cut_sec, eval = T}
df$sec2 <- obs_freq(x = df$second,
                     freq = seq(0, 59, 20))
df[, c("second", "sec2")]
```

## Aggregate data

now we need to add the column key_time, that it will be used
to aggregate other variables

```{r key_time, eval = T}
df$key_time <- ISOdatetime(year = df$year,
                           month = df$month,
                           day = df$day,
                           hour = df$hour,
                           min = df$minute,
                           sec = df$sec2,
                           tz = "UTC")
df[, c("timeUTC", "key_time")]

```

```{r aggregatingdata, eval = T}
df2 <- obs_agg(df, cols =  c("year",
                             "month",
                             "day",
                             "hour",
                             "minute",
                             "second",
                             "time",
                             "time_decimal",
                             "value",
                             "latitude",
                             "longitude",
                             "altitude_final",
                             "elevation",
                             "intake_height",
                             # "gps_altitude",
                             # "pressure",
                             # "pressure_altitude",
                             # "u", "v", "temperature",
                             "type_altitude"))
```
## Add local time

Now we add local time

```{r addlt, eval = T}
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("site_code", "timeUTC"),
          order = c(-1, 1))
df3

```

## Master

Before generating the receptors list, we have the databe with all the required information

```{r rename, eval = T}
master <- df3

```
We may replace missing values with a nine nines. Here is commented

#master[is.na(master)] <- 999999999

We transform the time variables to character and round coordinates with 4 digits

```{r round, eval = T}
master$timeUTC <- as.character(master$timeUTC)
master$local_time <- as.character(master$local_time)
master$latitude <- round(master$latitude, 4)
master$longitude <- round(master$longitude, 4)

```

## Save master
Finally we save the master file

```{r outfile, eval = T}
out <- tempfile()

```

### txt

```{r outtxt, eval = T}
message(paste0(out,"_", datasetid, ".txt\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".txt"),
       sep = " ")

```

### csv

```{r outcsv, eval = T}
message(paste0(out,"_", datasetid, ".csv\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".csv"),
       sep = ",")

```

### csvy

CSVY are csv files with a YAML header to include metadata in tabulated
text files

```{r csvy, eval = T}
cat("\nAdding notes in csvy:\n")
notes <- c(paste0("sector: ", datasetid),
           paste0("timespan: ", yy),
           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
           "data: Data averaged every 20 seconds",
           paste0("altitude: < ", max_altitude),
           "hours: All",
           "local_time: if var `site_utc2lst` is not available, calculated as",
           "longitude/15*60*60 (John Miller)")

cat(notes, sep = "\n")

message(paste0(out,"_", datasetid, ".csvy\n"))
obs_write_csvy(dt = master,
               notes = notes,
               out = paste0(out,"_", datasetid, ".csvy"))
```

```{r readcsvy, eval = T}
obs_read_csvy(paste0(out,"_", datasetid, ".csvy"))
```

## Receptors

Now we can do the last step which is generating the receptor list files.
Now we filter selected columns

```{r receptors, eval = T}
receptor <- master[, c("site_code",
                       "year",
                       "month",
                       "day",
                       "hour",
                       "minute",
                       "second",
                       "latitude",
                       "longitude",
                       "altitude_final",
                       "type_altitude")]

```

We can round altitude also

```{r round_alt, eval = T}
receptor$altitude_final <- round(receptor$altitude_final)
```

Now we can format time variables with two digits

```{r formatrec, eval = T}
receptor <- obs_format(receptor,
                        spf =  c("month", "day",
                                 "hour", "minute", "second"))
```

We have a column that indicate AGL or ASL

```{r aslagl, eval = T}
receptor_agl <- receptor[type_altitude == 0]
receptor_asl <- receptor[type_altitude == 1]
```

Finally, we save the receptors

```{r save_receptors, eval = T}
if(nrow(receptor_agl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_AGL.txt"), "\n")

  fwrite(x = receptor_agl,
         file = paste0(out, "_", datasetid, "_receptor_AGL.txt"),
         sep = " ")
}

if(nrow(receptor_asl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_ASL.txt"), "\n")

  fwrite(x = receptor_asl,
         file = paste0(out, "_", datasetid, "receptor_ASL.txt"),
         sep = " ")

}

```

## Plot

Finally, we just plot some data, run it locally

```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
png("../man/figures/obsplot_aircraftflask.png", width = 1500, height = 1000, res = 200)
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
dev.off()
```

```{r obs_plot, fig.width=7, fig.height=5, eval = F}
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
```

![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)

### 2024-08-22 examples/aircraft-insitu.Rmd
add plots to articles

--- 2024-08-22 examples/aircraft-flask.Rmd
@@
 ---
-title: "aircraft-flask"
-date: "`r Sys.Date()`"
+title: "aircraft-insitu"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{aircraft-pfp}
+  %\VignetteIndexEntry{aircraft-insitu}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ```{r readnc, eval = T}
-datasetid <- "aircraft-flask"
+datasetid <- "aircraft-insitu"
 df <- obs_read_nc(index = index,
@@
                              "intake_height",
-                             # "gps_altitude",
-                             # "pressure",
-                             # "pressure_altitude",
-                             # "u", "v", "temperature",
+                             "gps_altitude",
+                             "pressure",
+                             "pressure_altitude",
+                             "u", "v", "temperature",
                              "type_altitude"))
@@
 ```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_aircraftflask.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_aircraftinsitu.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircrafinsitu.png?raw=true)

### 2024-08-22 examples/aircraft-pfp.Rmd
add plots to articles

--- 2024-08-22 examples/aircraft-insitu.Rmd
@@
 ---
-title: "aircraft-insitu"
+title: "aircraft-pfp"
+date: "`r Sys.Date()`"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{aircraft-insitu}
+  %\VignetteIndexEntry{aircraft-pfp}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ```{r readnc, eval = T}
-datasetid <- "aircraft-insitu"
+datasetid <- "aircraft-pfp"
 df <- obs_read_nc(index = index,
@@
 ```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_aircraftinsitu.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_aircraftpfp.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircrafinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftpfp.png?raw=true)
+
+```{r savesf, fig.width=7, fig.height=5, eval = T, echo = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+png("../man/figures/obsplot_aircraftpfp_map.png", width = 1500, height = 1000, res = 200)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, fig.width=7, fig.height=5, eval = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftpfp_map.png?raw=true)

### 2024-08-22 examples/summary.Rmd
add plots to articles

---
title: "summary"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rtorf)
library(data.table)
```

## ObsPack summary

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

>  We first define the categories

```{r read1, eval = F}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)

```

```
Number of files of index: 429
              sector     N
              <char> <int>
 1:     aircraft-pfp    40
 2:  aircraft-insitu    15
 3:    surface-flask   106
 4:   surface-insitu   174
 5:   aircraft-flask     4
 6:          aircore     1
 7:      surface-pfp    33
 8:     tower-insitu    51
 9:  shipboard-flask     4
10: shipboard-insitu     1
11:    Total sectors   429
Detected 190 files with agl
Detected 239 files without agl

```

We can check the table now

```{r check, eval = F}
index
```

```
  1: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_aao_aircraft-pfp_1_allvalid.nc
  2: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_above_aircraft-insitu_1_allvalid.nc
  3: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_abp_surface-flask_1_representative.nc
  4: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_abt_surface-insitu_6_allvalid.nc
  5: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_acg_aircraft-pfp_1_allvalid.nc
 ---                                                                                                           
425: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_yon_surface-insitu_19_representative.nc
426: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_zep_surface-flask_1_representative.nc
427: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_zep_surface-insitu_442_allvalid-15magl.nc
428: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_zot_surface-flask_45_representative.nc
429: Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/ch4_zsf_surface-insitu_442_allvalid-3magl.nc
                                              name     n          sector   agl
                                            <char> <int>          <char> <num>
  1:            ch4_aao_aircraft-pfp_1_allvalid.nc     1    aircraft-pfp    NA
  2:       ch4_above_aircraft-insitu_1_allvalid.nc     2 aircraft-insitu    NA
  3:     ch4_abp_surface-flask_1_representative.nc     3   surface-flask    NA
  4:          ch4_abt_surface-insitu_6_allvalid.nc     4  surface-insitu    NA
  5:            ch4_acg_aircraft-pfp_1_allvalid.nc     5    aircraft-pfp    NA
 ---                                                                          
425:   ch4_yon_surface-insitu_19_representative.nc   425  surface-insitu    NA
426:     ch4_zep_surface-flask_1_representative.nc   426   surface-flask    NA
427: ch4_zep_surface-insitu_442_allvalid-15magl.nc   427  surface-insitu    15
428:    ch4_zot_surface-flask_45_representative.nc   428   surface-flask    NA
```

### 2024-08-22 examples/aircraft-flask.Rmd
add plots to articles

@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)
+
+```{r savesf, fig.width=7, fig.height=5, eval = T, echo = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+png("../man/figures/obsplot_aircraftflask_map.png", width = 1500, height = 1000, res = 200)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, fig.width=7, fig.height=5, eval = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask_map.png?raw=true)

### 2024-08-22 examples/aircraft-flask.Rmd
add plots to articles

--- 2024-08-22 examples/aircraft-flask.Rmd
@@
 
-```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
+```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
 png("../man/figures/obsplot_aircraftflask.png", width = 1500, height = 1000, res = 200)
@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)
+
+```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+png("../man/figures/obsplot_aircraftflask_map.png", width = 1500, height = 1000, res = 200)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, fig.width=7, fig.height=5, eval = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask_map.png?raw=true)

### 2024-08-22 examples/aircraft-pfp.Rmd
add plots to articles

@@
 
-```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
+```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
 png("../man/figures/obsplot_aircraftpfp.png", width = 1500, height = 1000, res = 200)
@@
 
-```{r obs_plot, fig.width=7, fig.height=5, eval = F}
+```{r obs_plot, fig.width=5, fig.height=3, eval = F}
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-```{r savesf, fig.width=7, fig.height=5, eval = T, echo = F}
+```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
 library(sf)
@@
 
-```{r sf, fig.width=7, fig.height=5, eval = F}
+```{r sf, fig.width=5, fig.height=3, eval = F}
 library(sf)

### 2024-08-22 examples/aircraft-insitu.Rmd
add plots to articles

@@
 
-```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
+```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
 png("../man/figures/obsplot_aircraftinsitu.png", width = 1500, height = 1000, res = 200)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircrafinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftinsitu.png?raw=true)
+
+```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+png("../man/figures/obsplot_aircraftinsitu_map.png", width = 1500, height = 1000, res = 200)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, fig.width=5, fig.height=3, eval = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftinsitu_map.png?raw=true)

### 2024-08-22 rscripts/README.R
add plots to articles

@@ -1,38 +1,0 @@
-
-# README ####
-# The directory `rscripts` contains the required files
-# to generate the list of receptors using `rtorf`
-
-# The scripts will run from a defined working directory
-
-
-# 1. Run index (summary)
-
-# The script r/index.R will create the tree directory
-# to store the processed ObsPack files
-# Also, it will create summaries and store them
-# for each category in the directory `obs`
-
-# 2. Categories in the directory r
-# These scripts need to be run using Rscript command
-#examples:
-# Rscript r/aircore_year.R 2020 obs/obs_aircore.csv.gz
-# Rscript r/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv.gz
-# Rscript r/flask_year.R 2020 obs/obs_flask.csv.gz
-# Rscript r/surface_year.R 2020 obs/obs_surface-insitu.csv.gz
-# Rscript r/tower_year.R 2020 obs/obs_tower-insitu.csv.gz
-
-# Windows users can adapt the following lines
-# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/tower_insitu_year.R 2020 obs/obs_tower-insitu.csv.gz
-# "C:/Users/sibarrae/My Documents/R/R-4.1.3/bin/Rscript.exe" r/aircore_year.R 2020 obs/obs_aircore.csv.gz
-
-
-# 3. Run Hysplit
-# Once if we have the receptor, we can
-# run Hysplit to generate footprints in NetCDF format
-
-# 4. Generate receptor info
-# Once we have the receptor list and the footprints were generated,
-# we can generate the receptor info, which
-# consists in text files.
-# The script is r/inputs_inv.R

### 2024-08-22 rscripts/r/aircore_year.R

@@ -1,173 +1,0 @@
-#example Rscript rscripts/aircore_year.R 2020 obs/obs_aircore.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-
-out <- paste0("master/master_aircore", yy)
-
-message("All missing values are filled with 999999")
-
-cat("                 _     rtorffilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(rtorf)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'aircore',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "aircore",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ###
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
[…]

### 2024-08-22 rscripts/r/aircraft_year.R

@@ -1,207 +1,0 @@
-#example Rscript rscripts/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-
-out <- paste0("master/master_aircraft_insitu_", yy)
-
-message("All missing values are filled with 999999")
-
-cat("                 _     rtorffilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(rtorf)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# evening <- 14 # for aircraft keep all the local hours
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'aircraft-insitu',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "aircraft-insitu",
-#               verbose = T)
-df  <- fread(ind)
-
-# 3 ECO 2020 ####
-(cat("Reading ECO 2020\n"))
-
-if(Sys.info()[["sysname"]] == "Windows") {
[…]

### 2024-08-22 rscripts/r/flask_non_noaa_year.R

@@ -1,205 +1,0 @@
-#example Rscript rscripts/flask_year.R 2020 obs/obs_flask.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_flask_no_noaa_", yy)
-
-
-
-message("All missing values are filled with 999999")
-
-cat("                 _     rtorffilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(rtorf)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'flask',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "flask",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ###
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
-
[…]

### 2024-08-22 rscripts/r/index.R

@@ -1,54 +1,0 @@
-wd <- paste0(getwd(), "/rscripts") #  you can change
-# if you change, define a new path
-# wd <- "path"
-# then set working directory
-setwd(wd)
-
-cat("This is your working directory", getwd(), "\n")
-
-library(data.table)
-library(rtorf)
-
-categories <- c(
-  #"aircraft-pfp",
-  "aircraft-insitu",
-  "surface-insitu",
-  "aircore",
-  #"surface-pfp",
-  "tower-insitu",
-  "shipboard-insitu",
-  "flask"
-)
-
-# ObsPack ####
-obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt/"
-
-# Create Directories ####
-# master have files with all columns
-dir.create("master", showWarnings = FALSE)
-
-# receptors have files with required columns
-dir.create("receptors", showWarnings = FALSE)
-
-# receptors have files with required columns
-dir.create("obs", showWarnings = FALSE)
-
-# 1 summary (previously named index) ####
-cat("Running obs_summary(obs)\n")
-index <- obs_summary(obs = obs)
-
-
-# 2 read ####
-
-for(i in seq_along(categories)) {
-
-cat(paste0("Running obs_read(index = index,
-               categories = ", categories[i], "\n"))
-
-df <- obs_read(index = index,
-               categories = categories[i],
-               verbose = T)
-
-fwrite(df, paste0("obs/obs_",categories[i], ".csv.gz"))
-
-}

### 2024-08-22 rscripts/r/inputs_inv.R

@@ -1,310 +1,0 @@
-
-library(colorout)
-library(data.table)
-library(rtorf)
-
-dir.create("invfiles")
-
-# aircraft insitu ####
-ma1 <- paste0("/path/master/master_aircraft_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skipping ", ma1[i], "\n")
-    next
-  }
-  dx <- fread(ma1[i])
-  dx$file <- na[i]
-  print(dim(dx))
-  lx[[i]] <- dx
-}
-master <- obs_list.dt(lx)
-# filtering between dec 2017 and jan 2021
-
-master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
-                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
-
-master$nc <- obs_footname(year = master$year,
-                          month = master$month,
-                          day = master$day,
-                          hour = master$hour,
-                          minute = master$minute,
-                          lat = master$latitude,
-                          lon = master$longitude,
-                          alt = master$altitude_final)
-
-path <-  "/path/aircraft/flask/" # param
-
-master$fnc <- paste0(path, master$nc)
-
-master$is_fnc <- file.exists(master$fnc)
-
-table(master$is_fnc)
-
-master <- master[is_fnc == TRUE]
-
-master <- master[!is.na(value)]
-
-master <- master[altitude_final > 10]
-
-dfaircraft <- obs_invfiles(master = master,
-                           path = path,
-                           Type = "continuous",
-                           SubType = "aircraft")
-obs_plot(dfaircraft$master,
-         time = "timeUTC",
-         y = "value",
-         colu = "site_code",
-         yfactor = 1e9,
-         cex = 0.5,
-         type = "l"
-)
-
-
-# tower insitu ####
-ma1 <- paste0("master/master_tower_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skipping ", ma1[i], "\n")
-    next
-  }
[…]

### 2024-08-22 rscripts/r/inputs_invs.R

@@ -1,314 +1,0 @@
-
-library(colorout)
-library(data.table)
-library(rtorffilter)
-
-# dir.create("invfiles")
-
-# aircraft insitu ####
-ma1 <- paste0("master/master_aircraft_insitu_", 2017:2021, ".csv")
-na <- gsub(pattern = ".csv", replacement = "", x = ma1)
-na <- gsub(pattern = "master/master_", replacement = "", x = na)
-
-lx <- list()
-
-for(i in seq_along(ma1)) {
-  if(!file.exists(ma1[i])) {
-    cat("skiiping ", ma1[i], "\n")
-    next
-  }
-  dx <- fread(ma1[i])
-  dx$file <- na[i]
-  print(dim(dx))
-  lx[[i]] <- dx
-}
-master <- obs_list.dt(lx)
-# filtering between dec 2017 and jan 2021
-
-master <- master[timeUTC > ISOdatetime(2017,12,1,0,0,0, tz = "UTC") &
-                   timeUTC < ISOdatetime(2021,2,1,0,0,0, tz = "UTC")]
-#  footprints_hera_hysplit.txt
-master$nc <- obs_footname(year = master$year,
-                          month = master$month,
-                          day = master$day,
-                          hour = master$hour,
-                          minute = master$minute,
-                          lat = master$latitude,
-                          lon = master$longitude,
-                          alt = master$altitude_final)
-
-path <-  "/work/noaa/co2/footprint/hysplit/CPO2021/hysplit4/aircraft/flask/" # param
-
-master$fnc <- paste0(path, master$nc)
-
-master$is_fnc <- file.exists(master$fnc)
-
-table(master$is_fnc)
-
-master <- master[is_fnc == TRUE]
-
-master <- master[!is.na(value)]
-
-master <- master[altitude_final > 10]
-
-dfaircraft <- obs_invfiles(master = master,
-                           path = path,
-                           Type = "continuous",
-                           SubType = "aircraft")
-plot(dfaircraft)
-
-obs_plot(dfsurface$master,
-         time = "timeUTC",
-         y = "value",
-         colu = "site_code",
-         n = c("BRA", "EGB", "EST", "ETL", "LLB"),
-         yfactor = 1e9,
-         cex = 0.5,
-         type = "l",
-         ylim = c(1800, 3000)
-)
-
-
-# Found  760 missing `value` on master
-# It is possible that this script must be run again to explude missing value
-
-
-# tower insitu ####
[…]

### 2024-08-22 rscripts/r/surface_insitu_year.R

@@ -1,236 +1,0 @@
-#example Rscript rscripts/surface_year.R 2020 obs/obs_surface-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index", call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_surface_insitu_", yy)
-
-
-message("All missing values are filled with 999999")
-
-cat("                 _     rtorffilter              \n")
-cat("               (`  ).                  _           \n")
-cat("              (     ).             .:(`  )`.       \n")
-cat(")           _(       '`.          :(   .    )      \n")
-cat("        .=(`(      .   )     .--  `.  (    ) )      \n")
-cat("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-cat("`.     `(       ) )       (   .  )     (   )  ._   \n")
-cat("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-cat(")  )  ( )       --'       `- __.'         :(      )) \n")
-cat(".-'  (_.'          .')                    `(    )  ))\n")
-cat("                  (_  )                     ` __.:'          \n")
-cat("\n")
-cat("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-
-library(data.table)
-library(rtorf)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14:15
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'surface-insitu',
-#               verbose = T)\n")
-#df <- obs_read(index = index,
-#               categories = "surface-insitu",
-#               verbose = T)
-
-df <- fread(ind)
-
-# 3 Filtering space and time ####
-cat("Filtering\n")
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
[…]

### 2024-08-22 rscripts/r/tower_insitu_year.R

@@ -1,238 +1,0 @@
-#example Rscript rscripts/tower_year.R 2020 obs/obs_tower-insitu.csv
-
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) != 2) {
-  stop("At least one argument must be supplied (year).n and obs_index",
-       call.=FALSE)
-} else {
-  yy <-  as.numeric(args[1] )
-  print(yy)
-
-  ind  <- as.character(args[2])
-  print(ind)
-
-}
-out <- paste0("master/master_tower_insitu_", yy)
-
-message("All missing values are filled with 999999")
-
-n01 <- ("                  _     rtorffilter              \n")
-n02 <- ("               (`  ).                  _           \n")
-n03 <- ("              (     ).             .:(`  )`.       \n")
-n04 <- (")           _(       '`.          :(   .    )      \n")
-n05 <- ("        .=(`(      .   )     .--  `.  (    ) )      \n")
-n06 <- ("       ((    (..__.:'-'   .+(   )   ` _`  ) )                 \n")
-n07 <- ("`.     `(       ) )       (   .  )     (   )  ._   \n")
-n08 <- ("  )      ` __.:'   )     (   (   ))     `-'.-(`  ) \n")
-n09 <- (")  )  ( )       --'       `- __.'         :(      )) \n")
-n10 <- (".-'  (_.'          .')                    `(    )  ))\n")
-n11 <- ("                  (_  )                     ` __.:'          \n")
-n12 <- ("\n")
-n13 <- ("--..,___.--,--'`,---..-.--+--.,,-,,..._.--..-._.-a:f--.\n")
-cat(c(n01, n02,n03, n04, n05,
-      n06, n07, n08, n09, n10,
-      n11, n12, n13))
-
-library(data.table)
-library(rtorf)
-
-# 0 filters ####
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14:15
-
-# 1 Index ####
-#cat("Running obs_index(obs)\n")
-#index <- obs_index(obs = obs)
-
-# 2 Read ####
-#cat("Running obs_read(index = index,
-#               categories = 'tower-insitu',
-#               verbose = T)\n")
-# df <- obs_read(index = index,
-#               categories = "tower-insitu",
-#               verbose = T)
-df <- fread(ind)
-
-# 3 Filtering space and time ####
-cat("Filterinig\n")
[…]

### 2024-08-22 examples/surface-insitu.Rmd
add add surface insitu article

---
title: "surface-insitu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{surface-insitu}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rtorf)
library(data.table)
```

```{r read1, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

Now we read the surface-insitu using the function `obs_read_nc`. 
`solar_time` is  included for surface, so we `TRUE` that argument.

```{r readnc, eval = T}
datasetid <- "surface-insitu"
df <- obs_read_nc(index = index,
                  categories = datasetid,
                  solar_time = TRUE,
                  verbose = TRUE)

```

Now we check the data

```{r checkdf, eval = T}
df
```

Now we can process the data. We first filter for observations
within our spatial domain:

## Checks and definitions

```{r spatial, eval = T}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
yy <- 2020
evening <- 14

```

We check altitude, intake_height, altitude_final and elevation.
altitude_final is a column from intake_height, added to match
column from obs_read text files.

```{r checkcols, eval = T}
df[, c("altitude", "altitude_final", "intake_height", "elevation",
       "dataset_selection_tag",
              "site_name")]
```

The temporal range of data is
```{r range_year, eval = T}
range(df$year)
```

We also check for dimensions of data

```{r dim_df, eval = T}
dim(df)
```

## Filters

```{r spatial_temporal_filter, eval = T}
df <- df[year == yy]

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
dim(df)
```

After filtering by space and time, we have `r round(nrow(df), 2)`
million observations. Towers can have observations at different
heights. Here we need to select one site with the observations
registered at the highest height. The column
with the height is named `altitude_final` and the
max altitude was named `max_altitude`.

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
dfa
```

### Key Time

Here we need to start time columns.
The function `obs_addtime` adds time columns
`timeUTC`, `timeUTC_start` which shows the start time
of each observation and `timeUTC_end`
which shows the end time for each observation.

```{r}
df2 <- obs_addtime(df)
```

Then we need a *key_time* to aggregate data.
This can be done using UTC, solar, or local time.
The normal approach is using afternoon solar or local time.

#### Hierarchy of solar or local time

1) Solar time
2) Local time with columns `site_utc2lst`
3) Local time longitude

> solar time (default)

Here we select the hours of interest and then
aggregate data by year, month and day of solar time.
In this way, we will have one information per day. however
this approach is not appropriate for aircraft
which are aggregated every 10 or 20 seconds. Hence
we need to aggregate data by one time column.
Also, this helps to generate the receptor info 
files including hour, minute and second.
Hence, *we need to add solar or local time column*.

```{r}
df2$solar_time <- obs_addstime(df2)

```

> local time with column `site_utc2lst`

Then we need to identify the local time with the function
`add_ltime`. This is important because to identifying
observations in the evening in local time.
`add_ltime` uses two methods, first identify
the time difference with utc by identifying the
metadata column "site_utc2lst". 
If solar time is not available
#now we need to cut solar time for the frequency needed. 
As we will work with

> local time longitude

If this information
is not available, with the aircrafts for instance,
the local time is calculated with an approximation
based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.
Then, the time is cut every two hours.
Now, we identify the local time to select evening hours.

#### Cut time

Now we have they key column time, we can cut it accordingly.

```{r}
df2$solar_time_cut <- cut(x = df2$solar_time,
                          breaks = "1 hour") |>
  as.character()

```

How we can check the solar time and the cut solar time.
Please note that solar_time_cut,
the column that it will be used to aggregate data

How we filter for the required solar time, in this case `r evening`.

```{r}
df3 <- df2[hour_st %in% evening]
df3[, c("solar_time", "solar_time_cut")]

```

Now there are `r nrow(df3)` observations and before filtering
`r nrow(df2)`.
At this point we can calculate the averages of several columns by
the cut time. The function `obs_agg` does this aggregation as shown
in the following lines of code. The argument `gby`
establish the function used to aggregate `cols`. I need to aggregate the data
by date (year, month, date), because it is already filtered by
the hours of interest. Then, I would have 1 observation per day.

As standard, let us define `key_time` as `solar_time`. 
The `obs_agg` function will aggregate the desired data by that column.

```{r}
df3$key_time <- df3$solar_time_cut
```

```{r}
df4 <- obs_agg(dt = df3,
               cols = c("value",
                        "latitude",
                        "longitude",
                        "site_utc2lst"),
               verbose = T,
               byalt = TRUE)
```

Now there are `r nrow(df4)` observations,
`r nrow(df) - nrow(df4)` less observations.
Here we add the column `max_altitude` to identify the max altitude
by site.

```{r}
df4[,
    max_altitude := max(altitude_final),
    by = site_code]
df4[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()

```

## Master

Before generating the receptors list, we have the database with all the required information

```{r rename, eval = T}
master <- df4

```
We may replace missing values with a nine nines. Here is commented

#master[is.na(master)] <- 999999999

We transform the time variables to character and round coordinates with 4 digits

```{r round, eval = T}
master$timeUTC <- as.character(master$timeUTC)
master$local_time <- as.character(master$local_time)
master$latitude <- round(master$latitude, 4)
master$longitude <- round(master$longitude, 4)

```

## Save master
Finally we save the master file

```{r outfile, eval = T}
out <- tempfile()

```

### txt

```{r outtxt, eval = T}
message(paste0(out,"_", datasetid, ".txt\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".txt"),
       sep = " ")

```

### csv

```{r outcsv, eval = T}
message(paste0(out,"_", datasetid, ".csv\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".csv"),
       sep = ",")

```

### csvy

CSVY are csv files with a YAML header to include metadata in tabulated
text files

```{r csvy, eval = T}
cat("\nAdding notes in csvy:\n")
notes <- c(paste0("sector: ", datasetid),
           paste0("timespan: ", yy),
           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
           "data: Data averaged every 20 seconds",
           paste0("altitude: < ", max_altitude),
           paste0("hours: ", evening),
           "local_time: used solar_time")

cat(notes, sep = "\n")

message(paste0(out,"_", datasetid, ".csvy\n"))
obs_write_csvy(dt = master,
               notes = notes,
               out = paste0(out,"_", datasetid, ".csvy"))
```

```{r readcsvy, eval = T}
obs_read_csvy(paste0(out,"_", datasetid, ".csvy"))
```

## Receptors

Now we can do the last step which is generating the receptor list files.
Now we filter selected columns

```{r receptors, eval = T}
receptor <- master[, c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour",
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "altitude_final",
                     "type_altitude",
                     "time_decimal")]

```

We can round altitude also

```{r round_alt, eval = T}
receptor$altitude_final <- round(receptor$altitude_final)
```

Now we can format time variables with two digits

```{r formatrec, eval = T}
receptor <- obs_format(receptor,
                        spf =  c("month", "day",
                                 "hour", "minute", "second"))
```

We have a column that indicate AGL or ASL

```{r aslagl, eval = T}
receptor_agl <- receptor[type_altitude == 0]
receptor_asl <- receptor[type_altitude == 1]
```

Finally, we save the receptors

```{r save_receptors, eval = T}
if(nrow(receptor_agl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_AGL.txt"), "\n")

  fwrite(x = receptor_agl,
         file = paste0(out, "_", datasetid, "_receptor_AGL.txt"),
         sep = " ")
}

if(nrow(receptor_asl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_ASL.txt"), "\n")

  fwrite(x = receptor_asl,
         file = paste0(out, "_", datasetid, "receptor_ASL.txt"),
         sep = " ")

}

```

## Plot

Finally, we just plot some data, run it locally

```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
png("../man/figures/obsplot_surfaceinsitu.png", width = 1500, height = 1000, res = 200)
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
dev.off()
```

```{r obs_plot, fig.width=5, fig.height=3, eval = F}
obs_plot(df4, time = "timeUTC", yfactor = 1e9)
```

![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetinsitu.png?raw=true)

```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
library(sf)
dx <- df4[, 
    lapply(.SD, mean),
    .SDcols = "value",
    by = .(latitude, longitude)]
x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
png("../man/figures/obsplot_surfaceinsitu_map.png", width = 1500, height = 1000, res = 200)
plot(x["value"], axes = T, reset = F)
maps::map(add = T)
dev.off()
```

```{r sf, fig.width=5, fig.height=3, eval = F}
library(sf)
dx <- df4[, 
    lapply(.SD, mean),
    .SDcols = "value",
    by = .(latitude, longitude)]
x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
plot(x["value"], axes = T, reset = F)
maps::map(add = T)
```

![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceinsitu_map.png?raw=true)

### 2024-08-26 examples/README_nc.Rmd
add surface-flask examples and articles

@@
 To maintain a coherent approach among text and NetCDF files, 
-we will add a column altitude_final and type_altitude
-
-```{r}
-df$altitude_final <- df$intake_height
-df$type_altitude <- ifelse(df$dataset_intake_ht_unit == "magl", 0, 1)
-```
+we will add a column altitude_final and type_altitude. This
+is done in `obs_read_nc` or `obs_read`.
 
@@
 ```
+
 Then we need a *key_time* to aggregate data.

### 2024-08-26 examples/surface-flask.Rmd
add surface-flask examples and articles

--- 2024-08-22 examples/surface-insitu.Rmd
@@
 ---
-title: "surface-insitu"
+title: "surface-flask"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{surface-insitu}
+  %\VignetteIndexEntry{surface-flask}
   %\VignetteEngine{knitr::rmarkdown}
@@
 cate = c("aircraft-pfp",
-         "aircraft-insitu",
          "aircraft-flask",
-         "surface-insitu",
+         "aircraft-flask",
+         "surface-flask",
          "surface-flask", 
          "surface-pfp",   
-         "tower-insitu",  
+         "tower-flask",  
          "aircore",       
-         "shipboard-insitu",
+         "shipboard-flask",
          "shipboard-flask")
@@
 
-Now we read the surface-insitu using the function `obs_read_nc`. 
+Now we read the surface-flask using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.
@@
 ```{r readnc, eval = T}
-datasetid <- "surface-insitu"
+datasetid <- "surface-flask"
 df <- obs_read_nc(index = index,
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_surfaceinsitu.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_surfaceflask.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetflask.png?raw=true)
 
@@
 x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_surfaceinsitu_map.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_surfaceflask_map.png", width = 1500, height = 1000, res = 200)
 plot(x["value"], axes = T, reset = F)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceinsitu_map.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceflask_map.png?raw=true)

### 2024-08-26 examples/surface-pfp.Rmd
add surface-pfpk examples and articles

--- 2024-08-22 examples/surface-insitu.Rmd
@@
 ---
-title: "surface-insitu"
+title: "surface-pfp"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{surface-insitu}
+  %\VignetteIndexEntry{surface-pfp}
   %\VignetteEngine{knitr::rmarkdown}
@@
 
-Now we read the surface-insitu using the function `obs_read_nc`. 
+Now we read the surface-pfp using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.
@@
 ```{r readnc, eval = T}
-datasetid <- "surface-insitu"
+datasetid <- "surface-pfp"
 df <- obs_read_nc(index = index,
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_surfaceinsitu.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_surfacepfp.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacepfp.png?raw=true)
 
@@
 x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_surfaceinsitu_map.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_surfacepfp_map.png", width = 1500, height = 1000, res = 200)
 plot(x["value"], axes = T, reset = F)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceinsitu_map.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacepfp_map.png?raw=true)

### 2024-08-26 examples/tower-insitu.Rmd
add tower-insitu examples and articles

--- 2024-08-22 examples/surface-insitu.Rmd
@@
 ---
-title: "surface-insitu"
+title: "tower-insitu"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{surface-insitu}
+  %\VignetteIndexEntry{tower-insitu}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ```{r readnc, eval = T}
-datasetid <- "surface-insitu"
+datasetid <- "tower-insitu"
 df <- obs_read_nc(index = index,
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_surfaceinsitu.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_towerinsitu.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_towerinsitu.png?raw=true)
 
@@
 x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_surfaceinsitu_map.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_towerinsitu_map.png", width = 1500, height = 1000, res = 200)
 plot(x["value"], axes = T, reset = F)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceinsitu_map.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_towerinsitu_map.png?raw=true)

### 2024-08-26 examples/aircore.Rmd
add aircore examples and articles

--- 2024-08-22 examples/aircraft-flask.Rmd
@@
 ---
-title: "aircraft-flask"
+title: "aircore"
 date: "`r Sys.Date()`"
@@
 vignette: >
-  %\VignetteIndexEntry{aircraft-pfp}
+  %\VignetteIndexEntry{aircore}
   %\VignetteEngine{knitr::rmarkdown}
@@
 
-obs <- "/media/sergio/ext5/obspack/obspack_ch4_1_GLOBALVIEWplus_v6.0_2023-12-01/data/nc/"
+obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
 index <- obs_summary(obs = obs, 
@@
 ```{r readnc, eval = T}
-datasetid <- "aircraft-flask"
+datasetid <- "aircore"
 df <- obs_read_nc(index = index,
@@
 ```{r checkcols, eval = T}
-df[, c("altitude", "altitude_final", "intake_height", "elevation")]
+df[, c("altitude", "altitude_final")]
 ```
@@
                              "altitude_final",
-                             "elevation",
-                             "intake_height",
+                             # "elevation",
+                             # "intake_height",
                              # "gps_altitude",
@@
 
-```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_aircraftflask.png", width = 1500, height = 1000, res = 200)
+```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
+png("../man/figures/obsplot_aircore.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircore.png?raw=true)
+
+```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+png("../man/figures/obsplot_aircore_map.png", width = 1500, height = 1000, res = 200)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+dev.off()
+```
+
+```{r sf, fig.width=7, fig.height=5, eval = F}
+library(sf)
+x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
+plot(x["value"], axes = T, reset = F)
+maps::map(add = T)
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircore_map.png?raw=true)

### 2024-08-26 examples/shipboard-insitu.Rmd
shipboard

--- 2024-08-26 examples/tower-insitu.Rmd
@@
 ---
-title: "tower-insitu"
+title: "shipboard-insitu"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{tower-insitu}
+  %\VignetteIndexEntry{shipboard-insitu}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ```{r readnc, eval = T}
-datasetid <- "tower-insitu"
+datasetid <- "shipboard-insitu"
 df <- obs_read_nc(index = index,
@@
 max_altitude <- 8000
-yy <- 2020
+yy <- 2019 #no 2020
 evening <- 14
@@
 
-## Filters
-
-```{r spatial_temporal_filter, eval = T}
-df <- df[year == yy]
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
-dim(df)
-```
-
-After filtering by space and time, we have `r round(nrow(df), 2)`
-million observations. Towers can have observations at different
-heights. Here we need to select one site with the observations
-registered at the highest height. The column
-with the height is named `altitude_final` and the
-max altitude was named `max_altitude`.
-
-```{r}
-dfa <- df[,
-          max(altitude_final),
-          by = site_code] |> unique()
-
-names(dfa)[2] <- "max_altitude"
-dfa
-```
-
 ### Key Time
@@
 ```{r}
-df2 <- obs_addtime(df)
+df <- obs_addtime(df)
 ```
@@
 
-#### Hierarchy of solar or local time
-
-1) Solar time
-2) Local time with columns `site_utc2lst`
-3) Local time longitude
-
-> solar time (default)
-
-Here we select the hours of interest and then
-aggregate data by year, month and day of solar time.
-In this way, we will have one information per day. however
-this approach is not appropriate for aircraft
-which are aggregated every 10 or 20 seconds. Hence
-we need to aggregate data by one time column.
-Also, this helps to generate the receptor info 
-files including hour, minute and second.
-Hence, *we need to add solar or local time column*.
-
-```{r}
-df2$solar_time <- obs_addstime(df2)
-
-```
-
-> local time with column `site_utc2lst`
-
-Then we need to identify the local time with the function
-`add_ltime`. This is important because to identifying
-observations in the evening in local time.
-`add_ltime` uses two methods, first identify
-the time difference with utc by identifying the
-metadata column "site_utc2lst". 
-If solar time is not available
-#now we need to cut solar time for the frequency needed. 
-As we will work with
-
-> local time longitude
-
-If this information
-is not available, with the aircrafts for instance,
-the local time is calculated with an approximation
-based on longitude:
-
-$$
-lt = UTC + longitude/15 * 60 * 60
-$$
-Where $lt$ is the local time, $UTC$ the time,
-$longitude$ the coordinate.
-Then, the time is cut every two hours.
-Now, we identify the local time to select evening hours.
-
-#### Cut time
-
-Now we have they key column time, we can cut it accordingly.
-
-```{r}
-df2$solar_time_cut <- cut(x = df2$solar_time,
-                          breaks = "1 hour") |>
-  as.character()
-
-```
-
-How we can check the solar time and the cut solar time.
-Please note that solar_time_cut,
-the column that it will be used to aggregate data
-
-How we filter for the required solar time, in this case `r evening`.
-
-```{r}
-df3 <- df2[hour_st %in% evening]
-df3[, c("solar_time", "solar_time_cut")]
-
-```
-
-Now there are `r nrow(df3)` observations and before filtering
-`r nrow(df2)`.
-At this point we can calculate the averages of several columns by
-the cut time. The function `obs_agg` does this aggregation as shown
-in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`. I need to aggregate the data
-by date (year, month, date), because it is already filtered by
-the hours of interest. Then, I would have 1 observation per day.
-
-As standard, let us define `key_time` as `solar_time`. 
-The `obs_agg` function will aggregate the desired data by that column.
-
-```{r}
-df3$key_time <- df3$solar_time_cut
-```
-
-```{r}
-df4 <- obs_agg(dt = df3,
-               cols = c("value",
-                        "latitude",
-                        "longitude",
-                        "site_utc2lst"),
-               verbose = T,
-               byalt = TRUE)
-```
-
-Now there are `r nrow(df4)` observations,
-`r nrow(df) - nrow(df4)` less observations.
-Here we add the column `max_altitude` to identify the max altitude
-by site.
-
-```{r}
-df4[,
-    max_altitude := max(altitude_final),
-    by = site_code]
-df4[,
-    c("site_code",
-      "altitude_final",
-      "max_altitude")] |> unique()
-
-```
-
-## Master
-
-Before generating the receptors list, we have the database with all the required information
-
-```{r rename, eval = T}
-master <- df4
-
-```
-We may replace missing values with a nine nines. Here is commented
-
-#master[is.na(master)] <- 999999999
-
-We transform the time variables to character and round coordinates with 4 digits
-
-```{r round, eval = T}
-master$timeUTC <- as.character(master$timeUTC)
-master$local_time <- as.character(master$local_time)
-master$latitude <- round(master$latitude, 4)
-master$longitude <- round(master$longitude, 4)
-
-```
-
-## Save master
-Finally we save the master file
-
-```{r outfile, eval = T}
-out <- tempfile()
-
-```
-
-### txt
-
-```{r outtxt, eval = T}
-message(paste0(out,"_", datasetid, ".txt\n"))
-fwrite(master,
-       paste0(out,"_", datasetid, ".txt"),
-       sep = " ")
-
-```
-
-### csv
-
-```{r outcsv, eval = T}
-message(paste0(out,"_", datasetid, ".csv\n"))
-fwrite(master,
-       paste0(out,"_", datasetid, ".csv"),
-       sep = ",")
-
-```
-
-### csvy
-
-CSVY are csv files with a YAML header to include metadata in tabulated
-text files
-
-```{r csvy, eval = T}
-cat("\nAdding notes in csvy:\n")
-notes <- c(paste0("sector: ", datasetid),
-           paste0("timespan: ", yy),
-           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
-           "data: Data averaged every 20 seconds",
-           paste0("altitude: < ", max_altitude),
-           paste0("hours: ", evening),
-           "local_time: used solar_time")
-
-cat(notes, sep = "\n")
-
-message(paste0(out,"_", datasetid, ".csvy\n"))
-obs_write_csvy(dt = master,
-               notes = notes,
-               out = paste0(out,"_", datasetid, ".csvy"))
-```
-
-```{r readcsvy, eval = T}
-obs_read_csvy(paste0(out,"_", datasetid, ".csvy"))
-```
-
-## Receptors
-
-Now we can do the last step which is generating the receptor list files.
-Now we filter selected columns
-
-```{r receptors, eval = T}
-receptor <- master[, c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour",
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "altitude_final",
-                     "type_altitude",
-                     "time_decimal")]
-
-```
-
-We can round altitude also
-
-```{r round_alt, eval = T}
-receptor$altitude_final <- round(receptor$altitude_final)
-```
-
-Now we can format time variables with two digits
-
-```{r formatrec, eval = T}
-receptor <- obs_format(receptor,
-                        spf =  c("month", "day",
-                                 "hour", "minute", "second"))
-```
-
-We have a column that indicate AGL or ASL
-
-```{r aslagl, eval = T}
-receptor_agl <- receptor[type_altitude == 0]
-receptor_asl <- receptor[type_altitude == 1]
-```
-
-Finally, we save the receptors
-
-```{r save_receptors, eval = T}
-if(nrow(receptor_agl) > 0) {
-  message(paste0(out, "_", datasetid, "_receptor_AGL.txt"), "\n")
-
-  fwrite(x = receptor_agl,
-         file = paste0(out, "_", datasetid, "_receptor_AGL.txt"),
-         sep = " ")
-}
-
-if(nrow(receptor_asl) > 0) {
-  message(paste0(out, "_", datasetid, "_receptor_ASL.txt"), "\n")
-
-  fwrite(x = receptor_asl,
-         file = paste0(out, "_", datasetid, "receptor_ASL.txt"),
-         sep = " ")
-
-}
-
-```
-
 ## Plot
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_towerinsitu.png", width = 1500, height = 1000, res = 200)
-obs_plot(df3, time = "timeUTC", yfactor = 1e9)
+png("../man/figures/obsplot_shipboardinsitu.png", width = 1500, height = 1000, res = 200)
+obs_plot(df, time = "timeUTC", yfactor = 1e9)
 dev.off()
@@
 ```{r obs_plot, fig.width=5, fig.height=3, eval = F}
-obs_plot(df4, time = "timeUTC", yfactor = 1e9)
+obs_plot(df, time = "timeUTC", yfactor = 1e9)
 ```
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_towerinsitu.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_shipboardinsitu.png?raw=true)
 
@@
 library(sf)
-dx <- df4[, 
+dx <- df[, 
     lapply(.SD, mean),
@@
 x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_towerinsitu_map.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_shipboardinsitu_map.png", width = 1500, height = 1000, res = 200)
 plot(x["value"], axes = T, reset = F)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_towerinsitu_map.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_shipboardinsitu_map.png?raw=true)

### 2024-08-26 examples/shipboard-flask.Rmd
shipboard flask

--- 2024-08-26 examples/surface-flask.Rmd
@@
 ---
-title: "surface-flask"
+title: "shipboard-flask"
 output: rmarkdown::html_vignette
 vignette: >
-  %\VignetteIndexEntry{surface-flask}
+  %\VignetteIndexEntry{shipboard-flask}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ```{r readnc, eval = T}
-datasetid <- "surface-flask"
+datasetid <- "shipboard-flask"
 df <- obs_read_nc(index = index,
@@
 
+The temporal range of data is
+```{r range_year, eval = T}
+range(df$year)
+```
+
 Now we can process the data. We first filter for observations
@@
 ```{r spatial, eval = T}
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
 yy <- 2020
@@
 
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
 dim(df)
@@
 ```{r}
-df3 <- df2[hour_st %in% evening]
+df3 <- df2
 df3[, c("solar_time", "solar_time_cut")]
-
 ```
@@
            paste0("timespan: ", yy),
-           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
-           "data: Data averaged every 20 seconds",
-           paste0("altitude: < ", max_altitude),
            paste0("hours: ", evening),
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_surfaceflask.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_shipboardflask.png", width = 1500, height = 1000, res = 200)
 obs_plot(df3, time = "timeUTC", yfactor = 1e9)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfacetflask.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_shipboardflask.png?raw=true)
 
@@
 x <- st_as_sf(dx, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_surfaceflask_map.png", width = 1500, height = 1000, res = 200)
+png("../man/figures/obsplot_shipboardflask_map.png", width = 1500, height = 1000, res = 200)
 plot(x["value"], axes = T, reset = F)
@@
 
-![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_surfaceflask_map.png?raw=true)
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_shipboardflask_map.png?raw=true)

### 2024-08-26 README.Rmd
website and readme, next fixing typos

@@
 
-There are `r nrow(index)` files in the ObsPack directory.
-The printed information also shows the total at the bottom,
-as the sum of the individual file by sector. This
-is to ensure that the sum of files is equal to the
-total number of files found, shown at the top.
-furthermore, the printed information
-also shows that there are 109 files with the
-`agl` explicitly mentioned in the name of the file.
+For each one of these dataset ids, check the articles in the documentation
 
-### Read data
-
-Once the summary is built, the function `obs_read`
-will read the files available in the
-index file previously generated.
-Here we
-selected the category "tower-insitu".
-The argument verbose prints which files are being read each time,
-by default.
-At the end, this function prints the total number
-of observations by type of altitude (agl or asl).
-
-```{r read2}
-df <- obs_read_nc(
-  index = index,
-  categories = "tower-insitu")
-
-```
-
-> intake_height or altitude_final
-
-The identification
-of the altitude and type is critical. The approach used here 
-consists of:
-
-1. Use intake_height
-2. Identify `agl` from the name of the tile.
-3. If `agl` is not present, `agl = altitude - site_elevation`.
-4. If there are some NA in elevation, will result some NA in `agl`
-5. A new column is added named `altitude_final` to store `agl` or `asl`
-6. Another column named `type_altitude` is added to identify `agl` as 0, or `asl` as 1.
-
-To maintain a coherent approach among text and NetCDF files, 
-we will add a column altitude_final and type_altitude. This
-is done in `obs_read_nc` or `obs_read`.
-
-Sometimes we need more information about the site. For instance,
-what do the observations start and end. Then, we added 
-the function `obs_table`, which calculates statistics
-summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode.
-
-```{r table}
-dft <- obs_table(df = df)
-```
-
-Now we can visualize the average of observations by site.
-
-(run it locally)
-
-```{r sf, echo = TRUE, message=FALSE, eval = FALSE}
-library(sf)
-sdft <- st_as_sf(dft[stat == "mean"], 
-                 coords = c("longitude", "latitude"), 
-                 crs = 4326)
-sdft$value <- sdft$value*1e09
-plot(sdft["value"], 
-     reset = FALSE, 
-     axes = TRUE, 
-     graticule = TRUE,
-     pch = 16, 
-     cex = 1.5, 
-     pal = cptcity::lucky(colorRampPalette = T), #random
-     main = NULL)
-maps::map(add = T)
-```
-
-Let us randomly select a couples of sites from the database.
-
-```{r uniques}
-usites <- unique(df$site_name)[sample(seq_along(unique(df$site_name)), 2)]
-usites
-```
-
-```{r table1}
-knitr::kable(dft[site_name %in% usites])
-```
-
-We added a function to plot the data read from ObsPack.
-The y-axis is the field `value` and the 
-x-axis is by default `time`. The data illustrated sorted by
-color is the field `site_code`, with the default
-number of 3 sites. The argument `pal` is to define the
-color palette, used by the internally imported function
-`cptcity::cpt`.
-
-(run it locally)
-
-```{r plot1, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack", eval = F}
-obs_plot(dt = df[site_name %in% usites], 
-         time = "time", 
-         yfactor = 1e9, 
-         cex = 0.5,
-         verbose = FALSE)
-```
-
-Here we can see `r round(nrow(df)/1000000, 2)` million observations 
-for `tower-insitu`. These observations are made between 
-`r range(df$year)[1]` and `r range(df$year)[2]`.
-
-### Filtering
-
-ObsPack includes global observations and sometimes we need to extract data
-for a specific region and periods of time.
-In this part we include spatial and temporal parameters
-to filter data.
-The year of interest is 2020, but we also included
-December of 2019 and January of 2021.
-At this stage, we can apply the spatial filter
-by using the coordinates.
-
-```{r filter}
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14
-
-yy <- 2020
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
-
-unique(df[, c("altitude_final", 
-              "site_elevation",
-              "elevation",
-              "dataset_selection_tag",
-              "site_name")])
-
-```
-
-After filtering by space and time, we have `r round(nrow(df), 2)`
-million observations. Towers can have observations at different
-heights. Here we need to select one site with the observations
-registered at the highest height. The column
-with the height is named `altitude_final` and the
-max altitude was named `max_altitude`.
-
-```{r altitude}
-dfa <- df[,
-          max(altitude_final),
-          by = site_code] |> unique()
-
-names(dfa)[2] <- "max_altitude"
-dfa
-```
-
-### Key Time
-
-Here we need to start time columns.
-The function `obs_addtime` adds time columns
-`timeUTC`, `timeUTC_start` which shows the start time
-of each observation and `timeUTC_end`
-which shows the end time for each observation.
-
-```{r addtime}
-df2 <- obs_addtime(df)
-```
-Then we need a *key_time* to aggregate data.
-This can be done using UTC, solar, or local time.
-The normal approach is using afternoon solar or local time.
-
-#### Hierarchy of solar or local time
-
-1) Solar time
-2) Local time with columns `site_utc2lst`
-3) Local time longitude
-
-> solar time (default)
-
-Here we select the hours of interest and then
-aggregate data by year, month and day of solar time.
-In this way, we will have one information per day. however
-this approach is not appropriate for aircraft
-which are aggregated every 10 or 20 seconds. Hence
-we need to aggregate data by one time column.
-Also, this helps to generate the receptor info 
-files including hour, minute and second.
-Hence, *we need to add solar or local time column*.
-
-```{r st}
-df2$solar_time <- obs_addstime(df2)
-
-```
-
-> local time with column `site_utc2lst`
-
-Then we need to identify the local time with the function
-`add_ltime`. This is important because to identifying
-observations in the evening in local time.
-`add_ltime` uses two methods, first identify
-the time difference with utc by identifying the
-metadata column "site_utc2lst". 
-If solar time is not available
-#now we need to cut solar time for the frequency needed. 
-As we will work with
-
-> local time longitude
-
-If this information
-is not available, with the aircrafts for instance,
-the local time is calculated with an approximation
-based on longitude:
-
-$$
-lt = UTC + longitude/15 * 60 * 60
-$$
-Where $lt$ is the local time, $UTC$ the time,
-$longitude$ the coordinate.
-Then, the time is cut every two hours.
-Now, we identify the local time to select evening hours.
-
-#### Cut time
-
-Now we have they key column time, we can cut it accordingly.
-
-```{r cut}
-df2$solar_time_cut <- cut(x = df2$solar_time,
-                          breaks = "1 hour") |>
-  as.character()
-
-```
-
-How we can check the solar time and the cut solar time.
-Please note that solar_time_cut,
-the column that it will be used to aggregate data
-
-How we filter for the required solar time, in this case `r evening`.
-
-```{r checkst}
-df3 <- df2[hour_st %in% evening]
-df3[, c("solar_time", "solar_time_cut")]
-
-```
-
-Now there are `r nrow(df3)` observations and before filtering
-`r nrow(df2)`.
-At this point we can calculate the averages of several columns by
-the cut time. The function `obs_agg` does this aggregation as shown
-in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`. I need to aggregate the data
-by date (year, month, date), because it is already filtered by
-the hours of interest. Then, I would have 1 observation per day.
-
-As standard, let us define `key_time` as `solar_time`. 
-The `obs_agg` function will aggregate the desired data by that column.
-
-```{r keytime}
-df3$key_time <- df3$solar_time_cut
-```
-
-```{r agg}
-df4 <- obs_agg(dt = df3,
-               cols = c("value",
-                        "latitude",
-                        "longitude"),
-               verbose = T,
-               byalt = TRUE)
-```
-
-Now there are `r nrow(df4)` observations,
-`r nrow(df) - nrow(df4)` less observations.
-Here we add the column `max_altitude` to identify the max altitude
-by site.
-
-```{r}
-df4[,
-    max_altitude := max(altitude_final),
-    by = site_code]
-df4[,
-    c("site_code",
-      "altitude_final",
-      "max_altitude")] |> unique()
-
-```
-
-### Saving master as text and csvy
-
-Now that we have all the required information, 
-we can save the files. Here, we name the data.frame 
-as master, because it contains all the information.
-This is important because some 
-fields can be used in the future, and for traceability.
-For convenience,
-time variables are transformed into character
-before writing into the disk.
-The separation is space " ".
-
-```{r, eval = T}
-master <- df4
-master$timeUTC <- as.character(master$timeUTC)
-fwrite(master, 
-       file =  paste0(tempdir(), "/tower_insitu_2020.txt"),
-       sep = " ")
-```
-
-The format Comma Separated Value with YAML (CSVY)[^3]
-consists in a typical CSV with a YAML header.
-The function`obs_write_csvy` includes the argument
-`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
-the output of the R function `str`, which
-provides a vertical summary of the data, known as
-structure.
-
-```{r, eval = T}
-csvy <- paste0(tempdir(), "/tower_insitu_2020.csvy")
-obs_write_csvy(dt = master,
-               notes = "tower 2020",
-               out = csvy
-)
-```
-
-To check the YAML header we read the first 38 lines of 
-the files that were generated. Here we can see
-the column names, type of data and first observations.
-The YAML header is delimited by the characters "---".
-
-```{r, eval = F}
-readLines(csvy)[1:34] # and more
-```
-
-### Saving receptors
-
-We need to filter some columns
-from the master files in a new object called receptors.
-This is needed because internally we run HYSPLIT [@hy]
-using the information from the receptors.
-In the case of a tower, we need to select 
-observations with the highest 
-altitude. The specific columns are selected
-as shown on the following code. We are selecting the 
-ending times, because later HYSPLIT is run backwards
-based on the time of measurement, between ending and
-starting times. The columns about time are formatted to
-have two characters. For instance, the month 1,
-is formatted as "01". We also need to filter for
-`type_altitude` equal 0, representing `agl `observations
-, or equal to 1, `asl`.
-
-```{r, eval = F}
-
-receptor <- master[altitude_final == max_altitude,
-                   c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour",
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "altitude_final",
-                     "type_altitude",
-                     "year_end",
-                     "month_end",
-                     "day_end",
-                     "hour_end",
-                     "minute_end",
-                     "second_end",
-                     "time_decimal")]
-
-receptor$altitude_final <- round(receptor$altitude_final)
-
-receptor <- obs_format(receptor)
-
-if(nrow(receptor_agl) > 0) {
-  fwrite(x = receptor_agl,
-         file = paste0(tempdir(), "/receptor_tower_insitu_2020_AGL.txt"),
-         sep = " ")}
-
-if(nrow(receptor_asl) > 0) {
-  fwrite(x = receptor_asl,
-         file = paste0(tempdir(), "/receptor_tower_insitu_2020_ASL.txt"),
-         sep = " ")}
-
-```
-
-## Application for other sectors
-In this package we are sharing scripts to process other sectors
-The scripts are available in the path 
-`https://github.com/ibarraespinosa/rtorf/tree/main/rscripts`
+https://noaa-gml.github.io/rtorf/
 

### 2024-08-26 examples/README_nc.Rmd
website and readme, next fixing typos

--- 2024-04-25 examples/README_nc.Rmd
@@
 [![R build
-status](https://github.com/ibarraespinosa/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)
+status](https://github.com/noaa-gml/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/noaa-gml/rtorf/actions)
 
-[![R-CMD-check](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml)
+[![R-CMD-check](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml)
 
@@
 
-There are `r nrow(index)` files in the ObsPack directory.
-The printed information also shows the total at the bottom,
-as the sum of the individual file by sector. This
-is to ensure that the sum of files is equal to the
-total number of files found, shown at the top.
-furthermore, the printed information
-also shows that there are 109 files with the
-`agl` explicitly mentioned in the name of the file.
+For each one of these dataset ids, check the articles in the documentation
 
-### Read data
-
-Once the summary is built, the function `obs_read`
-will read the files available in the
-index file previously generated.
-Here we selected the category "tower-insitu".
-The argument verbose prints which files are being read each time,
-by default.
-At the end, this function prints the total number
-of observations by type of altitude (agl or asl).
-
-```{r}
-df <- obs_read_nc(index = index,
-                  categories = "tower-insitu")
-
-```
-
-Sometimes we need more information about the site. For instance,
-what do the observations start and end. Then, we added 
-the function `obs_table`, which calculates statistics
-summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode.
-
-```{r}
-dft <- obs_table(df = df)
-```
-
-Now we can visualize the average of observations by site.
-
-```{r sf, echo = TRUE, message=FALSE}
-library(sf)
-sdft <- st_as_sf(dft[stat == "mean"], 
-                 coords = c("longitude", "latitude"), 
-                 crs = 4326)
-sdft$value <- sdft$value*1e09
-plot(sdft["value"], 
-     reset = FALSE, 
-     axes = TRUE, 
-     graticule = TRUE,
-     pch = 16, 
-     cex = 1.5, 
-     pal = cptcity::lucky(colorRampPalette = T), 
-     main = NULL)
-maps::map(add = T)
-```
-
-Let us randomly select a couples of sites from the database.
-
-```{r}
-usites <- unique(df$site_name)[sample(seq_along(unique(df$site_name)), 2)]
-usites
-```
-
-```{r}
-knitr::kable(dft[site_name %in% usites])
-```
-
-We added a function to plot the data read from ObsPack.
-The y-axis is the field `value` and the 
-x-axis is by default `time`. The data illustrated sorted by
-color is the field `site_code`, with the default
-number of 3 sites. The argument `pal` is to define the
-color palette, used by the internally imported function
-`cptcity::cpt`.
-
-```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
-obs_plot(dt = df[site_name %in% usites], 
-         time = "time", 
-         yfactor = 1e9, 
-         cex = 0.5)
-```
-
-Here we can see `r round(nrow(df)/1000000, 2)` million observations 
-for `tower-insitu`. These observations are made between 
-`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
-of the altitude and type is critical. The approach used here 
-consists of:
-
-1. Identify `agl` from the name of the tile.
-2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
-3. If `agl` is not present, `agl = altitude - elevation`.
-4. If there are some NA in elevation, will result some NA in `agl`
-5. A new column is added named `altitude_final` to store `agl` or `asl`
-6. Another column named `type_altitude` is added to identify `agl` or `asl`.
-7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"
-
-### Filtering
-
-ObsPack includes global observations and sometimes we need to extract data
-for a specific region and periods of time.
-In this part we include spatial and temporal parameters
-to filter data.
-The year of interest is 2020, but we also included
-December of 2019 and January of 2021.
-At this stage, we can apply the spatial filter
-by using the coordinates.
-
-```{r}
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14
-
-yy <- 2020
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-if(any(names(df) %in% "intake_height")) {
-  df <- df[intake_height < max_altitude &
-             latitude < north &
-             latitude > south &
-             longitude < east &
-             longitude > west]
-} else {
-  df <- df[altitude_final < max_altitude &
-             latitude < north &
-             latitude > south &
-             longitude < east &
-             longitude > west]
-}
-
-unique(df[, c("intake_height", "site_elevation", "elevation",
-              "dataset_selection_tag",
-              "site_name")])
-
-```
-
-After filtering by space and time, we have `r round(nrow(df), 2)`
-million observations. Towers can have observations at different
-heights. Here we need to select one site with the observations
-registered at the highest height. The column
-with the height is named `altitude_final` and the
-max altitude was named `max_altitude`.
-
-```{r}
-
-if(any(names(df) %in% "intake_height")) {
-  dfa <- df[,
-            max(intake_height),
-            by = site_code] |> unique()
-
-  names(dfa)[2] <- "max_altitude"
-} else {
-  dfa <- df[,
-            max(altitude_final),
-            by = site_code] |> unique()
-
-  names(dfa)[2] <- "max_altitude"
-}
-dfa
-```
-
-### Time
-
-Here we need to start time columns.
-The function `obs_addtime` adds time columns
-`timeUTC`, `timeUTC_start` which shows the start time
-of each observation and `timeUTC_end`
-which shows the end time for each observation.
-Then we need to identify the local time with the function
-`add_ltime`. This is important because to identifying
-observations in the evening in local time.
-`add_ltime` uses two methods, first identify
-the time difference with utc by identifying the
-metadata column "site_utc2lst". If this information
-is not available, with the aircrafts for instance,
-the local time is calculated with an approximation
-based on longitude:
-
-$$
-lt = UTC + longitude/15 * 60 * 60
-$$
-Where $lt$ is the local time, $UTC$ the time,
-$longitude$ the coordinate.
-Then, the time is cut every two hours.
-Now, we identify the local time to select evening hours.
-
-```{r}
-df2 <- obs_addtime(df)
-```
-
-```{r}
-df2 <- df2[hour_st %in% evening]
-
-```
-
-Now there are `r nrow(df2)` observations.
-At this point we can calculate the averages of several columns by
-the cut time. The function `obs_agg` does this aggregation as shown
-in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`. I need to aggregate the data
-by date (year, month, date), because it is already filtered by
-the hours of interest. Then, I would have 1 observation per day.
-
-```{r}
-df3 <- obs_agg(dt = df2,
-               gby = "mean",
-               cols = c("value",
-                        "latitude",
-                        "longitude",
-                        "type_altitude",
-                        "dif_time",
-                        "year_end",
-                        "site_utc2lst"),
-               verbose = FALSE,
-               byalt = TRUE)
-df5 <- obs_addltime(df4)
-```
-
-Now there are `r nrow(df5)` observations,
-`r nrow(df3) - nrow(df5)` less observations.
-Here we add the column `max_altitude` to identify the max altitude
-by site.
-
-```{r}
-df5[,
-    max_altitude := max(altitude_final),
-    by = site_code]
-df5[,
-    c("site_code",
-      "altitude_final",
-      "max_altitude")] |> unique()
-```
-
-### Saving master as text and csvy
-
-Now that we have all the required information, 
-we can save the files. Here, we name the data.frame 
-as master, because it contains all the information.
-This is important because some 
-fields can be used in the future, and for traceability.
-For convenience,
-time variables are transformed into character
-before writing into the disk.
-The separation is space " ".
-
-```{r, eval = T}
-master <- df5
-master$timeUTC <- as.character(master$timeUTC)
-master$timeUTC_end <- as.character(master$timeUTC_end)
-master$local_time <- as.character(master$local_time)
-
-fwrite(master, 
-       file =  paste0(tempdir(), "/tower_insitu_2020.txt"),
-       sep = " ")
-```
-
-The format Comma Separated Value with YAML (CSVY)[^3]
-consists in a typical CSV with a YAML header.
-The function`obs_write_csvy` includes the argument
-`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
-the output of the R function `str`, which
-provides a vertical summary of the data, known as
-structure.
-
-```{r, eval = T}
-csvy <- paste0(tempdir(), "/tower_insitu_2020.csvy")
-obs_write_csvy(dt = master,
-               notes = "tower 2020",
-               out = csvy
-)
-```
-
-To check the YAML header we read the first 38 lines of 
-the files that were generated. Here we can see
-the column names, type of data and first observations.
-The YAML header is delimited by the characters "---".
-
-```{r, eval = F}
-readLines(csvy)[1:38]
-```
-
-### Saving receptors
-
-We need to filter some columns
-from the master files in a new object called receptors.
-This is needed because internally we run HYSPLIT [@hy]
-using the information from the receptors.
-In the case of a tower, we need to select 
-observations with the highest 
-altitude. The specific columns are selected
-as shown on the following code. We are selecting the 
-ending times, because later HYSPLIT is run backwards
-based on the time of measurement, between ending and
-starting times. The columns about time are formatted to
-have two characters. For instance, the month 1,
-is formatted as "01". We also need to filter for
-`type_altitude` equal 0, representing `agl `observations
-, or equal to 1, `asl`.
-
-```{r, eval = F}
-
-receptor <- master[altitude_final == max_altitude,
-                   c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour",
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "altitude_final",
-                     "type_altitude",
-                     "year_end",
-                     "month_end",
-                     "day_end",
-                     "hour_end",
-                     "minute_end",
-                     "second_end")]
-receptor$altitude_final <- round(receptor$altitude_final)
-receptor <- obs_format(receptor)
-
-if(nrow(receptor_agl) > 0) {
-  fwrite(x = receptor_agl,
-         file = paste0(tempdir(), "/receptor_tower_insitu_2020_AGL.txt"),
-         sep = " ")}
-
-if(nrow(receptor_asl) > 0) {
-  fwrite(x = receptor_asl,
-         file = paste0(tempdir(), "/receptor_tower_insitu_2020_ASL.txt"),
-         sep = " ")}
-
-```
-
-## Application for other sectors
-In this package we are sharing scripts to process other sectors
-The scripts are available in the path 
-`https://github.com/ibarraespinosa/rtorf/tree/main/rscripts`
+https://noaa-gml.github.io/rtorf/
 

### 2024-08-26 examples/README_nc.Rmd
vert plot

@@
 
-![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
-
-[![R build
-status](https://github.com/noaa-gml/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/noaa-gml/rtorf/actions)
-
+![GitHub commit activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
 [![R-CMD-check](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml)
+![GitHub Repo stars](https://img.shields.io/github/stars/noaa-gml/rtorf)
 

### 2024-08-26 examples/aircore.Rmd
vert plot

@@
 
-Now we read the aircraft-pfp using the function `obs_read_nc`. To this date,
+Now we read the `aircore` using the function `obs_read_nc`. To this date,
 `solar_time` is not included for aircraft, so we `FALSE` that argument.
@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircore_map.png?raw=true)
+
+```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
+x <- df3
+x$ch4 <- x$value*1e+9
+png("../man/figures/obsplot_aircore_vert.png", width = 1500, height = 1000, res = 200)
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+dev.off()
+```
+
+```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
+x <- df3
+x$ch4 <- x$value*1e+9obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircore_vert.png?raw=true)

### 2024-08-26 examples/aircore.Rmd
revise airocore

@@
 x <- df3
-x$ch4 <- x$value*1e+9obs_plot(x, 
+x$ch4 <- x$value*1e+9
+obs_plot(x, 
          time = "ch4", 

### 2024-08-26 examples/aircraft-flask.Rmd
revise airocore

@@
 
-Now we read the aircraft-pfp using the function `obs_read_nc`. To this date,
+Now we read the `aircraft-flask` using the function `obs_read_nc`. To this date,
 `solar_time` is not included for aircraft, so we `FALSE` that argument.
@@
 
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
+# df <- df[altitude_final < max_altitude &
+#            latitude < north &
+#            latitude > south &
+#            longitude < east &
+#            longitude > west]
 dim(df)
@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask_map.png?raw=true)
+
+```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
+x <- df3
+x$ch4 <- x$value*1e+9
+png("../man/figures/obsplot_aircraftflask_vert.png", width = 1500, height = 1000, res = 200)
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+dev.off()
+```
+
+```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
+x <- df3
+x$ch4 <- x$value*1e+9
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask_vert.png?raw=true)

### 2024-08-26 examples/aircraft-flask.Rmd
vert plot

@@
 x$ch4 <- x$value*1e+9
-obs_plot(x, 
+obs_plot(x[site_code == "IAGOS"], 
          time = "ch4", 
          y = "altitude_final", 
-         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         colu = "month", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
          type = "b", 

### 2024-08-26 examples/aircraft-pfp.Rmd
vert plot

@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftpfp_map.png?raw=true)
+
+```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
+x <- df3
+x$ch4 <- x$value*1e+9
+png("../man/figures/obsplot_aircraftpfp_vert.png", width = 1500, height = 1000, res = 200)
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+dev.off()
+```
+
+```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
+x <- df3
+x$ch4 <- x$value*1e+9
+obs_plot(x[site_code == "IAGOS"], 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftpfp_vert.png?raw=true)

### 2024-08-26 examples/aircraft-pfp.Rmd
vert plot aircraft flask

@@
 x$ch4 <- x$value*1e+9
-obs_plot(x[site_code == "IAGOS"], 
+obs_plot(x, 
          time = "ch4", 
          y = "altitude_final", 
-         colu = "month", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         colu = "site_code", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
          type = "b", 

### 2024-08-26 examples/aircraft-flask.Rmd
check articles

--- 2024-08-26 examples/aircraft-flask.Rmd
@@
          y = "altitude_final", 
-         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         colu = "month", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
          type = "b", 

### 2024-08-26 examples/aircraft-insitu.Rmd
check articles

@@
 
-Now we read the aircraft-pfp using the function `obs_read_nc`. To this date,
+Now we read the `aircraft-insitu` using the function `obs_read_nc`. To this date,
 `solar_time` is not included for aircraft, so we `FALSE` that argument.
@@
 ![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftinsitu_map.png?raw=true)
+
+```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
+x <- df3
+x$ch4 <- x$value*1e+9
+png("../man/figures/obsplot_aircraftinsitu_vert.png", width = 1500, height = 1000, res = 200)
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+dev.off()
+```
+
+```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
+x <- df3
+x$ch4 <- x$value*1e+9
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
+```
+
+![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftinsitu_vert.png?raw=true)

### 2024-08-26 examples/surface-flask.Rmd
check articles

@@
 
-Now we read the surface-flask using the function `obs_read_nc`. 
+Now we read the `surface-flask` using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.
@@
            paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
-           "data: Data averaged every 20 seconds",
            paste0("altitude: < ", max_altitude),

### 2024-08-26 examples/surface-pfp.Rmd
check articles

@@
 
-Now we read the surface-pfp using the function `obs_read_nc`. 
+Now we read the `surface-pfp` using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.

### 2024-08-26 examples/tower-insitu.Rmd
check articles

@@
 
-Now we read the surface-insitu using the function `obs_read_nc`. 
+Now we read the `tower-insitu` using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.

### 2024-08-26 paper/paper/paper.Rmd
JOSS paper

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: "R Tools for ObsPack, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA' ObsPack"
tags:
  - R
  - GHG
  - NOAA
  - ObsPack
authors:
  - name: Sergio Ibarra-Espinosa
    orcid: 0000-0002-3162-1905
    affiliation: "1, 2" # (Multiple affiliations must be quoted)
  - name: Lei Hu
    orcid: 0000-0002-5624-3832
    affiliation: 2
affiliations:
 - name: Cooperative Institute for Research in Environmental Sciences, CU Boulder
   index: 1
 - name: NOAA Global Monitoring Laboratory
   index: 2
citation_author: Ibarra-Espinosa and Hu
date: 26 August 2024
year: 2017
bibliography: mybibfile.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

In this study, we present a new open-source R package `rtorf`, 
to read, process, select, and plot NOAA Observation Package (ObsPack)
data products. We use a methane ObsPack data product as an example
in this code base, but it can be easily modified to analyze ObsPack
products for other greenhouse gasses. The R package starts with 
creating a catalog of  all ObsPack files in each product. It then 
reads all files and creates one database. While reading each ObsPack 
file, it extracts site elevation and time zone information from the 
file header and calculates sampling altitude in meters above ground 
level and local time for individual samping events. Finally, it 
processes and selects observations for inverse modeling purposes. 
This package imports functions from data.table R package, which 
contains C bindings with parallel implementation via Open-MP 
[@dt]. rtorf provides functions 
to perform these tasks in a transparent and efficient way, 
supporting open-source communities in environmental sciences.

The world is experiencing an accelerated global warming
due to the accumulation of greenhouse gases (GHG)
since the industrial revolution [@us2018].
Greenhouse gas observations are critical to monitor the state of the
atmosphere, quantify present and historical emissions, and understand 
global climate change. 
During the 21th Conference of Parties (COP21), it was established the Paris 
Accord, a multilateral effort reduce greenhouse emissions
in order to limit the temperature increment of 1.5 degrees
[@rhodes20162015]. 
Methane is a greenhouse gas responsible for half of the 
temperature increase since preindustrial levels.
Furthermore, methane has a 9 years lifetime and a
global warming potential of 30 over 100 years [@epagwp], with
a current global radiative forcing of 0.650 $Wm^{-2}$ [@aggi].
Hence, in the 26 version of COP conference 
[@hunter2021glasgow], it was 
signed the Global Methane Pledge aiming reduce at least 
methane emissions 30\% from 2020 levels by 2030, with
U.S. as one of the parties [@wh]. Therefore, monitoring $CH_4$ observations,
emissions and sinks has become critical. NOAA 
ObsPack data has been used to 
support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases 
Observing SATellite (GOSAT) data. 
@saunois2016global.
At regional scale, @lu2022methane performed another studied focused
on north america using as priors local emissions inventories.

<!-- 2. Importance of greenhouse gases observations -->
<!-- 3. NOAA - mandate by congress - Obspack -->

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire,
evaluate and make available long-term records of atmospheric gases[^2].
To achieve that goal, GML gather own and other laboratories
data, releasing observation in a compendium named ObsPack
[@masarie2014obspack]. Specifically, the $CH_4$ ObsPack GLOBALVIEW+
is a comprehensive product consisting in observations from aircrafts, 
ships, surface stations, towers and aircores.
ObsPack include a descriptor named `datasetid` covering:
aircraft-pfp, aircraft-insitu, aircraft-flask, surface-insitu,
surface-flask, surface-pfp, tower-insitu, aircore, shipboard-insitu,
and shipboard-flask. ObsPack product generally contains hundreds of files, each of which 
has different sampling frequencies, hours, and attributes.
It takes time and effort to develop tools to read and process 
each ObsPack product and select observations of interest for specific 
modeling and data analysis purposes.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files [@masarie2014obspack]. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. 
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome. The `rtorf` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses. 
In this manuscript we present `rtorf`, an R package to read,
process and plot NOAA ObsPack data, a software useful and needed
for the community [@R]. For this release, we are 
focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process consists in creating a 
summary of the ObsPack files, reading them in an iteration process, 
filtering, and generating another output and plots.

# Installation

To install `rtorf`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment. `rtorf` is hosted at GitHub,
which allows the implementation of checking the package installation in a 
variety OS.

```{r install, eval = F}
remotes::install_github("noaa-gml/rtorf")
library(rtorf)
```

# Overview

`rtorf` is a collection of function organized together to read and 
process ObsPack files. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. As $CH_4$ ObsPACKGLOBALViewplus 5.1,
the product used in this manuscript, includes dataid, we produced
a guide for each of of them available at https://noaa-gml.github.io/rtorf.
Then, in this manuscript we present the processing of `aircraft-insitu`.
The obspack product in this case is `obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08`.

## Summary

We first call the libraries `rtorf` and `data.table`.  Most of objects returned
by `rtorf` are of class `data.table`. Then, we define the datasetid to be 
identified in the name of the files inside the directory with data. 
This process print a summary of the data and if the logical argument verbose
is `TRUE`, print the file being read in each iteration, default is `FALSE`.

```{r}
library(rtorf)
library(data.table)
cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

Once the index of file is built, we can read each file. As we are directing to the
nc directory in ObsPack, with NetCDF files inside, we use the function 
`obs_read_nc`. This function dumps the NetCDF information into a data.table 
with `long` format [@Silge2016]. As the global attributes is attributes in
the NetCDF would result in a data.table with too many columns, we used
the argument `att` equals to FALSE, which is default.
In ground-based datasetid the `solar_time` array is available.
This is useful to select specific observations, for more information check
the site documentation. At the moment, this array is not available for aircraft
observations, hence we select `FALSE`. In this case, we select `verbose` equal
to `TRUE`, to see the name of the files being read.

```{r}
datasetid <- "aircraft-insitu"
df <- obs_read_nc(index = index,
                  categories = datasetid,
                  att = FALSE,
                  solar_time = FALSE,
                  verbose = TRUE)
```

The resulting `data.table` contains `r ncol(df)` columns and 
`r nrow(df)` observations. 
Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Gb")`.
The data includes observations between `r range(df$year)[1]` and 
`r range(df$year)[2]`. Now we can define some parameters to filter our data,
like the year 2020 and spatially data below 8000 meters above sea level (masl) and
focused over north America.

```{r}
df <- df[year == 2020 &
           altitude_final < 8000 &
           latitude < 80 &
           latitude > 10 &
           longitude < -50 &
           longitude > -170]
```

Now we have a `data.table` that contains `r ncol(df)` columns and 
`r nrow(df)` observations. The size of `data.table` is 
`r format(object.size(df), unit = "Gb")`. We can add a column
of time in format "POSIXct" and cut the seconds every 20 seconds.
Usually, aircraft observations every 1 second. Then,
We can simplify the data by calculating averages every 20 seconds.
We perform this task by cutting time very 20 seconds. Then,
we add a new column with the mandatory name `key_time`, which 
will be used to aggregate data with "POSIXct" class, but every 20 seconds.

```{r}
df <- obs_addtime(df)
df$sec2 <- obs_freq(x = df$second,
                    freq = seq(0, 59, 20))
df$key_time <- ISOdatetime(year = df$year,
                           month = df$month,
                           day = df$day,
                           hour = df$hour,
                           min = df$minute,
                           sec = df$sec2,
                           tz = "UTC")
df[1, c("timeUTC", "key_time")]
```

now we can aggregate the data using the function `obs_agg`. The argument cols
indicate which columns will be averaged. Then, we add local time with the 
function `obs_addltime` and we re order the data.table.

```{r}
df2 <- obs_agg(df, 
               cols =  c("year", "month", "day", "hour", "minute", "second",
                         "time", "time_decimal",  "value",
                         "latitude", "longitude", "altitude_final",
                         "pressure", "u", "v", "temperature", "type_altitude"))
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("site_code", "timeUTC"),
          order = c(-1, 1))
```

## Solar or local time

Identifying the local time is important for atmospheric reasons. Sometimes
we need observations when the Planetary Boundary Layer is high,
so that the concentrations are well distributed, generaly 
around 2:00pm. In `rtorf` we use an hierarchical approach
based on the availability of critical information.
Basically, if the solar time array is available, we use
the functon `obs_addstime`. In the begative case
`rtorf` searches for the metadata `site_utc2lst` ot convert
UTC time to local. Finally, in the absence of the mentioned data,
we calculate an approximation of the local time
using the geographical coordinates, as :

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.

## Plots

Now we have the data processed and ready to be exported. `rtorf` includes 
a number of functions to save the data as takbulated format in text, csv and 
CSVY[^1] are csv files with a YAML header. This funcitons can be seen in the documentation.
In this last part of the manuscript we will show some visualizations. We included
a function named `obs_plot` which plots data in long format using R base functions.
Here we see data for the month of March 2020. This usefull function allows to plot several
sites and prints the x-axis range.

```{r, fig.width=5, fig.height=3}
obs_plot(df3[month == 3], time = "timeUTC", yfactor = 1e9, type = "b",
         xlab = "UTC time", ylab = expression(CH[4]~ppb))

```
Finally, we show some vertical profiles for the months of January and March of 2020.
We can see how during March of 2020 methane concentrations are lower than
January. This may be due the implementation of Lockdowns [@espinosa2023covid].
A manuscript focused on the impact of COVID-19 on methane emissions
will be submitted soon.

```{r}
x <- df3
x$ch4 <- x$value*1e+9
obs_plot(x, 
         time = "ch4", 
         y = "altitude_final", 
         colu = "month", 
         type = "b", 
         xlab = expression(CH[4]~ppb), 
         ylab = "altitude (m)")
```

## Future work

We are currently porting `rtorf` to python into a package named `pytorf`
with the same functionalities. Furthermore, we expect to implement
WebAsembly with Web-R[^2] in R and Python to run `rtorf` on the browser [@webr].

[^1]: https://csvy.org/
[^2]: https://docs.r-wasm.org/webr/latest/

# Acknowledgements

This project is funded by the NOAA Climate Program Office AC4 and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was supported by NOAA cooperative agreement NA22OAR4320151. Also, thanks to John Miller NOAA GML, Kenneth Schuldt NOAA GML, Kirk Thoning  NOAA GML.

# References

### 2024-08-27 paper/paper.Rmd
add paper
paper

--- 2023-04-20 README.Rmd
@@
 ---
-title: "robspack"
-output: github_document
+# Example from https://joss.readthedocs.io/en/latest/submitting.html
+title: "R Tools for ObsPack, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML
+' ObsPack"
+tags:
+  - R
+  - GHG
+  - NOAA
+  - ObsPack
+authors:
+  - name: Sergio Ibarra-Espinosa
+    orcid: 0000-0002-3162-1905
+    affiliation: "1, 2" # (Multiple affiliations must be quoted)
+  - name: Lei Hu
+    orcid: 0000-0002-5624-3832
+    affiliation: 2
+affiliations:
+ - name: Cooperative Institute for Research in Environmental Sciences, CU Boulder
+   index: 1
+ - name: NOAA Global Monitoring Laboratory
+   index: 2
+citation_author: Ibarra-Espinosa and Hu
+date: 26 August 2024
+year: 2024
+bibliography: mybibfile.bib
+output: rticles::joss_article
+csl: apa.csl
+journal: JOSS
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+# Summary
+
+In this study, we present a new open-source R package `rtorf`, 
+to read, process, select, and plot NOAA Observation Package (ObsPack)
+data products. We use a methane ObsPack data product as an example
+in this code base, but it can be easily modified to analyze ObsPack
+products for other greenhouse gasses. The R package starts with 
+creating a catalog of  all ObsPack files in each product. It then 
+reads all files and creates one database. While reading each ObsPack 
+file, it extracts site elevation and time zone information from the 
+file header and calculates sampling altitude in meters above ground 
+level and local time for individual samping events. Finally, it 
+processes and selects observations for inverse modeling purposes. 
+This package imports functions from data.table R package, which 
+contains C bindings with parallel implementation via Open-MP 
+[@dt]. rtorf provides functions 
+to perform these tasks in a transparent and efficient way, 
+supporting open-source communities in environmental sciences.
+
+The world is experiencing an accelerated global warming
+due to the accumulation of greenhouse gases (GHG)
+since the industrial revolution [@us2018].
+Greenhouse gas observations are critical to monitor the state of the
+atmosphere, quantify present and historical emissions, and understand 
+global climate change. 
+During the 21th Conference of Parties (COP21), it was established the Paris 
+Accord, a multilateral effort reduce greenhouse emissions
+in order to limit the temperature increment of 1.5 degrees
+[@rhodes20162015]. 
+Methane is a greenhouse gas responsible for half of the 
+temperature increase since preindustrial levels.
+Furthermore, methane has a 9 years lifetime and a
+global warming potential of 30 over 100 years [@epagwp], with
+a current global radiative forcing of 0.650 $Wm^{-2}$ [@aggi].
+Hence, in the 26 version of COP conference 
+[@hunter2021glasgow], it was 
+signed the Global Methane Pledge aiming reduce at least 
+methane emissions 30\% from 2020 levels by 2030, with
+U.S. as one of the parties [@wh]. Therefore, monitoring $CH_4$ observations,
+emissions and sinks has become critical. NOAA 
+ObsPack data has been used to 
+support many studies. For instance, 
+the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
+in agreement with other studies
+[@saunois2020global;@saunois2016global]
+@lu2021global, characterized global methane emissions in 
+between 2014 and 2017, including a comparison with Greenhouse gases 
+Observing SATellite (GOSAT) data. 
+@saunois2016global.
+At regional scale, @lu2022methane performed another studied focused
+on north america using as priors local emissions inventories.
+
+<!-- 2. Importance of greenhouse gases observations -->
+<!-- 3. NOAA - mandate by congress - Obspack -->
+
+The National Oceanic and Atmospheric Administration (NOAA) and its
+Global Monitoring Laboratory (GML) has the mission of acquire,
+evaluate and make available long-term records of atmospheric gases[^2].
+To achieve that goal, GML gather own and other laboratories
+data, releasing observation in a compendium named ObsPack
+[@masarie2014obspack]. Specifically, the $CH_4$ ObsPack GLOBALVIEW+
+is a comprehensive product consisting in observations from aircrafts, 
+ships, surface stations, towers and aircores.
+ObsPack include a descriptor named `datasetid` covering:
+aircraft-pfp, aircraft-insitu, aircraft-flask, surface-insitu,
+surface-flask, surface-pfp, tower-insitu, aircore, shipboard-insitu,
+and shipboard-flask. ObsPack product generally contains hundreds of files, each of which 
+has different sampling frequencies, hours, and attributes.
+It takes time and effort to develop tools to read and process 
+each ObsPack product and select observations of interest for specific 
+modeling and data analysis purposes.
+
+The NOAA ObsPack data is delivered to the public as NetCDF and 
+text files [@masarie2014obspack]. The structure of the files including descriptor fields
+depend on the type of file. For instance, the metadata
+from aircrafts is different than surface stations, but all the files
+include concentrations and other critical fields. 
+Given the complexity of ObsPack format, reading and analyzing the data
+can be cumbersome. The `rtorf` package provides the 
+GHG science and research community a transparent and efficient tool to 
+process ObsPack products for GHG modeling and analyses. 
+In this manuscript we present `rtorf`, an R package to read,
+process and plot NOAA ObsPack data, a software useful and needed
+for the community [@R]. For this release, we are 
+focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
+The general process consists in creating a 
+summary of the ObsPack files, reading them in an iteration process, 
+filtering, and generating another output and plots.
+
+# Installation
+
+To install `rtorf`, the user must have installed the R package 
+remotes and run the following script. This process will install 
+all the required dependencies, such as data.table, `cptcity`, an R 
+package with more than 7000 color palettes, and `lubridate`, a package 
+to manage time and dates [@lu;@cpt]. Then, we call the libraries
+to load the function into the environment. `rtorf` is hosted at GitHub,
+which allows the implementation of checking the package installation in a 
+variety OS.
+
+```{r install, eval = F}
+remotes::install_github("noaa-gml/rtorf")
+library(rtorf)
 ```
 
-![GitHub commit
-activity](https://img.shields.io/github/commit-activity/y/ibarraespinosa/robspack)
-[![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/R-CMD-check/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
-[![R build
-status](https://github.com/ibarraespinosa/robspack/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/robspack/actions)
+# Overview
 
-[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of green house gases observations
+`rtorf` is a collection of function organized together to read and 
+process ObsPack files. The general process 
+consists in create a 
+summary of the ObsPack files, reading them in an iteration process, 
+filter and generating another output. As $CH_4$ ObsPACKGLOBALViewplus 5.1,
+the product used in this manuscript, includes dataid, we produced
+a guide for each of of them available at https://noaa-gml.github.io/rtorf.
+Then, in this manuscript we present the processing of `aircraft-insitu`.
+The obspack product in this case is `obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08`.
 
-`robspack` only depends on `data.table`, which is basically parallel C, 
-so it can be installed in any machine.
+## Summary
 
-## installation
+We first call the libraries `rtorf` and `data.table`.  Most of objects returned
+by `rtorf` are of class `data.table`. Then, we define the datasetid to be 
+identified in the name of the files inside the directory with data. 
+This process print a summary of the data and if the logical argument verbose
+is `TRUE`, print the file being read in each iteration, default is `FALSE`.
 
-```{r, eval = F}
-remotes::install_github("ibarraespinosa/robspack")
+```{r}
+library(rtorf)
+library(data.table)
+cate = c("aircraft-pfp",
+         "aircraft-insitu",
+         "aircraft-flask",
+         "surface-insitu",
+         "surface-flask", 
+         "surface-pfp",   
+         "tower-insitu",  
+         "aircore",       
+         "shipboard-insitu",
+         "shipboard-flask")
+
+obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
+index <- obs_summary(obs = obs, 
+                     categories = cate)
 ```
 
-```{r, include = T, message=F, warning=F}
-library(robspack)
-library(data.table)
+Once the index of file is built, we can read each file. As we are directing to the
+nc directory in ObsPack, with NetCDF files inside, we use the function 
+`obs_read_nc`. This function dumps the NetCDF information into a data.table 
+with `long` format [@Silge2016]. As the global attributes is attributes in
+the NetCDF would result in a data.table with too many columns, we used
+the argument `att` equals to FALSE, which is default.
+In ground-based datasetid the `solar_time` array is available.
+This is useful to select specific observations, for more information check
+the site documentation. At the moment, this array is not available for aircraft
+observations, hence we select `FALSE`. In this case, we select `verbose` equal
+to `TRUE`, to see the name of the files being read.
+
+```{r}
+datasetid <- "aircraft-insitu"
+df <- obs_read_nc(index = index,
+                  categories = datasetid,
+                  att = FALSE,
+                  solar_time = FALSE,
+                  verbose = TRUE)
 ```
 
-## ObsPack summary
-
-The first step consists in constructing a summary for the ObsPack.
-This is required to read the data, but also, identify `agl`, 
-which is present in some of the file names.
-This function returns a `data.frame`.
-Optionally, the user can indicate a path to store the `data.frame`.
-`obs_summary` also prints a summary of the data. The second
-argument is the categories, and by default includes the categories
-shown below, to account for all the files.
-Then the summary `data.frame` contains 
-the columns `id` as the full path to each
-file, `name` which is the name or relative path of the file, `n`
-just an id, `sector` such as tower, and the column `agl`
-which indicates the `agl` indicated in the name of the file
-if available. To read the documentation of this function,
-the user must run `?obs_summary`.
+The resulting `data.table` contains `r ncol(df)` columns and 
+`r nrow(df)` observations. 
+Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Gb")`.
+The data includes observations between `r range(df$year)[1]` and 
+`r range(df$year)[2]`. Now we can define some parameters to filter our data,
+like the year 2020 and spatially data below 8000 meters above sea level (masl) and
+focused over north America.
 
 ```{r}
-
-categories <- c(
-  "aircraft-pfp",
-  "aircraft-insitu",
-  "surface-insitu",
-  "aircore",
-  "surface-pfp",
-  "tower-insitu",
-  "shipboard-insitu",
-  "flask"
-)
-obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
-index <- obs_summary(obs = obs)
-
+df <- df[year == 2020 &
+           altitude_final < 8000 &
+           latitude < 80 &
+           latitude > 10 &
+           longitude < -50 &
+           longitude > -170]
 ```
 
-There are 362 files in the ObsPack directory.
-The printed information also shows the total at the bottom,
-as the sum of the individual file by sector. This
-is to ensure that the sum of files is equal to the
-total number of files found, shown at the top.
-furthermore, the printed information
-also shows that there are 136 files with the
-`agl` explicitly mentioned in the name of the file.
+Now we have a `data.table` that contains `r ncol(df)` columns and 
+`r nrow(df)` observations. The size of `data.table` is 
+`r format(object.size(df), unit = "Gb")`. We can add a column
+of time in format "POSIXct" and cut the seconds every 20 seconds.
+Usually, aircraft observations every 1 second. Then,
+We can simplify the data by calculating averages every 20 seconds.
+We perform this task by cutting time very 20 seconds. Then,
+we add a new column with the mandatory name `key_time`, which 
+will be used to aggregate data with "POSIXct" class, but every 20 seconds.
 
-Sometimes we need more information about the site. For instance,
-what do the observations start and end. Then, we added 
-the function `obs_table`, which calculates statistics
-summary of "time" and other numeric variables by file name,
-sector, site, altitude and mode. For instance,
-the observations in the site "SCT" in
-South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
-"2020-12-31 23:30:00 UTC".
 ```{r}
-dft <- obs_table(index = index,
-                 categories = "tower-insitu",
-                 verbose = FALSE)
-dft[site_code == "SCT", ]$timeUTC |> 
-  range()
+df <- obs_addtime(df)
+df$sec2 <- obs_freq(x = df$second,
+                    freq = seq(0, 59, 20))
+df$key_time <- ISOdatetime(year = df$year,
+                           month = df$month,
+                           day = df$day,
+                           hour = df$hour,
+                           min = df$minute,
+                           sec = df$sec2,
+                           tz = "UTC")
+df[1, c("timeUTC", "key_time")]
 ```
 
-### Read data
-
-Once the summary is built, the function `obs_read`
-will read the files available in the
-index file previously generated.
-Here we selected the category "tower-insitu".
-The argument verbose prints which files are being read each time,
-by default.
-At the end, this function prints the total number
-of observations by type of altitude (agl or asl).
+now we can aggregate the data using the function `obs_agg`. The argument cols
+indicate which columns will be averaged. Then, we add local time with the 
+function `obs_addltime` and we re order the data.table.
 
 ```{r}
-df <- obs_read(index = index,
-               categories = "tower-insitu",
-               verbose = FALSE)
-
+df2 <- obs_agg(df, 
+               cols =  c("year", "month", "day", "hour", "minute", "second",
+                         "time", "time_decimal",  "value",
+                         "latitude", "longitude", "altitude_final",
+                         "pressure", "u", "v", "temperature", "type_altitude"))
+df3 <- obs_addltime(df2)
+setorderv(df3, cols = c("site_code", "timeUTC"),
+          order = c(-1, 1))
 ```
 
-We added a function to plot the data read from ObsPack.
-The y-axis is the field `value` and the 
-x-axis is by default `time`. The data illustrated sorted by
-color is the field `site_code`, with the default
-number of 3 sites. The argument `pal` is to define the
-color palette, used by the internally imported function
-`cptcity::cpt`.
+## Solar or local time
 
-```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
-obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
-```
-
-Here we can see `r round(nrow(df)/1000000, 2)` million observations 
-for `tower-insitu`. These observations are made between 
-`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
-of the altitude and type is critical. The approach used here 
-consists of:
-
-1. Identify `agl` from the name of the tile.
-2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
-3. If `agl` is not present, `agl = altitude - elevation`.
-4. If there are some NA in elevation, will result some NA in `agl`
-5. A new column is added named `altitude_final` to store `agl` or `asl`
-6. Another column named `type_altitude` is added to identify `agl` or `asl`.
-7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"
-
-### Filtering
-
-ObsPack includes global observations and sometimes we need to extract data
-for a specific region and periods of time.
-In this part we include spatial and temporal parameters
-to filter data.
-The year of interest is 2020, but we also included
-December of 2019 and January of 2021.
-At this stage, we can apply the spatial filter
-by using the coordinates.
-
-```{r}
-north <- 80
-south <- 10
-west <- -170
-east <- -50
-max_altitude <- 8000
-evening <- 14:15
-
-yy <- 2020
-df <- rbind(df[year == yy - 1 & month == 12],
-            df[year == yy],
-            df[year == yy + 1 & month == 1])
-
-df <- df[altitude_final < max_altitude &
-           latitude < north &
-           latitude > south &
-           longitude < east &
-           longitude > west]
-```
-
-After filtering by space and time, we have `r round(nrow(df), 2)`
-million observations. Towers can have observations at different
-heights. Here we need to select one site with the observations
-registered at the highest height. The column
-with the height is named `altitude_final` and the
-max altitude was named `max_altitude`.
-
-```{r}
-dfa <- df[,
-          max(altitude_final),
-          by = site_code] |> unique()
-
-names(dfa)[2] <- "max_altitude"
-```
-
-### Time
-
-Here we need to start time columns.
-The function `obs_addtime` adds time columns
-`timeUTC`, `timeUTC_start` which shows the start time
-of each observation and `timeUTC_end`
-which shows the end time for each observation.
-Then we need to identify the local time with the function
-`add_ltime`. This is important because to identifying
-observations in the evening in local time.
-`add_ltime` uses two methods, first identify
-the time difference with utc by identifying the
-metadata column "site_utc2lst". If this information
-is not available, with the aircrafts for instance,
-the local time is calculated with an approximation
-based on longitude:
+Identifying the local time is important for atmospheric reasons. Sometimes
+we need observations when the Planetary Boundary Layer is high,
+so that the concentrations are well distributed, generaly 
+around 2:00pm. In `rtorf` we use an hierarchical approach
+based on the availability of critical information.
+Basically, if the solar time array is available, we use
+the functon `obs_addstime`. In the begative case
+`rtorf` searches for the metadata `site_utc2lst` ot convert
+UTC time to local. Finally, in the absence of the mentioned data,
+we calculate an approximation of the local time
+using the geographical coordinates, as :
 
@@
 $longitude$ the coordinate.
-Then, the time is cut every two hours.
-Now, we identify the local time to select evening hours.
+
+## Plots
+
+Now we have the data processed and ready to be exported. `rtorf` includes 
+a number of functions to save the data as takbulated format in text, csv and 
+CSVY[^1] are csv files with a YAML header. This funcitons can be seen in the documentation.
+In this last part of the manuscript we will show some visualizations. We included
+a function named `obs_plot` which plots data in long format using R base functions.
+Here we see data for the month of March 2020. This usefull function allows to plot several
+sites and prints the x-axis range.
+
+```{r, fig.width=5, fig.height=3}
+obs_plot(df3[month == 3], time = "timeUTC", yfactor = 1e9, type = "b",
+         xlab = "UTC time", ylab = expression(CH[4]~ppb))
+
+```
+Finally, we show some vertical profiles for the months of January and March of 2020.
+We can see how during March of 2020 methane concentrations are lower than
+January. This may be due the implementation of Lockdowns [@espinosa2023covid].
+A manuscript focused on the impact of COVID-19 on methane emissions
+will be submitted soon.
 
 ```{r}
-df2 <- obs_addtime(df)
-df2$timeUTC <- cut(x = df2$timeUTC+3600,
-                   breaks = "2 hour") |>
-  as.character() |>
-  as.POSIXct(tz = "UTC")
-df3 <- obs_addltime(df2)
-df3 <- df3[lh %in% evening]
+x <- df3
+x$ch4 <- x$value*1e+9
+obs_plot(x, 
+         time = "ch4", 
+         y = "altitude_final", 
+         colu = "month", 
+         type = "b", 
+         xlab = expression(CH[4]~ppb), 
+         ylab = "altitude (m)")
 ```
 
-Now there are `r nrow(df3)` observations.
-At this point we can calculate the averages of several columns by
-the cut time. The function `obs_agg` does this aggregation as shown
-in the following lines of code. The argument `gby`
-establish the function used to aggregate `cols`, in this case
-the function mean by time and altitude. Finally, we
-add local time again.
+## Future work
 
-```{r}
-df4 <- obs_agg(dt = df3,
-               gby = "mean",
-               cols = c("value",
-                        "latitude",
-                        "longitude",
-                        "type_altitude",
-                        "dif_time",
-                        "year_end",
-                        "site_utc2lst"),
-               verbose = FALSE,
-               byalt = TRUE)
-df5 <- obs_addltime(df4)
-```
+We are currently porting `rtorf` to python into a package named `pytorf`
+with the same functionalities. Furthermore, we expect to implement
+WebAsembly with Web-R[^2] in R and Python to run `rtorf` on the browser [@webr].
 
-Now there are `r nrow(df5)` observations,
-`r nrow(df3) - nrow(df5)` less observations.
-Here we add the column `max_altitude` to identify the max altitude
-by site.
+[^1]: https://csvy.org/
+[^2]: https://docs.r-wasm.org/webr/latest/
 
-```{r}
-df5[,
-    max_altitude := max(altitude_final),
-    by = site_code]
-df5[,
-    c("site_code",
-      "altitude_final",
-      "max_altitude")] |> unique()
-```
+# Acknowledgements
 
-### Saving master as text and csvy
+This project is funded by the NOAA Climate Program Office AC4 and COM programs (NA21OAR4310233 / NA21OAR4310234).
+This research was supported by NOAA cooperative agreement NA22OAR4320151.
+Also, thanks to Arlyn Andrews,
+John Miller, Kenneth Schuldt, Kirk Thoning and Andy Jacobson from NOAA GML.
 
-Now that we have all the required information, 
-we can save the files. Here, we name the data.frame 
-as master, because it contains all the information.
-This is important because some 
-fields can be used in the future, and for traceability.
-For convenience,
-time variables are transformed into character
-before writing into the disk.
-The separation is space " ".
-
-```{r, eval = T}
-tmp <- tempfile()
-master <- df5
-master$timeUTC <- as.character(master$timeUTC)
-master$timeUTC_end <- as.character(master$timeUTC_end)
-master$local_time <- as.character(master$local_time)
-
-fwrite(master, 
-       file =  "paper/tower_insitu_2020.txt",
-       sep = " ")
-```
-
-The format Comma Separated Value with YAML (CSVY)[^3]
-consists in a typical CSV with a YAML header.
-The function`obs_write_csvy` includes the argument
-`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
-the output of the R function `str`, which
-provides a vertical summary of the data, known as
-structure.
-
-```{r, eval = T}
-obs_write_csvy(dt = master,
-          notes = "tower 2020",
-          out = "paper/tower_insitu_2020.csvy")
-```
-
-To check the YAML header we read the first 38 lines of 
-the files that were generated. Here we can see
-the column names, type of data and first observations.
-The YAML header is delimited by the characters "---".
-
-```{r, eval = F}
-readLines("paper/tower_insitu_2020.csvy")[1:38]
-```
-
-### Saving receptors
-
-We need to filter some columns
-from the master files in a new object called receptors.
-This is needed because internally we run HYSPLIT [@hy]
-using the information from the receptors.
-In the case of a tower, we need to select 
-observations with the highest 
-altitude. The specific columns are selected
-as shown on the following code. We are selecting the 
-ending times, because later HYSPLIT is run backwards
-based on the time of measurement, between ending and
-starting times. The columns about time are formatted to
-have two characters. For instance, the month 1,
-is formatted as "01". We also need to filter for
-`type_altitude` equal 0, representing `agl `observations
-, or equal to 1, `asl`.
-
-```{r, eval = F}
-
-receptor <- master[altitude_final == max_altitude,
-                   c("site_code",
-                     "year",
-                     "month",
-                     "day",
-                     "hour",
-                     "minute",
-                     "second",
-                     "latitude",
-                     "longitude",
-                     "altitude_final",
-                     "type_altitude",
-                     "year_end",
-                     "month_end",
-                     "day_end",
-                     "hour_end",
-                     "minute_end",
-                     "second_end")]
-receptor$altitude_final <- round(receptor$altitude_final)
-receptor <- obs_format(receptor)
-
-if(nrow(receptor_agl) > 0) {
-  fwrite(x = receptor_agl,
-         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
-  sep = " ")}
-
-if(nrow(receptor_asl) > 0) {
-  fwrite(x = receptor_asl,
-         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
-  sep = " ")}
-
-```
-
-## Recommendation for other applications
-The approach to generate receptors depends on each type of 
-observation and other considerations. For instance, aircraft with 
-continuous observations at each second can be filtered and averaged 
-every 20 seconds. In that way, the footprints are still 
-representative and it would not be necessary to run HYSPLIT 
-every second. Of course, it depends on the application 
-and objective of the study. For this manuscript, we are 
-presenting the generation of receptors based on tower observations.
-
-## Application for other sectors
-In this package we are sharing scripts to process other sectors
-The scripts are available in the path 
-`https://github.com/ibarraespinosa/robspack/tree/main/rscripts`
+# References

### 2024-09-23 R/helpers_legacy.R
adding some legacy code

@@ -1,0 +1,170 @@
+#' Helpers Legacy
+#'
+#'
+#' @title obs_id2pos
+#' @family helpers legacy
+#' @name obs_id2pos
+#' @description return ndata.frame based on footprint/receptor id
+#' @param id string
+#' @param sep string
+#' @param asdf Logical, to return as data.frame or not
+#' @return data.frame with time and location based on input
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' id <- '2002x08x03x10x45.00Nx090.00Ex00030'
+#' (obs_id2pos(id, asdf = TRUE) -> dx)
+#' id <- c('2002x08x03x10x00x45.000Nx090.000Ex00030',
+#'         '2002x08x03x10x55x45.335Sx179.884Wx00030')
+#' (obs_id2pos(id) -> dx)
+#' (obs_id2pos(id, asdf = TRUE) -> dx)
+#' (obs_id2pos(rep(id, 2)) -> dx)
+#' (obs_id2pos(rep(id, 2), asdf = TRUE) -> dx)
+#' }
+#' }
+obs_id2pos <- function(id, sep = "x", asdf = FALSE) {
+  # revers of id2pos function to read identifying label for
+  # multiple receptors (location&time) returns time as fractional
+  # julian day since 1/1/1960 and alt as altitude above ground in
+  # meters returns a 2d array for multiple receptors, a vector for
+  # a single receptor examples:
+  # id2pos('2002x08x03x10x45.00Nx090.00Ex00030') time lat lon alt
+  # yr mon day hr min 15555.42 45.00 90.00 30.00 2002 08 03 10 00
+  # id2pos(c('2002x08x03x10x00x45.000Nx090.000Ex00030','2002x08x03x10x55x45.335Sx179.884Wx00030'))
+  # time lat lon alt yr mon day hr min 15555.42 45.000 90.000 30
+  # 2002 08 03 10 00 15555.45 -45.335 -179.884 30 2002 08 03 10 55
+  # $Id: id2pos.r,v 1.6 2012/09/24 19:34:37 mellis Exp $
+  #---------------------------------------------------------------------------------------------------
+
+  # get format using separator location (assumes same format is
+  # used for all elements of id)
+  cpos <- as.numeric(gregexpr("x", id[1], fixed = TRUE)[[1]])
[…]

### 2024-09-24 R/helpers_legacy.R
add legacy functions

@@ -88,7 +88,7 @@ obs_id2pos <- function(id, sep = "x", asdf = FALSE) {
 
 
   xx <- pos[1:dim(pos)[1],
-                ,
+            ,
             drop = TRUE]
 
   if(asdf) {
@@ -98,7 +98,7 @@ obs_id2pos <- function(id, sep = "x", asdf = FALSE) {
     names(df) <- nx
     return(df)
 
-   } else {
+  } else {
     return(xx)
   }
 }
@@ -167,4 +167,85 @@ obs_julian<-function(m, d, y, origin.){
   }
   names(out) <- nms
   out
+}
+
+
+#' @title obs_info2id
+#' @family helpers legacy
+#' @name obs_info2id
+#' @description return footprint/receptor id
+#' @param yr  year
+#' @param mo  month
+#' @param dy day
+#' @param hr hour
+#' @param mn minute
+#' @param lat latitude
+#' @param lon longitude
+#' @param alt altitude above ground level
+#' @param sep character, default"x"
+#' @param long Logical, to add minute, and rounded with 2 decimals, instead of 4. default TRUE
+#' @return a string with the receptor id
+#' @export
+#' @examples {
+#' \dontrun{
+#' # Do not run
+#' obs_info2id(2002,8,3,10,45,-90,0.03) [1]
+#' }
+#' }
+obs_info2id <- function(yr,
+                        mo,
+                        dy,
+                        hr,
+                        mn = 0,
+                        lat,
+                        lon,
+                        alt,
+                        sep = "x",
+                        long = T) {
+  # function to create identifying label for single receptor
+  # (location&time) expects alt as altitude above ground in meters
+  # example: info2id(2002,8,3,10,45,-90,0.03) [1]
+  # '2002x08x03x10x+45.00x+090.00x00030' Caution: No rounding to
+  # nearest hour when long=F Cation: Arguments must be specified
+  # when long=F (cannot rely on order when mn is absent).
+  #---------------------------------------------------------------------------------------------------
+
+
+  # need leading zeros
+  hr <- substring(100 + hr, 2)  #2 digit hr
+  yr4 <- substring(10000 + yr, 2)  #4 digit year
+  mon <- substring(100 + mo, 2)  #2 digit mon
[…]

### 2024-09-24 R/helpers.R
fix obs_footname

@@ -186,11 +186,11 @@ obs_rbind <- function(dt1,
 #' data.table::fread(f, h = TRUE)
 #' }
 obs_write_csvy <- function(dt,
-                      notes,
-                      out = paste0(tempfile(), ".csvy"),
-                      sep = ",",
-                      nchar.max = 80,
-                      ...) {
+                           notes,
+                           out = paste0(tempfile(), ".csvy"),
+                           sep = ",",
+                           nchar.max = 80,
+                           ...) {
   sink(file = out)
   cat("---\n")
   cat("name: Metadata \n")
@@ -298,6 +298,7 @@ obs_trunc <- function(n, dec){
 #' @family helpers
 #' @name obs_footname
 #' @description return the expected name for the NetCDF footprint
+#' @param time POSIXct time to extract time variblaes
 #' @param year numeric number
 #' @param month numeric number
 #' @param day numeric number
@@ -306,6 +307,7 @@ obs_trunc <- function(n, dec){
 #' @param lat numeric number
 #' @param lon numeric number
 #' @param alt numeric number
+#' @param fullpath Logical, to add or not YYYY/MO/hysplit to id
 #' @param out outfile path.
 #' @param ... data.table::fwrite arguments.
 #' @note source https://stackoverflow.com/a/47015304/2418532
@@ -322,10 +324,14 @@ obs_trunc <- function(n, dec){
 #' generates the expected name.
 #'
 #' @export
-#' @examples \dontrun{
-#' # do not run
+#' @examples {
+#' obs_footname(time = Sys.time(),
+#'              lat = 1,
+#'              lon = 1,
+#'              alt = 0)
 #' }
-obs_footname <- function(year,
+obs_footname <- function(time = NULL,
+                         year,
                          month,
                          day,
                          hour,
@@ -333,6 +339,7 @@ obs_footname <- function(year,
                          lat,
                          lon,
                          alt,
+                         fullpath = FALSE,
                          out,
                          ...){
 
@@ -345,40 +352,92 @@ obs_footname <- function(year,
 
   agl <- alt
 
-  dt <- paste0(sprintf(year, fmt = '%02d'),
-               "/",
-               sprintf(month, fmt = '%02d'),
-               "/hysplit",
-               sprintf(year, fmt = '%02d'),
-               "x",
-               sprintf(month, fmt = '%02d'),
-               "x",
-               sprintf(day, fmt = '%02d'),
-               "x",
-               sprintf(hour, fmt = '%02d'),
-               "x",
[…]

### 2025-02-13 README.Rmd
adding some legacy code
README

@@
 
+Using git bash/powershell
+```{bash installb, eval = F}
+git clone https://github.com/noaa-gml/rtorf
+R CMD INSTALL rtorf
+```
+
+Using remotes R package
 ```{r install, eval = F}
@@
 
-I'm currently implementing a version in python:
+I'm currently implementing a version in python [pytorf](https://github.com/noaa-gml/pytorf):
 
@@
 ```
+## Special thanks to all the contributors
+
+[![Contributors](https://contrib.rocks/image?repo=noaa-gml/rtorf)](https://github.com/noaa-gml/rtorf/graphs/contributors)

### 2025-03-12 README.Rmd
readme
new logo! TORF!
fix #18 #16 #12, rename torf,

@@
 ---
-title: "R Tools for Obspack, Receptors and Footprints (rtorf)"
+title: "R Tools for Observations, Receptors and Footprints (rtorf)"
 output: github_document
@@
 
+<img src="man/figures/logo.png" align="right" alt="" width="220" />
+
 ![GitHub commit activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
@@
 
-[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of green house gases observations
+[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of greenhouse gases observations
 
@@
 
-obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
+obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
 index <- obs_summary(obs = obs, categories = cate)
@@
 
-```{r, include=F}
-data.frame(
-  stringsAsFactors = FALSE,
-  R = c("fex","invfile",
-        "obs_addltime","obs_addstime","obs_addtime","obs_agg",
-        "obs_find_receptors","obs_footname","obs_format","obs_freq",
-        "obs_index","obs_invfiles","obs_list.dt","obs_meta",
-        "obs_out","obs_plot","obs_rbind","obs_read","obs_read_csvy",
-        "obs_read_nc","obs_roundtime","obs_summary",
-        "obs_table","obs_trunc","obs_write_csvy","plot.invfile",
-        "print.invfile","sr","summary.invfile"),
-  description = c("File extension",
-                  "Methods for objects with class 'invfile'",
-                  "local hour (bsed on longitude and time)","Add solar time into obspack",
-                  "Add times into obspack","Aggregates ObsPack by time",
-                  "Compares expected receptors","Expected footprint name",
-                  "Formatting data","return numeric vector in intervals",
-                  "Summary of the ObsPack files (.txt)",
-                  "Generate files to perform inverse modeling","list.dt","Read obspack metadata",
-                  "outersect","Read obspack metadata","rbind obspack",
-                  "Read obspack (.txt)","reads CSVY","Read obspack (.nc)",
-                  "round seconds from \"POSIXct\" \"POSIXt\" classes",
-                  "Summary of the ObsPack files (.txt)","Obspack Table",
-                  "Trunc numbers with a desired number of decimals",
-                  "Generates YAML and write data.frame",
-                  "Methods for objects with class 'invfile'","Methods for objects with class 'invfile'",
-                  "Extacts n last characters",
-                  "Methods for objects with class 'invfile"),
-  Python = c("","","","","","","","",
-             "","","OK","","","","","","","","","OK","",
-             "","","","","","","","")
-) -> df
-
-```
-
-```{r, eval=TRUE, echo=FALSE}
-knitr::kable(df)
-
-```
 ## Special thanks to all the contributors

### 2025-03-12 examples/aircraft-pfp.Rmd
fix #18 #16 #12, rename torf,

@@
 
-obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
+obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
 index <- obs_summary(obs = obs, 
@@
                   categories = datasetid,
-                  solar_time = FALSE,
                   verbose = T)

### 2025-03-12 examples/shipboard-flask.Rmd
check articles
fix #18 #16 #12, rename torf,

@@
 
-Now we read the surface-flask using the function `obs_read_nc`. 
+Now we read the `shipboard-flask` using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.
@@
 df
-```
-
-The temporal range of data is
-```{r range_year, eval = T}
-range(df$year)
 ```

### 2025-03-12 examples/surface-insitu.Rmd
check articles
fix #18 #16 #12, rename torf,

@@
 
-Now we read the surface-insitu using the function `obs_read_nc`. 
+Now we read the `surface-insitu` using the function `obs_read_nc`. 
 `solar_time` is  included for surface, so we `TRUE` that argument.

### 2025-03-12 R/02_read_obs.R

@@ -342,7 +342,8 @@ obs_read <- function(index,
 #' "flask").
 #' @param att Logical, if global attributes should be added to data.table
 #' @param expr String expressions to filter data.table internally
-#' @param solar_time Logical, add solar time?
+#' @param solar_time Logical, add solar time? Default:
+#' if categories include aircraft, FALSE, otherwise, TRUE
 #' @param as_list Logical to return as list
 #' @param verbose Logical to show more information
 #' @param warnings Logical to show warnings when reading NetCDF, especially
@@ -362,7 +363,7 @@ obs_read_nc <- function(index,
                         categories = "flask",
                         att = FALSE,
                         expr = NULL,
-                        solar_time = TRUE,
+                        solar_time = if(grepl("aircraft", categories)) FALSE else TRUE,
                         as_list = FALSE,
                         verbose = FALSE,
                         warnings = FALSE){

### 2025-04-01 README.Rmd
readme
add capability to read CO2 and potentially other gases

@@
 
-The first step consists in constructing a summary for the ObsPack.
+The first step consists in constructing a summary for ObsPack (CH4, CO2 or other).
 This is required to read the data, but also, identify `agl`, 
@@
 
+e.g. for the `aircraft-pfp` we can read $CO_2$ and $CH_4$ [here](https://noaa-gml.github.io/rtorf/articles/aircraft-pfp.html)
+
 ## Implementation in python:
@@
 
-## Special thanks to all the contributors
+## Special thanks to all the
+
+**contributors**
 
 [![Contributors](https://contrib.rocks/image?repo=noaa-gml/rtorf)](https://github.com/noaa-gml/rtorf/graphs/contributors)
+
+and
+
+**Stargazers**
+
+<p>
+  <a href="https://github.com/noaa-gml/rtorf/stargazers">
+    <img src="http://reporoster.com/stars/dark/noaa-gml/rtorf"/>
+  </a>
+</p>

### 2025-04-01 examples/aircraft-pfp.Rmd
add capability to read CO2 and potentially other gases

@@
 
-obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
-index <- obs_summary(obs = obs, 
+obsch4 <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
+indexch4 <- obs_summary(obs = obsch4, 
                         categories = cate)
+```
+
+How let us read CO2
+
+```{r}
+obsco2 <- "Z:/obspack/obspack_co2_1_GLOBALVIEWplus_v9.1_2023-12-08/data/nc/"
+indexco2 <- obs_summary(obs = obsco2, 
+                        categories = cate)
+```
+
+We can merge co2 and ch4
+
+```{r}
+index <- rbind(indexch4,
+               indexco2)
 ```
@@
 
-now we can cut time every 20 seconds. We can chosse other frequency as well.
+now we can cut time every 20 seconds. We can choose other frequency as well.
 
@@
 df3 <- obs_addltime(df2)
-setorderv(df3, cols = c("site_code", "timeUTC"),
-          order = c(-1, 1))
+setorderv(df3, cols = c("dataset_calibration_scale", "site_code", "timeUTC"),
+          order = c(1, -1, 1))
 df3
@@
            paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
+           "gas: CO2 and CH4",
            "data: Data averaged every 20 seconds",
-           paste0("altitude: < ", max_altitude),
-           "hours: All",
-           "local_time: if var `site_utc2lst` is not available, calculated as",
-           "longitude/15*60*60 (John Miller)")
+           paste0("altitude: < ", max_altitude))
 
@@
 ```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-png("../man/figures/obsplot_aircraftpfp.png", width = 1500, height = 1000, res = 200)
-obs_plot(df3, time = "timeUTC", yfactor = 1e9)
+png("../man/figures/obsplot_aircraftpfp.png", 
+    width = 1500,
+    height = 1000,
+    res = 200)
+obs_plot(df3[dataset_calibration_scale == "WMO CO2 X2019"],
+         time = "timeUTC",
+         yfactor = 1e6)
 dev.off()
@@
 ```{r obs_plot, fig.width=5, fig.height=3, eval = F}
-obs_plot(df3, time = "timeUTC", yfactor = 1e9)
+obs_plot(df3[dataset_calibration_scale == "WMO CO2 X2019"],
+         time = "timeUTC",
+         yfactor = 1e6)
 ```
@@
 library(sf)
-x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
-png("../man/figures/obsplot_aircraftpfp_map.png", width = 1500, height = 1000, res = 200)
-plot(x["value"], axes = T, reset = F)
+x <- df3[dataset_calibration_scale == "WMO CO2 X2019",
+         mean(value*1e+6),
+         by = .(site_code,
+                latitude,
+                longitude)]
+x <- st_as_sf(x, 
+              coords = c("longitude",
+                         "latitude"), 
+              crs = 4326)
+png("../man/figures/obsplot_aircraftpfp_map.png",
+    width = 1500,
+    height = 1000,
+    res = 200)
+plot(x["V1"],
+     axes = T,
+     reset = F)
 maps::map(add = T)
@@
 library(sf)
-x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
-plot(x["value"], axes = T, reset = F)
+x <- df3[dataset_calibration_scale == "WMO CO2 X2019",
+         mean(value*1e+6),
+         by = .(site_code,
+                latitude,
+                longitude)]
+x <- st_as_sf(x, 
+              coords = c("longitude",
+                         "latitude"), 
+              crs = 4326)
+plot(x["V1"], 
+     axes = T, 
+     reset = F)
 maps::map(add = T)
@@
 ```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
-x <- df3
-x$ch4 <- x$value*1e+9
-png("../man/figures/obsplot_aircraftpfp_vert.png", width = 1500, height = 1000, res = 200)
+x <- df3[dataset_calibration_scale == "WMO CO2 X2019"]
+x$co2 <- x$value*1e+6
+png("../man/figures/obsplot_aircraftpfp_vert.png", 
+    width = 1500,
+    height = 1000,
+    res = 200)
 obs_plot(x, 
-         time = "ch4", 
+         time = "co2",
          y = "altitude_final", 
-         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
-         type = "b", 
-         xlab = expression(CH[4]~ppb), 
+         colu = "month", 
+         type = "p", 
+         xlab = expression(CO[2]~ppm), 
          ylab = "altitude (m)")
@@
 ```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
-x <- df3
-x$ch4 <- x$value*1e+9
+x <- df3[dataset_calibration_scale == "WMO CO2 X2019"]
+x$co2 <- x$value*1e+6
 obs_plot(x, 
-         time = "ch4", 
+         time = "co2", 
          y = "altitude_final", 
          colu = "site_code", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
-         type = "b", 
-         xlab = expression(CH[4]~ppb), 
+         type = "p", 
+         xlab = expression(CO[2]~ppm), 
          ylab = "altitude (m)")

### 2025-04-01 R/02_read_obs.R

@@ -435,7 +435,10 @@ obs_read_nc <- function(index,
     xx <- d[dim == 1]
 
     for(i in 1:nrow(xx)) {
-      dt[[xx$names[i]]] <- lv[[xx$names[i]]]
+      # add if to add variables with same length as ntow(dt)
+      if(length(lv[[xx$names[i]]]) == nrow(dt)) {
+        dt[[xx$names[i]]] <- lv[[xx$names[i]]]
+      }
     }
 
     if(solar_time) {

### 2025-04-01 R/04_aggregate_obs.R

@@ -3,14 +3,13 @@
 #' This function add timeUTC as POSIX class,
 #' local time and ending sampling time
 #'
-#'
-#'
 #' @param dt obspack data.table
 #' @param cols Character which defines columns to be aggregated
-#' @param byalt Logical, to aggregate by altitude or not (used in tower)
+#' @param by String with columns to be aggregated
+#' @param fn Function to be applied to the columns, default mean
 #' @param verbose logical to show more information
 #' @return A data.frame with with an index obspack.
-#' @importFrom data.table  second minute hour month year setDT setorderv mday
+#' @importFrom data.table  second minute hour month year setDT setorderv mday .SD
 #' @importFrom lubridate decimal_date
 #' @note By default add column timeUTC based on the input column `key_time`
 #' @export
@@ -36,7 +35,13 @@ obs_agg <- function(dt,
                              "v",
                              "temperature",
                              "type_altitude"),
-                    byalt = FALSE,
+                    by = c("key_time",
+                           "site_code",
+                           "altitude_final", # useful for towers. Name may change
+                           "type_altitude",
+                           "lab_1_abbr",
+                           "dataset_calibration_scale"),
+                    fn = "mean",
                     verbose = TRUE){
 
   if(!inherits(dt, "data.table")){
@@ -52,24 +57,6 @@ obs_agg <- function(dt,
   # Remember aircraft
   # Only aircraft need to be aggregated every 20 seconds
   # For hysplit, the ending times are not needed
-  .SD <- NULL
-  timecol <- NULL
-  site_code <- NULL
-  altitude_final <- NULL
-  type_altitude <- NULL
-  timeUTC <- NULL
-  key_time <- NULL
-  lab_1_abbr <- NULL
-  dataset_calibration_scale <- NULL
-  year_st <- NULL
-  month_st  <- NULL
-  day_st  <- NULL
-  intake_height  <- NULL
-  local_time  <- NULL
-  day <- NULL
-  gby <- NULL
-  # mean <- NULL
-
 
   # Here we need daily information
   # as this function assumes that the input is already preselected data
@@ -79,41 +66,20 @@ obs_agg <- function(dt,
   # or by local time
   # solar time names ending in _st
   # while local time, the POSIXct time vector `local_time`
-  # Then we have four options
 
-  if(byalt) {
-    if(verbose) cat("Selecting by alt\n")
-    # Remember tower
[…]

### 2025-04-01 R/10_obs_hysplit.R

@@ -384,6 +384,22 @@ cat("\n")
 #' @param setup Default SETUP.CFG
 #' @return A SETUP.CFG file
 #' @note The var description comes from hysplit 5.3 manual page 214
+#'  From the hysplit user guide:
+#' STILT mode
+#' The STILT model incorporates the variation of HYSPLIT developed by Lin et al.
+#' (2003 - JGR, VOL. 108, NO. D16, 4493, doi:10.1029/2002JD003161) that can be used to
+#' estimate upwind surface fluxes from atmospheric measurements.
+#' Two changes are introduced; the mass summation is divided by air density
+#' resulting in a mixing ratio output field (ICHEM=6) and the lowest
+#' concentration summation layer (concentration layer top depth) is permitted
+#' to vary with the mixed layer depth (ICHEM=9). The ICHEM=8 switch turns on
+#' both density and varying layer depth. Two text files of particle position
+#'  information (PARTICLE.DAT and PARTICLE_STILT.DAT) at each time step will
+#'  also be created unless the namelist parameter OUTDT defining the output
+#'  interval (min) is changed. PARTICLE_STILT.DAT follows the same format as STILT.
+#'  The footprint output in PARTICLE.DAT represents particles that were below
+#'  50% of the mixed layer height. The footprint in PARTICLE_STILT.DAT represent
+#'   particles that were below a user defined height (VEGHT).
 #' @export
 #' @examples {
 #' # Do not run

### 2025-05-20 paper/paper.Rmd
adding some legacy code
setup

@@
 
-```{r}
+```{r read_air}
 datasetid <- "aircraft-insitu"
@@
 `r nrow(df)` observations. The size of `data.table` is 
-`r format(object.size(df), unit = "Gb")`. We can add a column
+`r format(object.size(df), unit = "Gb")`. Sometimes the data can be alaready
+filtered every 20 seconds or for a different period of time.
+However, raw data can be availalb ein a second-by-second basis. 
+Under this case we may need to aggregate datah. In this example, we can add a column
 of time in format "POSIXct" and cut the seconds every 20 seconds.
@@
 Now we have the data processed and ready to be exported. `rtorf` includes 
-a number of functions to save the data as takbulated format in text, csv and 
+a number of functions to save the data as tabulated format in text, csv and 
 CSVY[^1] are csv files with a YAML header. This funcitons can be seen in the documentation.
@@
 ```
+# HYSPLIT
 

### 2025-05-20 paper/paper.tex
setup

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{mybibfile.bib}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
% \renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{C:/Users/sibarrae/AppData/Local/Programs/R/R-4.4.2/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\footnotesize{\sffamily Ibarra-Espinosa and Hu, (2024). R
Tools for ObsPack, Receptors and Footprints (rtorf) for processing
atmospheric observations in NOAA-GML '
ObsPack. \textit{Journal of Open Source Software}, (), . \href{https://doi.org/}{https://doi.org/}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={R Tools for ObsPack, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML ' ObsPack},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
%From Pandoc 3.1.8
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{R Tools for ObsPack, Receptors and Footprints (rtorf) for
processing atmospheric observations in NOAA-GML ' ObsPack}

        \author[1, 2]{Sergio Ibarra-Espinosa}
          \author[2]{Lei Hu}

      \affil[1]{Cooperative Institute for Research in Environmental
Sciences, CU Boulder}
      \affil[2]{NOAA Global Monitoring Laboratory}
  \date{\vspace{-5ex}}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{}{\color{linky}{Repository}} \ExternalLink
    \item \href{}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  {\bfseries Submitted:} \\
  {\bfseries Published:}

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
}

\section{Summary}\label{summary}

In this study, we present a new open-source R package \texttt{rtorf}, to
read, process, select, and plot NOAA Observation Package (ObsPack) data
products. We use a methane ObsPack data product as an example in this
code base, but it can be easily modified to analyze ObsPack products for
other greenhouse gasses. The R package starts with creating a catalog of
all ObsPack files in each product. It then reads all files and creates
one database. While reading each ObsPack file, it extracts site
elevation and time zone information from the file header and calculates
sampling altitude in meters above ground level and local time for
individual samping events. Finally, it processes and selects
observations for inverse modeling purposes. This package imports
functions from data.table R package, which contains C bindings with
parallel implementation via Open-MP (Dowle \& Srinivasan, 2021). rtorf
provides functions to perform these tasks in a transparent and efficient
way, supporting open-source communities in environmental sciences.

The world is experiencing an accelerated global warming due to the
accumulation of greenhouse gases (GHG) since the industrial revolution
(Reidmiller et al., 2018). Greenhouse gas observations are critical to
monitor the state of the atmosphere, quantify present and historical
emissions, and understand global climate change. During the 21th
Conference of Parties (COP21), it was established the Paris Accord, a
multilateral effort reduce greenhouse emissions in order to limit the
temperature increment of 1.5 degrees (Rhodes, 2016). Methane is a
greenhouse gas responsible for half of the temperature increase since
preindustrial levels. Furthermore, methane has a 9 years lifetime and a
global warming potential of 30 over 100 years (S.EPA, 2023), with a
current global radiative forcing of 0.650 \(Wm^{-2}\) (NOAA GML, 2023).
Hence, in the 26 version of COP conference (Hunter, Salzman, \& Zaelke,
2021), it was signed the Global Methane Pledge aiming reduce at least
methane emissions 30\% from 2020 levels by 2030, with U.S. as one of the
parties (U.S. White House, 2021). Therefore, monitoring \(CH_4\)
observations, emissions and sinks has become critical. NOAA ObsPack data
has been used to support many studies. For instance, the global methane
budget for the year 2017 was 596 \(Tgy^{-1}\), in agreement with other
studies (Saunois et al., 2016, 2020) Lu et al. (2021), characterized
global methane emissions in between 2014 and 2017, including a
comparison with Greenhouse gases Observing SATellite (GOSAT) data.
Saunois et al. (2016). At regional scale, Lu et al. (2022) performed
another studied focused on north america using as priors local emissions
inventories.

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire, evaluate
and make available long-term records of atmospheric gases\footnote{\url{https://docs.r-wasm.org/webr/latest/}}.
To achieve that goal, GML gather own and other laboratories data,
releasing observation in a compendium named ObsPack (Masarie, Peters,
Jacobson, \& Tans, 2014). Specifically, the \(CH_4\) ObsPack GLOBALVIEW+
is a comprehensive product consisting in observations from aircrafts,
ships, surface stations, towers and aircores. ObsPack include a
descriptor named \texttt{datasetid} covering: aircraft-pfp,
aircraft-insitu, aircraft-flask, surface-insitu, surface-flask,
surface-pfp, tower-insitu, aircore, shipboard-insitu, and
shipboard-flask. ObsPack product generally contains hundreds of files,
each of which has different sampling frequencies, hours, and attributes.
It takes time and effort to develop tools to read and process each
ObsPack product and select observations of interest for specific
modeling and data analysis purposes.

The NOAA ObsPack data is delivered to the public as NetCDF and text
files (Masarie et al., 2014). The structure of the files including
descriptor fields depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. Given the complexity
of ObsPack format, reading and analyzing the data can be cumbersome. The
\texttt{rtorf} package provides the GHG science and research community a
transparent and efficient tool to process ObsPack products for GHG
modeling and analyses. In this manuscript we present \texttt{rtorf}, an
R package to read, process and plot NOAA ObsPack data, a software useful
and needed for the community (R Core Team, 2024). For this release, we
are focused on the \(CH_4\) ObsPack GLOBALVIEW+ product. The general
process consists in creating a summary of the ObsPack files, reading
them in an iteration process, filtering, and generating another output
and plots.

\section{Installation}\label{installation}

To install \texttt{rtorf}, the user must have installed the R package
remotes and run the following script. This process will install all the
required dependencies, such as data.table, \texttt{cptcity}, an R
package with more than 7000 color palettes, and \texttt{lubridate}, a
package to manage time and dates (Grolemund \& Wickham, 2011;
Ibarra-Espinosa, 2017). Then, we call the libraries to load the function
into the environment. \texttt{rtorf} is hosted at GitHub, which allows
the implementation of checking the package installation in a variety OS.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"noaa{-}gml/rtorf"}\NormalTok{)}
\FunctionTok{library}\NormalTok{(rtorf)}
\end{Highlighting}
\end{Shaded}

\section{Overview}\label{overview}

\texttt{rtorf} is a collection of function organized together to read
and process ObsPack files. The general process consists in create a
summary of the ObsPack files, reading them in an iteration process,
filter and generating another output. As \(CH_4\) ObsPACKGLOBALViewplus
5.1, the product used in this manuscript, includes dataid, we produced a
guide for each of of them available at
\url{https://noaa-gml.github.io/rtorf}. Then, in this manuscript we
present the processing of \texttt{aircraft-insitu}. The obspack product
in this case is
\texttt{obspack\_ch4\_1\_GLOBALVIEWplus\_v5.1\_2023-03-08}.

\subsection{Summary}\label{summary-1}

We first call the libraries \texttt{rtorf} and \texttt{data.table}. Most
of objects returned by \texttt{rtorf} are of class \texttt{data.table}.
Then, we define the datasetid to be identified in the name of the files
inside the directory with data. This process print a summary of the data
and if the logical argument verbose is \texttt{TRUE}, print the file
being read in each iteration, default is \texttt{FALSE}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rtorf)}
\FunctionTok{library}\NormalTok{(data.table)}
\NormalTok{cate }\OtherTok{=} \FunctionTok{c}\NormalTok{(}\StringTok{"aircraft{-}pfp"}\NormalTok{,}
         \StringTok{"aircraft{-}insitu"}\NormalTok{,}
         \StringTok{"aircraft{-}flask"}\NormalTok{,}
         \StringTok{"surface{-}insitu"}\NormalTok{,}
         \StringTok{"surface{-}flask"}\NormalTok{, }
         \StringTok{"surface{-}pfp"}\NormalTok{,   }
         \StringTok{"tower{-}insitu"}\NormalTok{,  }
         \StringTok{"aircore"}\NormalTok{,       }
         \StringTok{"shipboard{-}insitu"}\NormalTok{,}
         \StringTok{"shipboard{-}flask"}\NormalTok{) }

\NormalTok{obs }\OtherTok{\textless{}{-}} \StringTok{"Z:/obspack/obspack\_ch4\_1\_GLOBALVIEWplus\_v5.1\_2023{-}03{-}08/data/nc/"}
\NormalTok{index }\OtherTok{\textless{}{-}} \FunctionTok{obs\_summary}\NormalTok{(}\AttributeTok{obs =}\NormalTok{ obs, }
                     \AttributeTok{categories =}\NormalTok{ cate)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Number of files of index: 429
##               sector     N
##               <char> <int>
##  1:     aircraft-pfp    40
##  2:  aircraft-insitu    15
##  3:    surface-flask   106
##  4:   surface-insitu   174
##  5:   aircraft-flask     4
##  6:          aircore     1
##  7:      surface-pfp    33
##  8:     tower-insitu    51
##  9:  shipboard-flask     4
## 10: shipboard-insitu     1
## 11:    Total sectors   429
## Detected 190 files with agl
## Detected 239 files without agl
\end{verbatim}

Once the index of file is built, we can read each file. As we are
directing to the nc directory in ObsPack, with NetCDF files inside, we
use the function \texttt{obs\_read\_nc}. This function dumps the NetCDF
information into a data.table with \texttt{long} format (Silge \&
Robinson, 2016). As the global attributes is attributes in the NetCDF
would result in a data.table with too many columns, we used the argument
\texttt{att} equals to FALSE, which is default. In ground-based
datasetid the \texttt{solar\_time} array is available. This is useful to
select specific observations, for more information check the site
documentation. At the moment, this array is not available for aircraft
observations, hence we select \texttt{FALSE}. In this case, we select
\texttt{verbose} equal to \texttt{TRUE}, to see the name of the files
being read.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{datasetid }\OtherTok{\textless{}{-}} \StringTok{"aircraft{-}insitu"}
\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{obs\_read\_nc}\NormalTok{(}\AttributeTok{index =}\NormalTok{ index,}
                  \AttributeTok{categories =}\NormalTok{ datasetid,}
                  \AttributeTok{att =} \ConstantTok{FALSE}\NormalTok{,}
                  \AttributeTok{solar\_time =} \ConstantTok{FALSE}\NormalTok{,}
                  \AttributeTok{verbose =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Searching aircraft-insitu...
## 1: ch4_above_aircraft-insitu_1_allvalid.nc
## 2: ch4_act_aircraft-insitu_428_allvalid-b200.nc
## 3: ch4_act_aircraft-insitu_428_allvalid-c130.nc
## 4: ch4_cob2003b_aircraft-insitu_59_allvalid.nc
## 5: ch4_eco_aircraft-insitu_1_allvalid.nc
## 6: ch4_hip_aircraft-insitu_59_allvalid.nc
## 7: ch4_iagos-caribic_aircraft-insitu_457_allvalid.nc
## 8: ch4_korus-aq_aircraft-insitu_428_allvalid-dc8.nc
## 9: ch4_man_aircraft-insitu_1_allvalid.nc
## 10: ch4_orc_aircraft-insitu_3_allvalid-merge10.nc
## 11: ch4_seac4rs_aircraft-insitu_428_allvalid-ER2.nc
## 12: ch4_seac4rs_aircraft-insitu_428_allvalid-dc8.nc
## 13: ch4_start08_aircraft-insitu_59_allvalid.nc
## 14: ch4_tom_aircraft-insitu_1_allvalid.nc
## 15: ch4_ugd_aircraft-insitu_1_allvalid.nc
\end{verbatim}

The resulting \texttt{data.table} contains 59 columns and 2041758
observations. Furthermore, the size of \texttt{data.table} is 1.4 Gb.
The data includes observations between 2003 and 2021. Now we can define
some parameters to filter our data, like the year 2020 and spatially
data below 8000 meters above sea level (masl) and focused over north
America.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ df[year }\SpecialCharTok{==} \DecValTok{2020} \SpecialCharTok{\&}
\NormalTok{           altitude\_final }\SpecialCharTok{\textless{}} \DecValTok{8000} \SpecialCharTok{\&}
\NormalTok{           latitude }\SpecialCharTok{\textless{}} \DecValTok{80} \SpecialCharTok{\&}
\NormalTok{           latitude }\SpecialCharTok{\textgreater{}} \DecValTok{10} \SpecialCharTok{\&}
\NormalTok{           longitude }\SpecialCharTok{\textless{}} \SpecialCharTok{{-}}\DecValTok{50} \SpecialCharTok{\&}
\NormalTok{           longitude }\SpecialCharTok{\textgreater{}} \SpecialCharTok{{-}}\DecValTok{170}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

Now we have a \texttt{data.table} that contains 59 columns and 236
observations. The size of \texttt{data.table} is 0 Gb. We can add a
column of time in format ``POSIXct'' and cut the seconds every 20
seconds. Usually, aircraft observations every 1 second. Then, We can
simplify the data by calculating averages every 20 seconds. We perform
this task by cutting time very 20 seconds. Then, we add a new column
with the mandatory name \texttt{key\_time}, which will be used to
aggregate data with ``POSIXct'' class, but every 20 seconds.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{obs\_addtime}\NormalTok{(df)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Adding timeUTC
## Adding timeUTC_start
## Adding timeUTC_end
## Found time_interval
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df}\SpecialCharTok{$}\NormalTok{sec2 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_freq}\NormalTok{(}\AttributeTok{x =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{second,}
                    \AttributeTok{freq =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{59}\NormalTok{, }\DecValTok{20}\NormalTok{))}
\NormalTok{df}\SpecialCharTok{$}\NormalTok{key\_time }\OtherTok{\textless{}{-}} \FunctionTok{ISOdatetime}\NormalTok{(}\AttributeTok{year =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{year,}
                           \AttributeTok{month =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{month,}
                           \AttributeTok{day =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{day,}
                           \AttributeTok{hour =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{hour,}
                           \AttributeTok{min =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{minute,}
                           \AttributeTok{sec =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{sec2,}
                           \AttributeTok{tz =} \StringTok{"UTC"}\NormalTok{)}
\NormalTok{df[}\DecValTok{1}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"timeUTC"}\NormalTok{, }\StringTok{"key\_time"}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                timeUTC            key_time
##                 <POSc>              <POSc>
## 1: 2020-01-08 22:59:55 2020-01-08 22:59:40
\end{verbatim}

now we can aggregate the data using the function \texttt{obs\_agg}. The
argument cols indicate which columns will be averaged. Then, we add
local time with the function \texttt{obs\_addltime} and we re order the
data.table.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df2 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_agg}\NormalTok{(df, }
               \AttributeTok{cols =}  \FunctionTok{c}\NormalTok{(}\StringTok{"year"}\NormalTok{, }\StringTok{"month"}\NormalTok{, }\StringTok{"day"}\NormalTok{, }\StringTok{"hour"}\NormalTok{, }\StringTok{"minute"}\NormalTok{, }\StringTok{"second"}\NormalTok{,}
                         \StringTok{"time"}\NormalTok{, }\StringTok{"time\_decimal"}\NormalTok{,  }\StringTok{"value"}\NormalTok{,}
                         \StringTok{"latitude"}\NormalTok{, }\StringTok{"longitude"}\NormalTok{, }\StringTok{"altitude\_final"}\NormalTok{,}
                         \StringTok{"pressure"}\NormalTok{, }\StringTok{"u"}\NormalTok{, }\StringTok{"v"}\NormalTok{, }\StringTok{"temperature"}\NormalTok{, }\StringTok{"type\_altitude"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Adding time
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df3 }\OtherTok{\textless{}{-}} \FunctionTok{obs\_addltime}\NormalTok{(df2)}
\FunctionTok{setorderv}\NormalTok{(df3, }\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(}\StringTok{"site\_code"}\NormalTok{, }\StringTok{"timeUTC"}\NormalTok{),}
          \AttributeTok{order =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\subsection{Solar or local time}\label{solar-or-local-time}

Identifying the local time is important for atmospheric reasons.
Sometimes we need observations when the Planetary Boundary Layer is
high, so that the concentrations are well distributed, generaly around
2:00pm. In \texttt{rtorf} we use an hierarchical approach based on the
availability of critical information. Basically, if the solar time array
is available, we use the functon \texttt{obs\_addstime}. In the begative
case \texttt{rtorf} searches for the metadata \texttt{site\_utc2lst} ot
convert UTC time to local. Finally, in the absence of the mentioned
data, we calculate an approximation of the local time using the
geographical coordinates, as :

\[
lt = UTC + longitude/15 * 60 * 60
\] Where \(lt\) is the local time, \(UTC\) the time, \(longitude\) the
coordinate.

\subsection{Plots}\label{plots}

Now we have the data processed and ready to be exported. \texttt{rtorf}
includes a number of functions to save the data as takbulated format in
text, csv and CSVY\footnote{\url{https://csvy.org/}} are csv files with
a YAML header. This funcitons can be seen in the documentation. In this
last part of the manuscript we will show some visualizations. We
included a function named \texttt{obs\_plot} which plots data in long
format using R base functions. Here we see data for the month of March
2020. This usefull function allows to plot several sites and prints the
x-axis range.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{obs\_plot}\NormalTok{(df3[month }\SpecialCharTok{==} \DecValTok{3}\NormalTok{], }\AttributeTok{time =} \StringTok{"timeUTC"}\NormalTok{, }\AttributeTok{yfactor =} \FloatTok{1e9}\NormalTok{, }\AttributeTok{type =} \StringTok{"b"}\NormalTok{,}
         \AttributeTok{xlab =} \StringTok{"UTC time"}\NormalTok{, }\AttributeTok{ylab =} \FunctionTok{expression}\NormalTok{(CH[}\DecValTok{4}\NormalTok{]}\SpecialCharTok{\textasciitilde{}}\NormalTok{ppb))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Found the following sites: 
## [1] IAGOS
## Plotting the following sites: 
## [1] IAGOS
\end{verbatim}

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-latex/unnamed-chunk-5-1.pdf}}
Finally, we show some vertical profiles for the months of January and
March of 2020. We can see how during March of 2020 methane
concentrations are lower than January. This may be due the
implementation of Lockdowns (Espinosa et al., 2023). A manuscript
focused on the impact of COVID-19 on methane emissions will be submitted
soon.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x }\OtherTok{\textless{}{-}}\NormalTok{ df3}
\NormalTok{x}\SpecialCharTok{$}\NormalTok{ch4 }\OtherTok{\textless{}{-}}\NormalTok{ x}\SpecialCharTok{$}\NormalTok{value}\SpecialCharTok{*}\FloatTok{1e+9}
\FunctionTok{obs\_plot}\NormalTok{(x, }
         \AttributeTok{time =} \StringTok{"ch4"}\NormalTok{, }
         \AttributeTok{y =} \StringTok{"altitude\_final"}\NormalTok{, }
         \AttributeTok{colu =} \StringTok{"month"}\NormalTok{, }
         \AttributeTok{type =} \StringTok{"b"}\NormalTok{, }
         \AttributeTok{xlab =} \FunctionTok{expression}\NormalTok{(CH[}\DecValTok{4}\NormalTok{]}\SpecialCharTok{\textasciitilde{}}\NormalTok{ppb), }
         \AttributeTok{ylab =} \StringTok{"altitude (m)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Found the following sites: 
## [1] 1 3
## Plotting the following sites: 
## [1] 1 3
\end{verbatim}

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-latex/unnamed-chunk-6-1.pdf}}

\subsection{Future work}\label{future-work}

We are currently porting \texttt{rtorf} to python into a package named
\texttt{pytorf} with the same functionalities. Furthermore, we expect to
implement WebAsembly with Web-R\footnote{\url{https://docs.r-wasm.org/webr/latest/}}
in R and Python to run \texttt{rtorf} on the browser (Moon, 2020).

\section{Acknowledgements}\label{acknowledgements}

This project is funded by the NOAA Climate Program Office AC4 and COM
programs (NA21OAR4310233 / NA21OAR4310234). This research was supported
by NOAA cooperative agreement NA22OAR4320151. Also, thanks to Arlyn
Andrews, John Miller, Kenneth Schuldt, Kirk Thoning and Andy Jacobson
from NOAA GML.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-dt}
Dowle, M., \& Srinivasan, A. (2021). \emph{Data.table: Extension of
`data.frame`}. Retrieved from
\url{https://CRAN.R-project.org/package=data.table}

\bibitem[\citeproctext]{ref-espinosa2023covid}
Espinosa, S. I., Hu, L., Miller, S., Harkins, C., McDonald, B. C., Oh,
Y., Bruhwiler, L., et al. (2023). COVID-19 impacts on the US methane
emissions. \emph{AGU23}.

\bibitem[\citeproctext]{ref-lu}
Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with
{lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3),
1--25. Retrieved from \url{https://www.jstatsoft.org/v40/i03/}

\bibitem[\citeproctext]{ref-hunter2021glasgow}
Hunter, D. B., Salzman, J. E., \& Zaelke, D. (2021). Glasgow climate
summit: Cop26. \emph{UCLA School of Law, Public Law Research Paper},
(22-02).

\bibitem[\citeproctext]{ref-cpt}
Ibarra-Espinosa, S. (2017). \emph{Cptcity: Incorporating the cpt-city
archive into r}. Retrieved from
\url{https://CRAN.R-project.org/package=cptcity}

\bibitem[\citeproctext]{ref-lu2022methane}
Lu, X., Jacob, D. J., Wang, H., Maasakkers, J. D., Zhang, Y., Scarpelli,
T. R., Shen, L., et al. (2022). Methane emissions in the united states,
canada, and mexico: Evaluation of national methane emission inventories
and 2010--2017 sectoral trends by inverse analysis of in situ
(GLOBALVIEWplus CH\textless{} sub\textgreater{}
4\textless/sub\textgreater{} ObsPack) and satellite (GOSAT) atmospheric
observations. \emph{Atmospheric Chemistry and Physics}, \emph{22}(1),
395--418.

\bibitem[\citeproctext]{ref-lu2021global}
Lu, X., Jacob, D. J., Zhang, Y., Maasakkers, J. D., Sulprizio, M. P.,
Shen, L., Qu, Z., et al. (2021). Global methane budget and trend,
2010--2017: Complementarity of inverse analyses using in situ
(GLOBALVIEWplus CH\textless{} sub\textgreater{}
4\textless/sub\textgreater{} ObsPack) and satellite (GOSAT)
observations. \emph{Atmospheric Chemistry and Physics}, \emph{21}(6),
4637--4657.

\bibitem[\citeproctext]{ref-masarie2014obspack}
Masarie, K., Peters, W., Jacobson, A., \& Tans, P. (2014). ObsPack: A
framework for the preparation, delivery, and attribution of atmospheric
greenhouse gas measurements. \emph{Earth System Science Data},
\emph{6}(2), 375--384.

\bibitem[\citeproctext]{ref-webr}
Moon, K.-W. (2020). \emph{Webr: Data and functions for web-based
analysis}. Retrieved from \url{https://CRAN.R-project.org/package=webr}

\bibitem[\citeproctext]{ref-aggi}
NOAA GML. (2023). THE NOAA ANNUAL GREENHOUSE GAS INDEX (AGGI).
\emph{NOAA}. GML. Retrieved from
\url{https://gml.noaa.gov/aggi/aggi.html}

\bibitem[\citeproctext]{ref-R}
R Core Team. (2024). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-us2018}
Reidmiller, D. R., Avery, C. W., Easterling, D. R., Kunkel, K. E.,
Lewis, K. L., Maycock, T. K., \& Stewart, B. C. (2018). Impacts, risks,
and adaptation in the united states: Fourth national climate assessment,
volume II. doi:\href{https://doi.org/doi:\%2010.7930/NCA4.2018}{doi:
10.7930/NCA4.2018}

\bibitem[\citeproctext]{ref-rhodes20162015}
Rhodes, C. J. (2016). The 2015 paris climate change conference: COP21.
\emph{Science progress}, \emph{99}(1), 97--104.

\bibitem[\citeproctext]{ref-saunois2016global}
Saunois, M., Bousquet, P., Poulter, B., Peregon, A., Ciais, P.,
Canadell, J. G., Dlugokencky, E. J., et al. (2016). The global methane
budget 2000--2012. \emph{Earth System Science Data}, \emph{8}(2),
697--751.

\bibitem[\citeproctext]{ref-saunois2020global}
Saunois, M., Stavert, A. R., Poulter, B., Bousquet, P., Canadell, J. G.,
Jackson, R. B., Raymond, P. A., et al. (2020). The global methane budget
2000--2017. \emph{Earth system science data}, \emph{12}(3), 1561--1623.

\bibitem[\citeproctext]{ref-epagwp}
S.EPA, U. (2023). Understanding global warming potentials. \emph{EPA}.
Environmental Protection Agency. Retrieved from
\url{https://www.epa.gov/ghgemissions/understanding-global-warming-potentials\#:~:text=Methane\%20(CH4)\%20is\%20estimated,more\%20energy\%20than\%20CO2.}

\bibitem[\citeproctext]{ref-Silge2016}
Silge, J., \& Robinson, D. (2016). Tidytext: Text mining and analysis
using tidy data principles in r. \emph{Journal of Open Source Software},
\emph{1}(3), 37.
doi:\href{https://doi.org/10.21105/joss.00037}{10.21105/joss.00037}

\bibitem[\citeproctext]{ref-wh}
U.S. White House. (2021, September). Joint US-EU press release on the
global methane pledge. \emph{The White House}. The United States
Government. Retrieved from
\url{https://www.whitehouse.gov/briefing-room/statements-releases/2021/09/18/joint-us-eu-press-release-on-the-global-methane-pledge/}

\end{CSLReferences}

\end{document}

### 2025-05-20 R/10_obs_hysplit.R

@@ -110,36 +110,36 @@ obs_hysplit_control <- function(df,
                                 radiactive_decay = 0,
                                 pol_res = 0,
                                 control = "CONTROL"){
-if(!missing(df)) {
-  # lat lon height
-  lat <- df$latitude
-  lon <- df$longitude
-  agl <- df$altitude
+  if(!missing(df)) {
+    # lat lon height
+    lat <- df$latitude
+    lon <- df$longitude
+    agl <- df$altitude
 
-  lat <- sprintf(lat, fmt = '%#.4f')
+    lat <- sprintf(lat, fmt = '%#.4f')
 
-  lon <- sprintf(lon, fmt = '%#.4f')
+    lon <- sprintf(lon, fmt = '%#.4f')
 
 
-  yr <- df$year
-  mo <- df$month
-  dy <- df$day
-  ho <- df$hour
-  mi <- df$minute
+    yr <- df$year
+    mo <- df$month
+    dy <- df$day
+    ho <- df$hour
+    mi <- df$minute
 
-} else {
-  # lat lon height
-  lat <- sprintf(lat, fmt = '%#.4f')
+  } else {
+    # lat lon height
+    lat <- sprintf(lat, fmt = '%#.4f')
 
-  lon <- sprintf(lon, fmt = '%#.4f')
+    lon <- sprintf(lon, fmt = '%#.4f')
 
-  yr <- year
-  mo <- month
-  dy <- day
-  ho <- hour
-  mi <- minute
+    yr <- year
+    mo <- month
+    dy <- day
+    ho <- hour
+    mi <- minute
 
-}
+  }
 
   start_loc <- paste(lat, lon, sprintf("%#.1f", agl))
 
@@ -208,7 +208,7 @@ if(!missing(df)) {
     cat(nmet)
   }
 
-cat("\n")
+  cat("\n")
 
   for(j in seq_along(met)) {
     metx <- met[j]
@@ -413,6 +413,9 @@ cat("\n")
 #' output frequency should be an even multiple of the time step and be evenly
 #' divisible into 60. In STILT emulation mode, the time step is forced to one minute.
 #' @param extra_params more parameters
+#' @param bypass_params named vector of characters to bypass all other arguments. If this
+#' list is available, only the content of this list will be used to
+#' write SETUP.CFG file and not other arguments.
 #' @param setup Default SETUP.CFG
 #' @return A SETUP.CFG file
[…]

### 2025-05-20 R/10_obs_hysplit.R
add examples for obs_hysplit_setup function to demonstrate default and bypass parameter usage

@@ -438,8 +438,13 @@ obs_hysplit_control <- function(df,
 #' @export
 #' @examples {
 #' # Do not run
+#' # default
 #' setup_file <- tempfile()
 #' obs_hysplit_setup(setup = setup_file)
+#' cat(readLines(setup_file),sep =  "\n")
+#' # bypass
+#' setup_file <- tempfile()
+#' obs_hysplit_setup(bypass_params = c(lala = 1), setup = setup_file)
 #' cat(readLines(setup_file),sep =  "\n")
 #' }
 #
@@ -475,7 +480,7 @@ obs_hysplit_setup <- function(idsp = 2,
     for(i in seq_along(bypass_params)) {
 
       bp <- paste0(" ",
-                   eval(parse(text = bypass_params[i])),
+                   names(bypass_params)[i],
                    " = ",
                    bypass_params[i],
                    ",")

### 2025-05-28 paper/paper.Rmd
fix obs_hysplit_control when use time params instead of df. I needed

@@
 # Example from https://joss.readthedocs.io/en/latest/submitting.html
-title: "R Tools for ObsPack, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML
-' ObsPack"
+title: "R Tools for Observations, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML"
 tags:
@@
 citation_author: Ibarra-Espinosa and Hu
-date: 26 August 2024
-year: 2024
+date: 21 May 2025
+year: 2025
 bibliography: mybibfile.bib
@@
 file header and calculates sampling altitude in meters above ground 
-level and local time for individual samping events. Finally, it 
+level and local time for individual events. Finally, it 
 processes and selects observations for inverse modeling purposes. 
@@
 methane emissions 30\% from 2020 levels by 2030, with
-U.S. as one of the parties [@wh]. Therefore, monitoring $CH_4$ observations,
+U.S. as one of the initial parties [@wh]. Therefore, monitoring $CH_4$ observations,
 emissions and sinks has become critical. NOAA 
@@
 @saunois2016global.
-At regional scale, @lu2022methane performed another studied focused
+At regional scale, @lu2022methane and @lei performed another studied focused
 on north america using as priors local emissions inventories.
@@
 Global Monitoring Laboratory (GML) has the mission of acquire,
-evaluate and make available long-term records of atmospheric gases[^2].
+evaluate and make available long-term records of atmospheric gases[^1].
 To achieve that goal, GML gather own and other laboratories
@@
 which allows the implementation of checking the package installation in a 
-variety OS.
+variety OS. To have a general view of rtorf, the reader can view the 
+online diagram of the package[^2].
 
@@
 remotes::install_github("noaa-gml/rtorf")
-library(rtorf)
 ```
@@
 library(data.table)
-cate = c("aircraft-pfp",
-         "aircraft-insitu",
-         "aircraft-flask",
-         "surface-insitu",
-         "surface-flask", 
-         "surface-pfp",   
-         "tower-insitu",  
-         "aircore",       
-         "shipboard-insitu",
-         "shipboard-flask")
+cate = c("aircraft-pfp","aircraft-insitu", "aircraft-flask", 
+         "surface-insitu", "surface-flask", "surface-pfp", "tower-insitu", 
+         "aircore", "shipboard-insitu","shipboard-flask")
 
@@
 datasetid <- "aircraft-insitu"
-df <- obs_read_nc(index = index,
-                  categories = datasetid,
-                  att = FALSE,
-                  solar_time = FALSE,
-                  verbose = TRUE)
+df <- obs_read_nc(index = index, categories = datasetid,
+                  att = FALSE, solar_time = FALSE, verbose = TRUE)
 ```
@@
 like the year 2020 and spatially data below 8000 meters above sea level (masl) and
-focused over north America.
+focused over north America. Nevertheless, it can be modified to any region.
 
 ```{r}
-df <- df[year == 2020 &
-           altitude_final < 8000 &
-           latitude < 80 &
-           latitude > 10 &
-           longitude < -50 &
-           longitude > -170]
+df <- df[year == 2020 & altitude_final < 8000 & latitude < 80 &
+           latitude > 10 & longitude < -50 & longitude > -170]
 ```
@@
                     freq = seq(0, 59, 20))
-df$key_time <- ISOdatetime(year = df$year,
-                           month = df$month,
-                           day = df$day,
-                           hour = df$hour,
-                           min = df$minute,
-                           sec = df$sec2,
+df$key_time <- ISOdatetime(year = df$year, month = df$month, day = df$day,
+                           hour = df$hour, min = df$minute, sec = df$sec2,
                            tz = "UTC")
@@
 df2 <- obs_agg(df, 
-               cols =  c("year", "month", "day", "hour", "minute", "second",
-                         "time", "time_decimal",  "value",
+               cols =  c("year", "month", "day", "hour", "minute",
+                         "second", "time", "time_decimal",  "value",
                          "latitude", "longitude", "altitude_final",
-                         "pressure", "u", "v", "temperature", "type_altitude"))
+                         "pressure", "u", "v", "temperature",
+                         "type_altitude"))
 df3 <- obs_addltime(df2)
@@
 we need observations when the Planetary Boundary Layer is high,
-so that the concentrations are well distributed, generaly 
-around 2:00pm. In `rtorf` we use an hierarchical approach
+so that the concentrations are well distributed, in genera 
+around 2:00pm, when planetary boundary layer is higher. 
+In `rtorf` we use an hierarchical approach
 based on the availability of critical information.
 Basically, if the solar time array is available, we use
-the functon `obs_addstime`. In the begative case
-`rtorf` searches for the metadata `site_utc2lst` ot convert
+the function `obs_addstime`. In the negative case
+`rtorf` searches for the metadata `site_utc2lst` to convert
 UTC time to local. Finally, in the absence of the mentioned data,
@@
 a number of functions to save the data as tabulated format in text, csv and 
-CSVY[^1] are csv files with a YAML header. This funcitons can be seen in the documentation.
+CSVY[^3] are csv files with a YAML header. This functions can be seen in the documentation.
 In this last part of the manuscript we will show some visualizations. We included
 a function named `obs_plot` which plots data in long format using R base functions.
-Here we see data for the month of March 2020. This usefull function allows to plot several
+Here we see data for the month of March 2020. This useful function allows to plot several
 sites and prints the x-axis range.
 
-```{r, fig.width=5, fig.height=3}
-obs_plot(df3[month == 3], time = "timeUTC", yfactor = 1e9, type = "b",
-         xlab = "UTC time", ylab = expression(CH[4]~ppb))
+```{r, eval = F}
+obs_plot(df3[month == 3], time = "timeUTC", yfactor = 1e9, 
+         type = "b", xlab = "UTC time", ylab = expression(CH[4]~ppb))
+```
 
-```
 Finally, we show some vertical profiles for the months of January and March of 2020.
@@
 
-```{r}
+```{r, eval = F}
 x <- df3
 x$ch4 <- x$value*1e+9
-obs_plot(x, 
-         time = "ch4", 
-         y = "altitude_final", 
-         colu = "month", 
-         type = "b", 
-         xlab = expression(CH[4]~ppb), 
-         ylab = "altitude (m)")
+obs_plot(x, time = "ch4",  y = "altitude_final",  colu = "month", 
+         type = "b",  xlab = expression(CH[4]~ppb),  ylab = "altitude (m)")
 ```
+
+```{r, echo=FALSE, out.width="90%", fig.cap="a) Time series, b) Monthly observations by altitude", fig.show='hold',fig.align='center'}
+knitr::include_graphics(c("paper_files/obsplot_ts.png",
+                          "paper_files/obsplot_alt.png"))
+```
+
 # HYSPLIT
+
+`rtorf` also provides functions to run HYSPLIT (Hybrid Single-Particle Lagrangian Integrated Trajectory) model [@stein2015noaa], through `obs_hysplit_control`, `obs_hysplit_ascdata`
+and `obs_hysplit_setup`. These function were designed to be used inside programs
+and run using `rslurm` for parallel processing[^4].
+This capability is particularly valuable for in-depth analysis of atmospheric observations, helping to interpret measurement data in the context of air mass histories and contributing to more robust emission quantifications and atmospheric model evaluations.
 
@@
 
-We are currently porting `rtorf` to python into a package named `pytorf`
-with the same functionalities. Furthermore, we expect to implement
-WebAsembly with Web-R[^2] in R and Python to run `rtorf` on the browser [@webr].
+We are currently porting `rtorf` to python into a package named `pytorf`[^5].
 
-[^1]: https://csvy.org/
-[^2]: https://docs.r-wasm.org/webr/latest/
+[^1]: https://gml.noaa.gov/
+[^2]: https://gitdiagram.com/noaa-gml/rtorfml
+[^3]: https://csvy.org/
+[^4]: https://noaa-gml.github.io/rtorf/articles/hysplit.ht
+[^5]: https://github.com/noaa-gml/pytorf
 
@@
 This research was supported by NOAA cooperative agreement NA22OAR4320151.
-Also, thanks to Arlyn Andrews,
-John Miller, Kenneth Schuldt, Kirk Thoning and Andy Jacobson from NOAA GML.
+Also, thanks to Arlyn Andrews, John Miller, Kenneth Schuldt, Kirk Thoning and Andy Jacobson from NOAA GML.
+
+```{r, echo=FALSE, message=FALSE }
+invisible(system("cp paper.md ../"))
+```
 

### 2025-05-28 R/10_obs_hysplit.R

@@ -138,6 +138,7 @@ obs_hysplit_control <- function(df,
     dy <- day
     ho <- hour
     mi <- minute
+    agl <- alt
 
   }
 

### 2025-07-23 paper/paper.Rmd
paper

@@
 `r nrow(df)` observations. 
-Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Gb")`.
+Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Mb")`.
 The data includes observations between `r range(df$year)[1]` and 
@@
 `r nrow(df)` observations. The size of `data.table` is 
-`r format(object.size(df), unit = "Gb")`. Sometimes the data can be alaready
+`r format(object.size(df), unit = "Mb")`. Sometimes the data can be already
 filtered every 20 seconds or for a different period of time.
-However, raw data can be availalb ein a second-by-second basis. 
-Under this case we may need to aggregate datah. In this example, we can add a column
+However, raw data can be available in a second-by-second basis. 
+Under this case we may need to aggregate data. In this example, we can add a column
 of time in format "POSIXct" and cut the seconds every 20 seconds.
@@
 
+Now we have the data processed and ready to be exported. `rtorf` includes 
+a number of functions to save the data as tabulated format in text, csv and 
+CSVY[^3] are csv files with a YAML header. These functions can be seen in the documentation.
+
 ## Plots
 
-Now we have the data processed and ready to be exported. `rtorf` includes 
-a number of functions to save the data as tabulated format in text, csv and 
-CSVY[^3] are csv files with a YAML header. This functions can be seen in the documentation.
 In this last part of the manuscript we will show some visualizations. We included
@@
 Finally, we show some vertical profiles for the months of January and March of 2020.
-We can see how during March of 2020 methane concentrations are lower than
-January. This may be due the implementation of Lockdowns [@espinosa2023covid].
+"We can see that methane concentrations during March of 2020 are lower than in January.
+This may be due the implementation of Lockdowns [@espinosa2023covid].
 A manuscript focused on the impact of COVID-19 on methane emissions
@@
 
-```{r, echo=FALSE, out.width="90%", fig.cap="a) Time series, b) Monthly observations by altitude", fig.show='hold',fig.align='center'}
+```{r, echo=FALSE, out.width="90%", fig.cap=" Example visualizations generated by a) Time series of CH₄ concentrations for the IAGOS site on March 2, 2020.
+(b) Vertical profiles of CH₄ concentrations from January (labeled '1') and March (labeled '3') 2020, illustrating the package's ability to compare different time periods.", fig.show='hold',fig.align='center'}
 knitr::include_graphics(c("paper_files/obsplot_ts.png",

### 2025-07-24 R/11_obs_convolve.R
adding foot to obs_convolve

@@ -12,7 +12,7 @@
 #' @param flux_format string with date format 'Ymd.nc'
 #' @param factor number to multiply fluxes.
 #' @param fn string with function to aggregate convolved fluxes, e.g. `mean`, `sum`, `max`, etc.
-#' @param as_list Logical, to retunr list of arrays instead of writting to NetCDF
+#' @param as_list Logical, to return list of arrays
 #' @param verbose Logical, to display more information
 #' @note main assumption is that fluxes have same spatial dimensions as footprints
 #' @export
@@ -197,12 +197,13 @@ obs_convolve <- function(foot_path = "AAA",
       conv_ocn <- apply(conv_ocn, c(1,2), fn)
       conv_fossil <- apply(conv_fossil, c(1,2), fn)
       conv_fire <- apply(conv_fire, c(1,2), fn)
-
+      foot <- apply(foot, c(1,2), fn)
     }
 
 
     if(!as_list)  {
       simplify2array(list(
+        foot = foot,
         conv_total = conv_total,
         conv_bio = conv_bio,
         conv_ocn = conv_ocn,
@@ -212,6 +213,7 @@ obs_convolve <- function(foot_path = "AAA",
       return(out)
     } else{
       return(list(
+        foot = foot,
         conv_total = conv_total,
         conv_bio = conv_bio,
         conv_ocn = conv_ocn,

### 2025-09-16 R/11_obs_convolve.R
add na.rm in obs_convolve fn

@@ -298,8 +298,8 @@ obs_convolve <- function(foot_path = "AAA",
 
   if(!is.null(fn)) {
 
-    conv <- apply(conv, c(1,2), fn)
-    foot <- apply(foot, c(1,2), fn)
+    conv <- apply(conv, c(1,2), fn, na.rm = T)
+    foot <- apply(foot, c(1,2), fn, na.rm = T)
   }
 
 

### 2025-09-17 README.Rmd
addding legacy code to be refactored in the future. Keeping here for tracking

@@
 </p>
+
+### Note about legacy code.
+
+I received the task of reformat legacy code used to read PARTICLE.DAT and generate
+NetCDF. This code is really old and since STILT and other have permissive licenses,
+it is here, however, some code is not exported to NAMESPACE. 
+A future version of rtorf, will have a new design for the legacy code,
+a modern and efficient approach, keeping the essential tasks.

### 2025-09-17 R/helpers_legacy.R

@@ -1088,42 +1088,42 @@ obs_POSIX_to_decimal <- function(posix.dates,
   # internally this is kept as a standard numeric quantity, generally
   # R*8.)
 
-if(legacy) {
+  if(legacy) {
 
-  if(!any(class(posix.dates)=="POSIXt")) {
-    stop("obs_POSIX_to_decimal Expecting one of class(posix.dates) == 'POSIXt'")
+    if(!any(class(posix.dates)=="POSIXt")) {
+      stop("obs_POSIX_to_decimal Expecting one of class(posix.dates) == 'POSIXt'")
+    }
+
+    #  retval <- numeric(0)
+    #  ndates <- length(posix.dates)
+
+    #  for (idate in 1:ndates) {
+    #    year <- as.numeric(format(posix.dates[idate],format="%Y"))
+    #    y0 <- as.numeric(ISOdatetime(year,1,1,0,0,0,tz=tz))
+    #    y1 <- as.numeric(ISOdatetime(year+1,1,1,0,0,0,tz=tz))
+    #    retval <- c(retval,year+(as.numeric(posix.dates[idate])-y0)/(y1-y0))
+    #  }
+
+    # alternative code as of 3 Aug 2009:  much faster, as there is no looping.
+
+    #   There's a potential problem here in that I'm not sure how the R
+    #   date-time classes handle a vector of times with different time
+    #   zone attributes among the members.  It would appear that the
+    #   attribute belonging to the vector overrides any such attribute
+    #   assigned to a member.  This is probably because it is the vector
+    #   object which can have attributes, not the member.  This code
+    #   assumes that all members of the vector have the same time zone
+    #   attribute.
+
+
+    lt <- as.POSIXlt(posix.dates)
+    y0 <- as.numeric(ISOdatetime(lt$year+1900,1,1,0,0,0,tz=attributes(lt)$tzone))
+    y1 <- as.numeric(ISOdatetime(lt$year+1901,1,1,0,0,0,tz=attributes(lt)$tzone))
+    return(lt$year+1900+(as.numeric(posix.dates)-y0)/(y1-y0))
+  } else {
+    return(lubridate::decimal_date(posix.dates,
+                                   tz = tz))
   }
-
-  #  retval <- numeric(0)
-  #  ndates <- length(posix.dates)
-
-  #  for (idate in 1:ndates) {
[…]

### 2025-11-12 README.Rmd
fix obs_hysplit_control and add verbose

@@
 library(data.table)
+utils::packageVersion("rtorf")
 ```
@@
 
-cate = c("aircraft-pfp",
+cate = c(
+  "aircraft-pfp",
   "aircraft-insitu",
@@
   "shipboard-insitu",
-         "shipboard-flask")
+  "shipboard-flask"
+)
 

### 2025-11-12 R/10_obs_hysplit.R

@@ -63,6 +63,7 @@
 #' @param radiactive_decay days
 #' @param pol_res Pollutant Resuspension (1/m)
 #' @param control name of the file, default "CONTROL"
+#' @param verbose logical, to show more information
 #' @return A CONTROL file
 #' @export
 #' @examples {
@@ -127,7 +128,8 @@ obs_hysplit_control <- function(
   wrhcicbc = c(0, 0, 0),
   radiactive_decay = 0,
   pol_res = 0,
-  control = "CONTROL"
+  control = "CONTROL",
+  verbose = FALSE
 ) {
   if (!missing(df)) {
     if (
@@ -175,6 +177,10 @@ obs_hysplit_control <- function(
 
   start_loc <- paste(lat, lon, sprintf("%#.1f", agl))
 
+  if (nchar(yr) != 4) {
+    stop("year must have for digits")
+  }
+
   rel_start <- paste(
     substr(yr, 3, 4), #I need to confirm
     sprintf(mo, fmt = '%02d'),
@@ -183,24 +189,38 @@ obs_hysplit_control <- function(
     sprintf(mi, fmt = '%02d')
   )
 
+  if (verbose) {
+    print(paste0("start: ", rel_start))
+  }
+
   if (missing(start_time)) {
     start_time <- rel_start
   }
 
   if (!is.character(start_time)) {
-    stop("start_time neds to be hcaracter with format yy mm dd hh mm")
+    stop("start_time neds to be character with format yy mm dd hh mm")
   }
   nmodels <- length(met)
 
   hydate <- as.Date(ISOdate(
     year = yr,
     month = mo,
-    day = dy + 1,
+    day = dy,
     hour = ho,
     min = mi,
     sec = 0,
     tz = "UTC"
   ))
+
+  if (verbose) {
+    print(paste0("hydate: ", hydate))
+  }
+
+  if (verbose) {
+    print(paste("location:", lat, lon, sprintf("%#.1f", agl)))
+  }
+
+  hydate <- hydate + 1
 
   # magick ####
 
